### 2016-01-18 hpc_101/paper.tex
Initial Commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Template Heavily modified by David Schanzenbach
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------
%\documentclass[t]{beamer}
\documentclass[t,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{etoolbox}
\usepackage{lmodern}% http://ctan.org/pkg/lm
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{anyfontsize}
\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{calc}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usetheme{ITS_CI}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\let\Tiny=\tiny
\definecolor{links}{HTML}{2A1B81}
\hypersetup{colorlinks,linkcolor=,urlcolor=links}

\makeatletter
\newcommand{\setlistspacing}[2]{\def\@ld{#1}\expandafter\def\csname
@list\romannumeral\@ld \endcsname{\leftmargin\csname
leftmargin\romannumeral\@ld \endcsname
              \topsep    #2
              \parsep    0\p@   \@plus\p@
              \itemsep   #2}}
\makeatother

\makeatletter
\patchcmd{\beamer@sectionintoc}{\vskip1.5em}{\vskip0.5em}{}{}
\makeatother

\newlength{\okinalen}
\setlength{\okinalen}{\widthof{'}}
\newcommand{\okina}{\hbox to.666\okinalen{\hss`\hss}}
\newcommand{\ctilde}{{\fontfamily{ptm}\selectfont\texttildelow}}

\newcommand{\regtrademark}{\fontsize{5}{6}\selectfont \textsuperscript{\textregistered}}
\newcommand{\numlessfootnotetxt}[1]{ \let\thefootnote\relax\footnotetext{#1}}

\newcommand{\semitransp}[2][35]{\color{fg!#1}#2}
\newcommand{\btVFill}{\vskip0pt plus 1filll}
\newcommand{\ddash}{-{}-}

\newcommand{\hawaii}{Hawai{\okina}i}
\newcommand{\lustre}{Lustre{\regtrademark}}
\newcommand{\intel}{Intel{\regtrademark}}
\newcommand{\cray}{Cray{\regtrademark}}

\AtBeginSection[]
{
   \begin{frame}
       \frametitle{Outline}
       \tableofcontents[currentsection, hideallsubsections]
   \end{frame}
}

%\AtBeginSubsection[]
%{
%   \begin{frame}
%       \frametitle{Outline}
%       \tableofcontents[currentsection,currentsubsection]
%   \end{frame}
%}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[On-boarding]{HPC On-boarding} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
Department of Cyberinfrastructure \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Outline} % Table of contents slide, comment this block out to remove it
\tableofcontents[hideallsubsections] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

\section{Introduction}

\subsection{Parallel Computing}
\begin{frame}
	\frametitle{Parallel Computing}
	\begin{itemize}
		\item High Performance Compute
		\begin{itemize}
			\item Each separate process can send and receive data amongst other processes (MPI \& OpenMP)
			\item If processes must communicate for the overall program to proceed, high-speed networking is needed (only MPI)
		\end{itemize}
		\item High Throughput Compute
		\begin{itemize}
			\item \emph{Pleasantly Parallel} -- Processes are independent and no communication is necessary 
		\end{itemize}
	\end{itemize}
\end{frame}

\section{University of Hawai'i Cluster}

\subsection{Cray CS300}
\begin{frame}
\frametitle{Cray CS300 -- History}
\begin{itemize}
		\item Brought online Fall 2014 -- Spring 2015
		\item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
		\item As of October 2015, more than 200 users have been granted access to the cluster
	\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{Cray CS300 -- Compute Nodes}
	\begin{itemize}
		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
			\begin{itemize}
				\item Two 10 core {\intel} processors (\emph{20 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx128$GB of useable RAM
			\end{itemize}
		\item 6 large memory nodes
			\begin{itemize}
				\item Four 10 core {\intel} processors (\emph{40 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx1$TB of useable RAM
			\end{itemize}			
		\item CentOS Linux
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage}
	Two storage options are currently available on the Cray CS300.
	\begin{enumerate}
		\item {\lustre}
		\item ValueStorage
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item {\lustre} is a high performance parallel filesystem
		\item The Cray CS300 has $\approx582$TB of storage space
		\item Shared between all compute nodes and login nodes
		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage $\rightarrow$ ValueStorage}
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Currently only available for purchase by cluster users and only accessible via the login nodes
		\item ValueStorage will eventually be available for purchase by everyone
		\item ValueStorage owners will eventually be able to mount as a network drive (CIFS) on laptop, workstations, servers
		\item Purchased in 0.5TB increments
	\end{itemize}
	\btVFill

	\begin{center}
	\textbf{ValueStorage Pricing}~\\ 
		\resizebox{0.75\textwidth}{!}{%
		\begin{tabular}{l r ||l r }
			\toprule
			\textbf{Product} & \textbf{Annual Cost}~~ & ~~\textbf{Product} & \textbf{Annual Cost}\\
			\midrule
			\midrule
			0.5TB & \$65.00~~ & ~~0.5TB + Replication  & \$130.00 \\
			\bottomrule
		\end{tabular}
		}
		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
		~\\	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

\subsubsection{Layout}
\begin{frame}
	\frametitle{Cray CS300 -- Layout}
	\includegraphics[width=0.95\textwidth]{images/high_level}
\end{frame}

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for all Faculty, Staff, and Students affiliate with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, lm.q, sb.q, kill.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

\subsection{Ownership Models}

\subsubsection{Condo Model}
\begin{frame}
	\frametitle{Condo Model}
	\begin{itemize}
		\item The Condo model allows users to buy nodes (\emph{condos}) to incorporate into the cluster	
		\item Node owners are provided with priority access to their purchased hardware
		\item All nodes have a \emph{5 year warranty}
		\begin{itemize}
			\item Once a nodes warranty has expired, it will be removed from the cluster
		\end{itemize}
		\item Condo owners are given early access to purchased nodes
		  \begin{itemize}
                    \item Access is granted as soon as the funds land in our accounts
		    \item The 5 year warranty will not begin until newly ordered nodes are installed
		\end{itemize}
		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
	\end{itemize}
\end{frame}

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In some cases, users will not require owning a node, but still need priority access
	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
	 \item SUs come in two variates:
		 \begin{itemize}
			 \item[--] Standard node units 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
\end{frame}

\subsubsection{Price Card}
\begin{frame}
	\frametitle{Price Card}
	\begin{center}
	{\large \textbf{Condo Price Card}} \\
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{l r ||l r}
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
			\midrule
			\midrule
			Standard & \$6,600.00  & Standard + 2 GPUs (Nvidia{\regtrademark} K40) & \$13,600.00 \\
			\hline
			Large Memory & \$33,900.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
			\bottomrule
		\end{tabular}
		}
	\end{center}
	\smallskip
	\begin{center}
		{\large \textbf{Service Unit Price Card}}\\
		\begin{tabular}{l c r }
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
			\midrule
			\midrule
			Standard node & \$0.50 per SU &  1,000~SU (\$500.00)\\
			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
			\bottomrule
		\end{tabular}
		\btVFill
	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

\section{Cluster Interaction}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item{Connecting to a cluster @ UH}
			\begin{itemize}
			\item Login to the cluster
			\item Verify user permissions
			\end{itemize}
    {\semitransp[25]{\item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compiliers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Connecting to a cluster @ UH}
\begin{frame}
	\frametitle{Connecting to a cluster @ UH}
	\begin{description}[\setlength{\leftmargini}{0pt}]
	\item[] \begin{itemize}
		\item To connect to the cluster, we utilize a client which communicates using the Secure Shell (\emph{SSH}) protocol 
		\item Linux and MacOSX, typically have a SSH client already installed
		\item Windows typically does not come with an SSH client installed
		\item Windows 10 may come pre-installed with a SSH client, but it might not be stable
		\item Suggested SSH clients for Windows include:
		\begin{itemize}
			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
		\end{itemize}
	\end{itemize}
	\item[] ~\\
	\item[] \begin{itemize}
		\item The Cray CS300 has two login nodes:
		\begin{itemize}
			\item uhhpc1.its.hawaii.edu
			\item uhhpc2.its.hawaii.edu
		\end{itemize}
	\end{itemize}
	\end{description}
	{\large Let's attempt to login!}
\end{frame}

\subsubsection{Login to the Cray CS300}
\begin{frame}
\frametitle{Login to the Cray CS300}
	\begin{center}\textbf{Windows}\end{center}
	\hrule~\\
	\begin{itemize}
		\item If SSH 3.2.9 installed (Lab PCs have it installed)
		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
		\item Click ``Quick Connect'' and enter the following information:
			\begin{itemize}
			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
			\item[] \textbf{User Name:} Your UH User name e.g., user99
			\item[] \textbf{Port:} 22
			\end{itemize}
		\item Press ``Connect''
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Login to the Cray CS300}
	\begin{center}\textbf{Mac \& Linux}\end{center}
	\hrule~\\
	\begin{itemize}
		\item Open a terminal window
		\item Enter one of the following:
		\begin{itemize}
			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
		\end{itemize}
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{On Initial Login~\ldots}
	Validate that all system permissions are correct for your user~\\
	\begin{enumerate}
		\item Test that you can list files in your home: `\texttt{ls -la}'    
		\item Test making a file in your home: `\texttt{touch test.txt}'    
		\item Go into \ctilde{}/lus: `\texttt{cd \ctilde{}/lus/}'  
		\item Test making a file in your lus directory: `\texttt{touch test.txt}'   
		\item Go into \ctilde{}/apps:  `\texttt{cd \ctilde{}/apps/}'     
		\item Test making a file in your apps directory: `\texttt{touch test.txt}'   
	\end{enumerate}
	\begin{alertblock}{Result}
		\begin{center}Did you get any errors?  Let us know if you did\end{center}
	\end{alertblock}
	\btVFill

	\small Notes:
		\begin{itemize}\tiny
		\item On login you are placed in /home/$<$username$>$/
		\item By default, \ctilde{} is equivalent to /home/$<$username$>$/
		\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    \item User directories
		{\semitransp[25]{\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{User Directories}
\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Home}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ ls -l 
total 0
lrwxrwxrwx 1 user99 user99 23 Jan 15 20:38 \textcolor{teal}{apps} -> /lus/scratch/usr/user99
lrwxrwxrwx 1 user99 user99 19 Jan 15 20:38 \textcolor{teal}{lus} -> /lus/scratch/user99
lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item \ctilde{}/ is not on the {\lustre} filesystem and \textbf{should not be used for job data!}
		\item \ctilde{}/lus/ is a symlink to the {\lustre} scratch
		\begin{itemize}\tiny
			\item This is where all your job data files should live
			\item Items in this directory \emph{\textbf{are}} subject to our 90 day purge policy
		\end{itemize}
		\item \ctilde{}/apps/ is a symlink to where programs should be stored
		\begin{itemize}\tiny
			\item Items in this directory \emph{\textbf{are not}} subject to our 90 day purge policy
			\item Directory is monitored for abuse
		\end{itemize}
		\item \ctilde{}/purge/ is typically a dead symlink
		\begin{itemize}\tiny
			\item Symlink becomes active if the user has files that are part of the next automatic purge
			\item When \ctilde{}/purge/ is active, the directory containing two files -- \emph{purge\_list.txt} \& \emph{totals.txt}
			\item An email is sent to users if they have files that will be purged
			\item Email notification is sent out 14 days before the purge takes place
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Filesystems}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ df -h
Filesystem                                  Size  Used Avail Use% Mounted on
10.10.0.3:/ha_cluster/home                  1.8T  888G  851G  52% /home
10.12.0.51@o2ib:10.12.0.52@o2ib:/scratch    582T  429T  125T  78% /lus/scratch
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item /home/$<$username$>$ exists on a NFS mounted filesystem
		\begin{itemize}\tiny
			\item Only has 1.8TB of useable space
			\item Using all this space may cause problems for the entire cluster
			\item Not a high performance filesystem and small in size
		\end{itemize}
		\item /lus/scratch/ is the {\lustre} filesystem
		\begin{itemize}\tiny
			\item Has 582TB of useable space
			\item \ctilde{}/apps/, \ctilde{}/lus/, and \ctilde{}/purge/ all point to directories on this filesystem
			\item High performance and a lot more space for users to use
			\item No hard or soft quotas are in place
			\item Utilization is managed through the 90 day purge policy
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    {\semitransp[25]{\item User directories}}
		\item Transferring files
		\begin{itemize}
			\item Globus
		\end{itemize}
		{\semitransp[25]{
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Transferring Files}
\begin{frame}
	\frametitle{Available File Transfer Protocols}
	\begin{itemize}
		\item The cluster has the following options for transferring files:
		\begin{itemize}
			\item[--] scp (RCP$+$SSH protocol)
			\item[--] rsync (rsync protocol with SSH transport)
			\item[--] SFTP (SSH FTP protocol)
			\item[--] \href{https://www.globus.org/}{Globus} (Grid FTP protocol)
		\end{itemize}
		\item All options are widely used, and have clients that can be found for on most major operating systems
%               \item Please see the the CI website for links to several different clients?                 
	\end{itemize}
	\btVFill
	\begin{center}
		SFTP, scp, and rsync are fairly common on Linux systems,~\\but Globus is not as common~\ldots
	\end{center}

\end{frame}

\subsection{Globus}
\begin{frame}
	\frametitle{Globus}
	\begin{block}{What is Globus?}\footnotesize
	The Globus transfer service provides high-performance, secure, file transfer and synchronization between endpoints.~\\~\\
	Globus handles all the difficult aspects of data transfer, allowing application users to easily start and manage transfers between endpoints, while automatically tuning parameters to maximize bandwidth usage, managing security configurations, providing automatic fault recovery, and notifying users of completion and problems. 
	\end{block}
	\begin{definition}\tiny
	An \textbf{\emph{endpoint}} is one of the two file transfer locations -- either the source or the destination -- between which files can move.~\\Once a resource (such as a server, cluster, storage system, laptop, or other system) is defined as an endpoint, it will be available to authorized users who can transfer files to or from this endpoint.
	\end{definition}
\numlessfootnotetxt{\tiny \url{https://www.globus.org/file-transfer}}
\end{frame}

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the Cray CS300 are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the Cyberinfrastructure website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		}}
		\item Software
		\begin{itemize}
			\item Modules
			\item Acquiring software
			\item Compilers
		\end{itemize}
		{\semitransp[25]{
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Software}
\begin{frame}
	\frametitle{Modules}
	\begin{block}{Modules}\tiny
	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark

	\end{block}
	\begin{itemize}
		\item We globally install popular/frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\tiny
			\item `module avail' -- list installed modules
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module unload $<$module name$>$' -- Remove named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
		\begin{itemize}\footnotesize
			\item Two ways to install software on the cluster
			\begin{itemize}\tiny
				\item Download compatible binaries
				\item Compile software from source
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

\begin{frame}
	\frametitle{Acquiring Software -- Binaries and/or Source}
	\begin{itemize}
		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the Cray CS300
		\begin{itemize}
			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
		\end{itemize}
		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Compilers}
	\begin{itemize}
		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
		\item Compiling must take place on a compute node
		\begin{itemize}
			\item Interactive sessions are useful for compiling software
			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
		\end{itemize}
		\item {\intel} compilers are recommended for best performance
		\begin{itemize}
			\item {\intel} 2013 compilers:
			\begin{itemize}
				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
			\end{itemize}
			\item {\intel} 2016 compilers:
			\begin{itemize}
				\item We have 2 floating seats for {\intel} 2016 compiler
				\item intel\_2016/ics 
				\item intel\_2016/impi 
			\end{itemize}
		\end{itemize}			
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		}}
		\item Managing user jobs
		\begin{itemize}
			\item Job scheduler
			\item Using SLURM
			\item Partition layout
			\item Submitting jobs (Examples)
		\end{itemize}		
	\end{itemize}
\end{frame}

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads work together (OpenMP, pthreads, TBB)
	\item Require a lot of cores to process a lot of data in an identical manner, yet none of the inputs have dependencies on another (HTC)
\end{itemize}
The Cray CS300 is capable of handle many different types of jobs, but in a multi-user environment with so many users, how do we turn chaos into organized chaos?
~\\
~\\
This looks like a job for a \textbf{\emph{job scheduler}}!
\end{frame}

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing).
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize
    \item Interfaces which help to define workflows and/or job dependencies
    \item Automatic submission of executions
    \item Interfaces to monitor the executions
    \item Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the Cray CS300, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}.
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{Using SLURM}
\begin{frame}
\frametitle{How to use SLURM}
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}
	\item sbatch -- 
	\item srun --
	\item srun.x11 --
	\item sacct --
	\item scontrol --
	\item scancel --
\end{itemize}  
~\\
Let us look at how the nodes are partitioned, as well as some examples of submitting jobs and create job scripts
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\frametitle{Partition layout}
\end{frame}

\frametitle{SLURM Job Scripts}
\begin{frame}
\frametitle{Interactive Job with SLURM}
	\begin{block}{Interactive session (no X11)}\tiny
	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00 \ddash{}pty /bin/bash
	\end{block}

	\begin{block}{Interactive session (with X11)}\tiny
		\begin{enumerate}
			\item Connect via SSH using the -Y option, X11 forwarding enabled
			\item run srun.x11 to start a session on a node
		\end{enumerate}
		~\\
		~\\
		$[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00	~\\
		$[$compute-0001 \ctilde$]$\$ xterm	~\\
	\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

\#!/bin/sh
\#SBATCH \ddash{}job-name=MPI\_example
\#SBATCH \ddash{}partition=exclusive.q
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00 
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=4
\#SBATCH \ddash{}tasks-per-node=20
\#SBATCH \ddash{}cpus-per-task=1
\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes                                                                                                                                             
\#SBATCH \ddash{}error=hello-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#SBATCH \ddash{}output=hello-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
\#\# Intel MPI manual: \href{https://software.intel.com/en-us/mpi-refman-lin-html}{https://software.intel.com/en-us/mpi-refman-lin-html}
export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
export I\_MPI\_FABRICS=tmi  
export I\_MPI\_PMI\_LIBRARY=/opt/local/slurm/default/lib64/libpmi.so

cd \$SLURM\_SUBMIT\_DIR
srun  -n \$\{SLURM\_NTASKS\}  ./hello\_mpi.intel 
\end{semiverbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

\#!/bin/sh
\#SBATCH \ddash{}job-name=example
\#SBATCH \ddash{}partition=community.q
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=1
\#SBATCH \ddash{}tasks-per-node=1
\#SBATCH \ddash{}cpus-per-task=5
\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes                                                                                                                                              
\#SBATCH \ddash{}error=hello-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#SBATCH \ddash{}output=hello-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}

export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}

cd \$SLURM\_SUBMIT\_DIR
./hello\_world
\end{semiverbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm

\#!/bin/sh
\#SBATCH \ddash{}job-name=example
\#SBATCH \ddash{}partition=community.q
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00 
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=1
\#SBATCH \ddash{}tasks-per-node=1
\#SBATCH \ddash{}cpus-per-task=5
\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes                                                                                                                                               
\#SBATCH \ddash{}error=buzz-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#SBATCH \ddash{}output=buzz-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}

export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}

cd \$SLURM\_SUBMIT\_DIR
./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.flower -o output\_\$\{SLURM_ARRAY_TASK_ID\}.honey
\end{semiverbatim}
\end{frame}

\section{Cluster Etiquette}
\begin{frame}
\frametitle{Cluster Etiquette}
\end{frame}

\section{Policies}
\begin{frame}
	\frametitle{Overview}
	\begin{itemize}
		\item Login node usage policy
		\item Scratch filesystem purge policy
	\end{itemize}
\end{frame}

\subsection{Login Node Usage Policy}
\begin{frame}
\frametitle{Login Node Usage Policy}
\btVFill
\begin{center}
\footnotesize \textbf{\emph{Login node usage policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
\end{center}
\end{frame}

\subsection{Lustre Filesystem Purge Policies}
\begin{frame}
\frametitle{{\lustre} Filesystem Purge Policies}
Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.  To accomplish this, we have currently implemented a purge policy, for any file that is older than 90 days.
\begin{itemize}
 \item Which of my directories are subject to the purge policy?
   \begin{itemize}
   \item \textbf{Answer:} All files and folders found in \ctilde/lus/ are subject to the 90 day purge policy.  Symlinks are excluded from the purge, and are not followed by the purge bot.  
   \end{itemize}
\item How frequently are files for purging identified?
   \begin{itemize}
   \item \textbf{Answer:} Current frequency is every other month, but the frequency many increase or decrease based on the fill rate of the {\lustre} filesystem 
	\end{itemize}
\end{itemize}
\btVFill
\begin{center}
\footnotesize \textbf{\emph{Purge policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
\end{center}

\end{frame}

\section{Frequently Asked Questions}
\begin{frame}
\frametitle{FAQ}
\begin{itemize}
	\item Are files on the cluster backed up?
	\begin{itemize}\footnotesize
	\item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.  It is up to you as the user to validate and maintain your own backups.  The Cyberinfrastructure team takes no responsibility for any data that is lost due to human or mechanical error.  
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\Huge{\centerline{Questions?}}
\end{frame}

\end{document}

### 2016-01-18 matlab_dcs_101/paper.tex
added base template as itself, as well as imported the Matlab slide deck

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass[t]{beamer}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usetheme{ITS_CI}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{anyfontsize}
\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyphenat}
\usetikzlibrary{arrows,shapes}

\newlength{\okinalen}
\setlength{\okinalen}{\widthof{'}}
\newcommand{\okina}{\hbox to.666\okinalen{\hss`\hss}}

\newcommand{\CNAME}{the~Cray~CS300}
\newcommand{\TOTWRKS}{64}
\newcommand{\ctilde}{{\fontfamily{ptm}\selectfont\texttildelow}}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\newcommand{\tab}[1]{\hspace{.2\textwidth}\rlap{#1}}

\definecolor{links}{HTML}{2A1B81}
\hypersetup{colorlinks,linkcolor=,urlcolor=links}

\AtBeginSection[]
               {
                 \begin{frame}
                   \frametitle{Outline}
                   \tableofcontents[currentsection]
                 \end{frame}
               }

\AtBeginSubsection[]
                  {
                    \begin{frame}
                      \frametitle{Outline}
                      \tableofcontents[currentsection,currentsubsection]
                    \end{frame}
                  }

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[MATLAB on \CNAME]{MATLAB Distributed Compute on \CNAME } % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of Hawai\okina{}i -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
	Information Technology Services \\
	Department of Cyberinfrastructure \\
	University of Hawai\okina{}i \\ % Your institution for the title page
	\medskip
	\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
	\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
  \titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
  \frametitle{Overview} % Table of contents slide, comment this block out to remove it
  \tableofcontents[] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

\section{MATLAB Distributed Compute}

\begin{frame}
	\frametitle{MATLAB on the Cluster}
	\begin{itemize}
		\item MATLAB Distributed Compute Server (DCS) R2015b is installed on \CNAME
		\item We currently hold a license for \TOTWRKS{} \emph{Workers}.
			\begin{itemize}
				\item One worker is equal to an instance of MATLAB
			\end{itemize}
		\item In order to use MATLAB on \CNAME, you will need to own the MATLAB product with the Parallel Computing Toolbox
		\item The workers on \CNAME~contain all the MATLAB toolboxes
		\begin{itemize}
			\item \textbf{This does not mean you get free access to all toolboxes}
			\item \textbf{You will only have access to the MATLAB toolboxes you have purchased}
		\end{itemize}            
	\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Pre-requisites to using MATLAB on \CNAME{}}
  In order to use MATLAB on the cluster you must meet the follow requirements
  \begin{enumerate}
  \item Own a copy of MATLAB, with a license for the Parallel Computing Toolbox
  \item Have gone through the On-boarding course for \CNAME
  \end{enumerate}
\end{frame}

\section{Configuration}

\begin{frame}
  \frametitle{Required Files}
  In order to submit MATLAB jobs to \CNAME, a packet of several scripts, and a configuration file are required:
  \begin{itemize}
  \item Download the zip file containing the MATLAB setup files here: \url{http://go.hawaii.edu/YH}
  \item Download should contain:
    \begin{itemize}
    \item A directory named slurm\_matlab, which contains several .m and .sh files
    \item A settings file named uhhpc.settings
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Setting up MATLAB}
  \begin{itemize}
  \item Unzip setup files
  \item Move the \emph{slurm\_matlab} directory to a location it can reside permanently
    \begin{itemize}
    \item This directory will be added to your MATLAB path
    \end{itemize}
  \item Start up MATLAB
  \end{itemize}
  The next few Slides will cover what needs to be done in MATLAB to allow communication with \CNAME~from MATLAB
\end{frame}

\subsection{Setting up the MATLAB path}
\begin{frame}
  \frametitle{Adding to the MATLAB Path}
  \begin{picture}(320,250)
    \onslide<1->{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/set_path_button}}}
    \only<1>{\put (126,236.5){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.6,0) -- (0.6,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}

    \onslide<2-3>{\put (-8,92){\includegraphics[width=0.4\linewidth]{screenshots/Set_path_window}}}
    \only<2>{\put (-4.25,167.2){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (1.2,0) -- (1.2,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}

    \onslide<3>{\put (55,135){\includegraphics[width=0.35\linewidth]{screenshots/select_path}}}
    \only<3>{\put (78.5,151.7){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (2.2,0) -- (2.2,0.25) -- (0,0.25) -- (0,-0.01);\end{tikzpicture}}}
    \only<3>{\put (133.5,136.5){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.7,0) -- (0.7,0.22) -- (0,0.22) -- (0,-0.01);\end{tikzpicture}}}

    \onslide<4->{\put (-19,92){\includegraphics[width=0.4\linewidth]{screenshots/Save_addition}}}
    \only<4>{\put (22.3,166.5){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (3.2,0) -- (3.2,0.3) -- (0,0.3) -- (0,-0.01);\end{tikzpicture}}}
    \only<4>{\put (22.3,92.8){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (1.33,0) -- (1.33,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}

    \put(-10,82){
      \begin{minipage}[t]{0.95\linewidth}
        {
          \only<1>{\normalsize{Click the ``\textbf{Set Path}'' button under the ``\textbf{Home}'' tab in\\the MATLAB UI}}
          \only<2>{\normalsize{In the Set Path dialogue, click ``\textbf{Add Folder \ldots}'' }}
          \only<3>{\normalsize{In the Add Folder to Path dialogue, navigate to the directory\\the \textbf{slurm\_matlab} folder resides in and select and \textbf{Open} it}}
          \only<4>{\normalsize{The newly added path should appear in the ``\textbf{MATLAB search path:}'' list.  Once it added, go ahead and click ``\textbf{Save}'', then ``\textbf{Close}''}}
        }
      \end{minipage}
    }
  \end{picture}
\end{frame}

\subsection{Importing \CNAME~MATLAB profile}
\begin{frame}
  \frametitle{Cluster Profile Setup}
  \begin{picture}(320,250)
    \onslide<1,3->{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/set_path_button}}}
    \only<1>{\put (126,232){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.6,0) -- (0.6,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}
    \onslide<2>{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/manage_profile}}}
    \only<2>{\put (128,214){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (1,0) -- (1,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}
    \onslide<3-4>{\put(-8,92){\includegraphics[width=0.4\linewidth]{screenshots/CPM_1}}}
    \only<3>{\put (9.2,163.5){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.3,0) -- (0.3,0.46) -- (0,0.46) -- (0,-0.01);\end{tikzpicture}}}
    \onslide<4>{\put(55,135){\includegraphics[width=0.35\linewidth]{screenshots/CPM_2}}}
    \only<4>{\put (57,165){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (1.5,0) -- (1.5,0.2) -- (0,0.2) -- (0,-0.01);\end{tikzpicture}}}
    \only<4>{\put (133.5,136.5){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.7,0) -- (0.7,0.22) -- (0,0.22) -- (0,-0.01);\end{tikzpicture}}}
    \onslide<5,9>{\put(-8,92){\includegraphics[width=0.4\linewidth]{screenshots/CPM_3}}}
    \only<5>{\put (-8.2,149.65){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (1.15,0) -- (1.15,0.15) -- (0,0.15) -- (0,-0.01);\end{tikzpicture}}}
    \only<5>{\put (115.6,92.4){\begin{tikzpicture}\draw[red,thick,solid] (0,0) -- (0.4,0) -- (0.4,0.16) -- (0,0.16) -- (0,-0.01);\end{tikzpicture}}}
    \onslide<6>{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/CPM_4}}}
    \onslide<7>{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/CPM_4}}}
    \onslide<8>{\put(-8,92){\includegraphics[width=0.8\linewidth]{screenshots/CPM_4}}}
    \put(-20,82){
      \begin{minipage}[t]{0.95\linewidth}
        {
          \only<1>{\normalsize{Click the ``\textbf{Parallel}'' button to bring up the Parallel drop down menu}}
          \only<2>{\normalsize{Select \textbf{Manage Cluster Profiles} in the Parallel drop down menu }}
          \only<3>{\normalsize{In the \textbf{Cluster Profile Manager} dialogue box, click on\\the ``\textbf{Import}'' icon}}
          \only<4>{\normalsize{Navigate to and \textbf{Open} the uhhpc-matlab.settings file\\that was downloaded}}
          \only<5>{\normalsize{Select the \textbf{uhhpc} cluster profile, and press the ``\textbf{Edit}'' button.\\For ease of editing, you may want to increase the window size}}
          \only<6>{\footnotesize{Each of the functions in the \textbf{SUBMIT FUNCTIONS} section, have the follow format:}\\
            \textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}} } 
          }
          \only<7>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}}}\\Replace the remote storage dir with the path that matches your users lus folder location i.e., /home/$<$username$>$/lus/}}
          \only<8>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}}}\\Adjust the slurm partition and slurm timelimit to fit the partition you wish to run on, and finally click \textbf{done}}}
					\only<9>{\normalsize{The profile should now be configured for your user on \CNAME }}
        }
      \end{minipage}
    }
  \end{picture}
\end{frame}

\begin{frame}
  \frametitle{Cluster Profile}
	\begin{itemize}
	\item To simplify the use of MATLAB on the cluster, we suggest setting the UHHPC cluster profile as the default profile
	\begin{itemize}
		\item Many of the parallel toolbox commands will use the default cluster when an alternate cluster object is not provided
	\end{itemize}
	\item If the UHHPC cluster profile is not set as the default, the cluster profile name, \textbf{`uhhpc'}, must be provided when creating the cluster object
	\item If you plan to change partitions or run in multiple partitions on \CNAME, you will need to modify the existing profile, or create duplicate profiles with the required changes to the remote storage location, partition and timelimit
	\end{itemize}

\end{frame}

%------------------------------------------------
\section{Basic Commands \& Testing}

\subsection{Commands}
\begin{frame}
\frametitle{Commands}
\begin{itemize}
\item \href{http://www.mathworks.com/help/distcomp/pctconfig.html}{pctconfig} -- Configure settings for Parallel Computing Toolbox client session
  \begin{itemize}
  \item[-] Is not persistent
  \item[-] Must be set before any interaction with the parallel toolbox occurs
  \item[-] Used to set the correct 'hostname' in MATLAB\\e.g.,~$>$~pctconfig(`hostname', `192.168.0.1')
		\begin{itemize}
		\item[-] A valid `hostname' is required for validation to pass as well as for many of the parallel functions to work
		\end{itemize}
  \end{itemize}
\item \href{http://www.mathworks.com/help/distcomp/parcluster.html}{parcluster} -- Creates a cluster object which can be passed to functions used to create parallel tasks
\item \href{http://www.mathworks.com/help/distcomp/gcp.html}{gcp} -- Get current parallel pool
\item \href{http://www.mathworks.com/help/distcomp/addattachedfiles.html}{addAttachedFiles} -- Attach files or folders to parallel pool
\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Commands}
	\begin{itemize}
	\item \href{http://www.mathworks.com/help/distcomp/batch.html}{batch} -- Run MATLAB script or function on worker
		\begin{itemize}
		\item Non-interactive.  MATLAB can be closed
		\end{itemize}
	\item \href{http://www.mathworks.com/help/distcomp/parpool.html}{parpool} -- Create parallel pool on cluster
	\item \href{http://www.mathworks.com/help/distcomp/spmd.html}{spmd} -- Execute code in parallel on workers of parallel pool
	\end{itemize}
	For more information and other commands, visit the\\\href{http://www.mathworks.com/help/distcomp}{\textbf{Distributed toolbox help manual}} on the MathWorks website
\end{frame}

\subsection{Validation}
\begin{frame}
\frametitle{Simple Tests}
\begin{itemize}
\item Simplest test is to use ``\emph{validate}'' in the \textbf{Cluster Profile Manager}
\item Before testhing, make sure to reduce the value for ``Number of workers available to cluster''~(NumWorkers) value to something lower than \TOTWRKS{}
  \begin{itemize}
  \item[-] Typically we test with 2 -- 10 workers
  \end{itemize}
\item NumWorkers must be lowered since validate uses this as the number of workers to request
\item You also must make sure the hostname in MATLAB is set correctly (using~pctconf), or the parpool test may fail
\item Remember to reset the NumWorkers value back to \TOTWRKS{} or you may get errors saying not enough workers when requesting more than the current NumWorkers value
\end{itemize}\end{frame}

\section{Demo}
\begin{frame}
\frametitle{Demo}
\huge{\centerline{Live Demo \ldots}}
\end{frame}

\section{Resources}
\begin{frame}
\frametitle{Resources}
\begin{itemize}
\item Newest revision of these slides -- \url{http://go.hawaii.edu/H5}
\item \href{https://rc.fas.harvard.edu/resources/documentation/software/parallel-matlab-pct-dcs/}{Parallel MATLAB (PCT and DCS)} @ Harvard
\begin{itemize}
	\item Not all sections pertain to our cluster, but some sections provide some information on running jobs with the parallel toolbox
	\item The \href{https://rc.fas.harvard.edu/resources/documentation/software/parallel-matlab-pct-dcs/\#Submitting_DCS_jobs_from_within_MATLAB}{``Submitting DCS jobs from within MATLAB''} section covers using the batch command to submit parallel jobs to a cluster
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\huge{\centerline{Questions?}}
\end{frame}

%----------------------------------------------------------------------------------------

\end{document}

### 2016-01-19 hpc_101/paper.tex
Filled in section about slurm partitions.  Also created a partition layout image.

@@
 
+\newcommand{\ci}{Cyberinfrastructure}
+
 \AtBeginSection[]
@@
 Information Technology Services \\
-Department of Cyberinfrastructure \\
+Department of {\ci} \\
 University of {\hawaii} \\ % Your institution for the title page
@@
 			\item Using SLURM
-			\item Partition layout			
+			\item Partitions			
 			\item Submitting jobs (Examples)
@@
 ~\\ 
-For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the Cyberinfrastructure website for more information:~\\
+For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
 {\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
@@
 			\item Using SLURM
-			\item Partition layout
+			\item Partitions
 			\item Submitting jobs (Examples)
@@
 
-\subsection{Using SLURM}
+\subsection{SLURM}
+\begin{frame}
+\frametitle{SLURM}
+\begin{block}{How are jobs scheduled?}
+User submitted jobs are assigned a priority using a fairshare algorithm.
+Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job.
+\end{block}
+\end{frame}
+
 \begin{frame}
@@
 \begin{frame}
-\frametitle{Partition layout}
+\footnotesize
+\frametitle{Partitions}
+\begin{block}{What is a partition?}\footnotesize
+A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.
+Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc.
+Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
+\end{block}
+\begin{itemize}
+	\item The Cray CS300 currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
+	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
+	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
+	\begin{itemize}\footnotesize
+		\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
+		\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
+	\end{itemize}
+
+\end{itemize}
+\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
+\end{frame}
+
+\begin{frame}
+\frametitle{Partitions}
+\includegraphics[width=0.95\textwidth]{images/partitions}
 \end{frame}
@@
 It is up to you as the user to validate and maintain your own backups.
-The Cyberinfrastructure team takes no responsibility for any data that is lost due to human or mechanical error.
+The {\ci} team takes no responsibility for any data that is lost due to human or mechanical error.
 	\end{itemize}

### 2016-01-19 hpc_101/paper.tex
adjustment of the partition image. added section describing the different slurm commands.

@@
 
+\newcommand{\craycs}{Cray~CS300}
+
 \newcommand{\ci}{Cyberinfrastructure}
@@
 
-\subsection{Cray CS300}
-\begin{frame}
-\frametitle{Cray CS300 -- History}
+\subsection{{\craycs}}
+\begin{frame}
+\frametitle{{\craycs} -- History}
 \begin{itemize}
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Compute Nodes}
+	\frametitle{{\craycs}0 -- Compute Nodes}
 	\begin{itemize}
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Storage}
-	Two storage options are currently available on the Cray CS300.
+	\frametitle{{\craycs} -- Storage}
+	Two storage options are currently available on the {\craycs}
 	\begin{enumerate}
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Storage $\rightarrow$ {\lustre}}
+	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
 	\begin{itemize}
 		\item {\lustre} is a high performance parallel filesystem
-		\item The Cray CS300 has $\approx582$TB of storage space
+		\item The {\craycs} has $\approx582$TB of storage space
 		\item Shared between all compute nodes and login nodes
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Storage $\rightarrow$ ValueStorage}
+	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
 	\begin{itemize}
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Network}
+	\frametitle{{\craycs} -- Network}
 	\begin{itemize}
@@
 \begin{frame}
-	\frametitle{Cray CS300 -- Layout}
+	\frametitle{{\craycs} -- Layout}
 	\includegraphics[width=0.95\textwidth]{images/high_level}
@@
 	\item[] \begin{itemize}
-		\item The Cray CS300 has two login nodes:
+		\item The {\craycs} has two login nodes:
 		\begin{itemize}
@@
 
-\subsubsection{Login to the Cray CS300}
-\begin{frame}
-\frametitle{Login to the Cray CS300}
+\subsubsection{Login to the {\craycs}}
+\begin{frame}
+\frametitle{Login to the {\craycs}}
 	\begin{center}\textbf{Windows}\end{center}
@@
 \begin{frame}
-\frametitle{Login to the Cray CS300}
+\frametitle{Login to the {\craycs}}
 	\begin{center}\textbf{Mac \& Linux}\end{center}
@@
 	\frametitle{Globus}\footnotesize
-If you already use Globus, the endpoints for the Cray CS300 are:
+If you already use Globus, the endpoints for the {\craycs} are:
 \begin{itemize}
@@
 	\begin{itemize}
-		\item We globally install popular/frequently requested software packages and create modules for all users to access
+		\item We globally install frequently requested software packages and create modules for all users to access
 		\item Access to modules is via the \textbf{module} command
-		\begin{itemize}\tiny
+		\begin{itemize}\footnotesize
 			\item `module avail' -- list installed modules
@@
 		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
-		\begin{itemize}\footnotesize
-			\item Two ways to install software on the cluster
-			\begin{itemize}\tiny
-				\item Download compatible binaries
-				\item Compile software from source
-			\end{itemize}
-		\end{itemize}
+		%\begin{itemize}\footnotesize
+			%\item Two ways to install software on the cluster
+			%\begin{itemize}\tiny
+				%\item Download compatible binaries
+				%\item Compile software from source
+			%\end{itemize}
+		%\end{itemize}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the Cray CS300
+		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the {\craycs}
 		\begin{itemize}
@@
 User jobs all come in different shapes and sizes:  
-\begin{itemize}
+\begin{itemize}\footnotesize
 	\item Require multiple nodes working in concert towards a common goal (MPI)
-	\item Require a single node, in which they use multiple threads work together (OpenMP, pthreads, TBB)
-	\item Require a lot of cores to process a lot of data in an identical manner, yet none of the inputs have dependencies on another (HTC)
-\end{itemize}
-The Cray CS300 is capable of handle many different types of jobs, but in a multi-user environment with so many users, how do we turn chaos into organized chaos?
-~\\
-~\\
-This looks like a job for a \textbf{\emph{job scheduler}}!
+	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
+	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
+\end{itemize}
+\bigskip
+The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we turn this chaos into organized chaos?
+\btVFill
+\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
 \end{frame}
@@
 	\begin{definition}{\footnotesize
-		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing).
+		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
 		}
@@
 	\end{itemize}
-	For the Cray CS300, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}.
+	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
 	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
@@
 User submitted jobs are assigned a priority using a fairshare algorithm.
-Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job.
+Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job
 \end{block}
-\end{frame}
-
-\begin{frame}
-\frametitle{How to use SLURM}
+%The following slides that cover slurm will cover
+%\begin{enumerate}
+%\item Several useful SLURM commands
+%\item Resource partitions
+%\item example interactive jobs and batch job scripts
+%\end{enumerate}
+\end{frame}
+
+\begin{frame}
+\frametitle{SLURM commands}\footnotesize
 SLURM has a series of commands, each of which allow users to interact with the job scheduler
-\begin{itemize}
-	\item sbatch -- 
-	\item srun --
-	\item srun.x11 --
-	\item sacct --
-	\item scontrol --
-	\item scancel --
-\end{itemize}  
-~\\
-Let us look at how the nodes are partitioned, as well as some examples of submitting jobs and create job scripts
+\begin{itemize}\tiny
+	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
+	\item \emph{\textbf{srun.x11}} -- Used to submit an interactive job with x11 support
+	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
+	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
+	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
+	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
+	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
+	\item \emph{\textbf{scontrol}} -- The administrative tool used to view and/or modify Slurm state. Note that many scontrol commands can only be executed as user root
+\end{itemize}
+\begin{itemize}
+	\item[--] Examples usage of the SLURM commands can be seen on schedmd's \href{http://slurm.schedmd.com/quickstart.html}{quickstart}
+	\item[--] Each command should have a `man' page, or displays help when the -h flag is used
+\end{itemize}
+\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
 \end{frame}
@@
 \begin{itemize}
-	\item The Cray CS300 currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
+	\item The {\craycs} currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
 	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
@@
 \begin{frame}[fragile]
-\frametitle{SLURM Submission Script File (MPI Job)}
+\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
 \begin{semiverbatim}\tiny
@@
 \begin{frame}[fragile]
-\frametitle{SLURM Submission Script File (Non-MPI Job)}
+\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
 \begin{semiverbatim}\tiny
@@
 \begin{frame}[fragile]
-\frametitle{SLURM Submission Script File (Job Array)}
+\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
 \begin{semiverbatim}\tiny

### 2016-01-20 hpc_101/paper.tex
addition of globus walk through, and partial table of partition details

@@
 Information Technology Services \\
-Department of {\ci} \\
+{\ci} \\
 University of {\hawaii} \\ % Your institution for the title page
@@
 \begin{itemize}
-		\item Brought online Fall 2014 -- Spring 2015
  \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
+ \item Deilvered in October of 2014
+ \item Accepted in December of 2014 
+ \item Early adopter testing started in January 2015
+ \item Opened for general use in April 2015
  \item As of October 2015, more than 200 users have been granted access to the cluster
@@
 				\item Diskless -- Some RAM is used for the Operating System
-				\item $\approx128$GB of useable RAM
+				\item 128GB of physical RAM
+				\item $\approx110$GB of useable RAM
 			\end{itemize}
@@
 				\item Diskless -- Some RAM is used for the Operating System
-				\item $\approx1$TB of useable RAM
+				\item 1TB of physical RAM
+				\item $\approx1008$GB of useable RAM
 			\end{itemize}			
@@
 		\item \textbf{Data is not backed up!  Users are responsible for their own data}
+                \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
 	\end{itemize}
@@
 		\item ValueStorage will eventually be available for purchase by everyone
-		\item ValueStorage owners will eventually be able to mount as a network drive (CIFS) on laptop, workstations, servers
+		\item ValueStorage owners will potentially be able to mount as a network drive (CIFS) on laptop, workstations, servers
 		\item Purchased in 0.5TB increments
@@
 			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
+                        \item low latency ($\approx1.3\mu$s)
 			\item Utilizes the \emph{fat tree network topology}
@@
 	\begin{itemize}
-		\item The initial investment in the cluster provides resources for all Faculty, Staff, and Students affiliate with the University of {\hawaii}
+		\item The initial investment in the cluster provides resources for all faculty, staff, and students affiliated with the University of {\hawaii}
 		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, lm.q, sb.q, kill.q, htc.q}
@@
 	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
-	 \item SUs come in two variates:
+	 \item SUs come in two flavors:
 		 \begin{itemize}
@@
 	\frametitle{Globus}\footnotesize
-If you already use Globus, the endpoints for the {\craycs} are:
-\begin{itemize}
-\item hawaii\#UHHPC1
-\item hawaii\#UHHPC1
-\end{itemize}
+\begin{block}{How do I get Globus?}
+To utilize Globus, follow the following steps:
+\begin{itemize}\tiny
+\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
+\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
+\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
+\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
+\begin{itemize}\tiny
+ \item Each of the login nodes is also an endpoint: \textbf{\emph{hawaii\#UHHPC1}} \& \textbf{\emph{hawaii\#UHHPC1}}
+\end{itemize}
+\item Once you select an endpoint, a login window will pop up.
+You can access the UHHPC endpoints by simply using your UH username and password.
+Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’.
+\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
+\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
+\end{itemize}
+\end{block}
 ~\\
-For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
+These instructions can also be found on the {\ci} website:~\\
 {\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Globus}
+        \begin{itemize}
+          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
+          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
+          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
+            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
+\end{itemize} 
 \end{frame}
@@
 \bigskip
-The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we turn this chaos into organized chaos?
+The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we impose order on this chaos?
 \btVFill
@@
 	\begin{itemize}\footnotesize
-    \item Interfaces which help to define workflows and/or job dependencies
-    \item Automatic submission of executions
-    \item Interfaces to monitor the executions
-    \item Priorities and/or queues to control the execution order of unrelated jobs
+
+        \item The ability for the user to interact with the schduler to define workflows and/or define job dependencies
+        \item[--] Interfaces which help to define workflows and/or job dependencies
+
+        \item The ability to automatically manage the scheduling of user jobs
+        \item[--] Automatic submission of executions
+
+        \item The functionality to allow users to monitor submitted jobs
+        \item[--] Interfaces to monitor the executions
+
+        \item The capability to control and prioritize the execution order of unrelated jobs
+        \item[--] Priorities and/or queues to control the execution order of unrelated jobs
 	\end{itemize}
@@
 	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
-	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
+\end{itemize}
+\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
+\end{frame}
+
+\begin{frame}
+\frametitle{Partitions}
+\includegraphics[width=0.95\textwidth]{images/partitions}
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Partition Details}\tiny
+\begin{tabular}{c c c c c c c}
+  \toprule                                                                    
+\textbf{Partition} & \textbf{Time} & \textbf{Nodes} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt\\Mode} & \textbf{Mem per CPU} \\
+\midrule 
+community.q & \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} & \begin{tabular}{c c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
+exclusive.q & \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
+\bottomrule 
+\end{tabular}
+
+%% \begin{columns}[t]
+%%   \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=community.q
+%%    DefaultTime=00:10:00
+%%    MaxTime=3-00:00:00 
+%%    MinNodes=1
+%%    MaxNodes=20 
+%%    Nodes=prod2-[0003-0085]
+%%    Priority=10
+%%    Shared=NO 
+%%    PreemptMode=OFF
+%%    DefMemPerCPU=3250 
+%%    MaxMemPerCPU=6500
+%% \end{semiverbatim}
+%%   \end{column}
+
+%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=exclusive.q
+%%    DefaultTime=00:10:00 
+%%    MaxTime=3-00:00:00 
+%%    MinNodes=1 
+%%    MaxNodes=20 
+%%    Nodes=prod2-[0003-0085]
+%%    Priority=10
+%%    Shared=EXCLUSIVE 
+%%    PreemptMode=OFF
+%%    DefMemPerNode=UNLIMITED 
+%%    MaxMemPerCPU=6500
+%% \end{semiverbatim}
+%% \end{column}
+
+%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=lm.q
+%%    DefaultTime=00:10:00
+%%    MaxTime=3-00:00:00 
+%%    MinNodes=1 
+%%    MaxNodes=1 
+%%    Nodes=lm2-[0001-0006]
+%%    Priority=10
+%%    Shared=NO
+%%    PreemptMode=REQUEUE
+%%    DefMemPerNode=UNLIMITED 
+%%    MaxMemPerCPU=262100
+%% \end{semiverbatim}
+%% \end{column} 
+%% \end{columns}
+
+%% \begin{columns}[t]
+%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=kill.q
+%%    DefaultTime=00:10:00
+%%    MaxTime=3-00:00:00
+%%    MinNodes=1
+%%    MaxNodes=12
+%%    Nodes=prod2-[0085-0176]
+%%    Priority=10
+%%    Shared=NO
+%%    PreemptMode=REQUEUE
+%%    DefMemPerCPU=3250
+%%    MaxMemPerCPU=6500
+
+%% \end{semiverbatim}
+%% \end{column}
+
+%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=htc.q
+%%    DefaultTime=00:10:00
+%%    MaxTime=3-00:00:00 
+%%    MinNodes=1 
+%%    MaxNodes=1 
+%%    Nodes=prod2-[0001-0176]
+%%    Priority=1
+%%    Shared=NO
+%%    PreemptMode=REQUEUE
+%%    DefMemPerCPU=3250 
+%%    MaxMemPerCPU=6500
+%% \end{semiverbatim}
+%% \end{column}
+
+%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
+%% PartitionName=sb.q
+%%    DefaultTime=00:05:00
+%%    MaxTime=01:00:00
+%%    MinNodes=1
+%%    MaxNodes=2
+%%    Nodes=prod2-[0177-0178]
+%%    Priority=10 
+%%    Shared=NO 
+%%    PreemptMode=REQUEUE
+%%    DefMemPerCPU=3250 
+%%    MaxMemPerCPU=6500
+%% \end{semiverbatim}
+%% \end{column} 
+%% \end{columns}
+
+More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
 \begin{itemize}\footnotesize
@@
 
-\end{itemize}
-\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
-\end{frame}
-
-\begin{frame}
-\frametitle{Partitions}
-\includegraphics[width=0.95\textwidth]{images/partitions}
 \end{frame}

### 2016-01-20 hpc_101/paper.tex
completiong of partition detail table

@@
 %----------------------------------------------------------------------------------------
+
+%https://tex.stackexchange.com/questions/74353/what-commands-are-there-for-horizontal-spacing
+%https://tex.stackexchange.com/questions/80212/controling-padding-using-the-xstring-package
+%https://tex.stackexchange.com/questions/2441/how-to-add-a-forced-line-break-inside-a-table-cell
+%https://en.wikibooks.org/wiki/LaTeX/Tables#The_tabularx_package
+%https://tex.stackexchange.com/questions/312/correctly-typesetting-a-tilde
+%https://tex.stackexchange.com/questions/34580/escape-character-in-latex
+%https://tex.stackexchange.com/questions/54180/how-do-i-write-something-at-the-end-of-the-slide-in-beamer
+%http://code.hammerpig.com/add-footnote-number-latex.html
+%https://tex.stackexchange.com/questions/51452/reduce-spacing-in-table-of-contents-beamer
+%https://www.sharelatex.com/learn/Using_colours_in_LaTeX
+%http://www.math-linux.com/latex-26/article/how-to-make-a-presentation-with-latex-introduction-to-beamer
+%https://stackoverflow.com/questions/3594329/semi-transparent-text-in-beamer-pdflatex
+%https://www.sharelatex.com/blog/2013/08/16/beamer-series-pt3.html
+
 %\documentclass[t]{beamer}
@@
 \usepackage{calc}
-
+\usepackage{makecell}
+\usepackage{xstring}
 \mode<presentation> {
@@
 \newcommand{\ci}{Cyberinfrastructure}
+
+\newcommand\padfrom[2]{\StrLen{#2}[\templen]\StrGobbleLeft{#1}{\templen}#2}
 
@@
 
-\begin{frame}[fragile]
-\frametitle{Partition Details}\tiny
-\begin{tabular}{c c c c c c c}
+\begin{frame}
+\frametitle{Partition Details}
+\resizebox{0.95\textwidth}{!}{%
+\begin{tabular}{l || c || c || c || c || c || c}
 \toprule                                                                    
-\textbf{Partition} & \textbf{Time} & \textbf{Nodes} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt\\Mode} & \textbf{Mem per CPU} \\
-\midrule 
-community.q & \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} & \begin{tabular}{c c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
-exclusive.q & \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
+\textbf{Partition} & \textbf{Time} & \textbf{Nodes per job} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt Mode} & \textbf{Memory per CPU (MB)} \\
+\toprule                                                                    
+\toprule                                                                    
+%\midrule
+%\midrule
+community.q & \begin{tabular}{l r} \textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+exclusive.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min}& \padfrom{\ \ \ }{1}~\\ \textbf{Max}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+kill.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+htc.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+lm.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
+\hline
+sb.q & \begin{tabular}{l r}\textbf{Default} & 0-00:05:00~\\ \textbf{Max} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \bottomrule 
 \end{tabular}   
-
-%% \begin{columns}[t]
-%%   \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=community.q
-%%    DefaultTime=00:10:00
-%%    MaxTime=3-00:00:00 
-%%    MinNodes=1
-%%    MaxNodes=20 
-%%    Nodes=prod2-[0003-0085]
-%%    Priority=10
-%%    Shared=NO 
-%%    PreemptMode=OFF
-%%    DefMemPerCPU=3250 
-%%    MaxMemPerCPU=6500
-%% \end{semiverbatim}
-%%   \end{column}
-
-%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=exclusive.q
-%%    DefaultTime=00:10:00 
-%%    MaxTime=3-00:00:00 
-%%    MinNodes=1 
-%%    MaxNodes=20 
-%%    Nodes=prod2-[0003-0085]
-%%    Priority=10
-%%    Shared=EXCLUSIVE 
-%%    PreemptMode=OFF
-%%    DefMemPerNode=UNLIMITED 
-%%    MaxMemPerCPU=6500
-%% \end{semiverbatim}
-%% \end{column}
-
-%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=lm.q
-%%    DefaultTime=00:10:00
-%%    MaxTime=3-00:00:00 
-%%    MinNodes=1 
-%%    MaxNodes=1 
-%%    Nodes=lm2-[0001-0006]
-%%    Priority=10
-%%    Shared=NO
-%%    PreemptMode=REQUEUE
-%%    DefMemPerNode=UNLIMITED 
-%%    MaxMemPerCPU=262100
-%% \end{semiverbatim}
-%% \end{column} 
-%% \end{columns}
-
-%% \begin{columns}[t]
-%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=kill.q
-%%    DefaultTime=00:10:00
-%%    MaxTime=3-00:00:00
-%%    MinNodes=1
-%%    MaxNodes=12
-%%    Nodes=prod2-[0085-0176]
-%%    Priority=10
-%%    Shared=NO
-%%    PreemptMode=REQUEUE
-%%    DefMemPerCPU=3250
-%%    MaxMemPerCPU=6500
-
-%% \end{semiverbatim}
-%% \end{column}
-
-%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=htc.q
-%%    DefaultTime=00:10:00
-%%    MaxTime=3-00:00:00 
-%%    MinNodes=1 
-%%    MaxNodes=1 
-%%    Nodes=prod2-[0001-0176]
-%%    Priority=1
-%%    Shared=NO
-%%    PreemptMode=REQUEUE
-%%    DefMemPerCPU=3250 
-%%    MaxMemPerCPU=6500
-%% \end{semiverbatim}
-%% \end{column}
-
-%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
-%% PartitionName=sb.q
-%%    DefaultTime=00:05:00
-%%    MaxTime=01:00:00
-%%    MinNodes=1
-%%    MaxNodes=2
-%%    Nodes=prod2-[0177-0178]
-%%    Priority=10 
-%%    Shared=NO 
-%%    PreemptMode=REQUEUE
-%%    DefMemPerCPU=3250 
-%%    MaxMemPerCPU=6500
-%% \end{semiverbatim}
-%% \end{column} 
-%% \end{columns}
-
-More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
-\begin{itemize}\footnotesize
-\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
-\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
-\end{itemize}
-
+}
+\\
+\bigskip
+Partition details also available on the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command:
+\\
+\texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
 \end{frame}

### 2016-01-20 hpc_101/paper.tex
change table line

@@
 \textbf{Partition} & \textbf{Time} & \textbf{Nodes per job} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt Mode} & \textbf{Memory per CPU (MB)} \\
-\toprule                                                                    
-\toprule                                                                    
-%\midrule
-%\midrule
+%\toprule                                                                    
+%\toprule                                                                    
+\midrule
+\midrule
 community.q & \begin{tabular}{l r} \textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\

### 2016-01-20 hpc_101/paper.tex
all slides in hpc_101 now have text.  Chnaged the main slide on the matlab ones to state the correct group

@@
 \newcommand{\ci}{Cyberinfrastructure}
+\newcommand{\citeam}{CI--Team}
 
@@
 
-\title[On-boarding]{HPC On-boarding} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[HPC--101]{HPC--101\\Onboarding} % The short title appears at the bottom of every slide, the full title is only on the title page
 
@@
 \frametitle{Cluster Etiquette}
+\begin{itemize}
+\item Never run anything on the login nodes
+\item Try to request only the resources that your job needs.  Leaving unneeded resources available, allows other users to use them
+\item Test what your application in an interactive session before scheduling a long running job
+\begin{itemize}
+\item[--] This can help identify what resources your job will require, and potentially aid you in correctly sizing what resources you request.
+\end{itemize}
+\item Use the nodes in the sb.q to validate your slurm script runs as intended
+\begin{itemize}
+\item[--] This helps to reduce the chance of you having a small error that kills your job prematurely after waiting for resources 
+\end{itemize} 
+\item Always run jobs from within {\ctilde}/lus/ directory to ensure output is written there and does not fill up the system filesystem
+\end{itemize}
 \end{frame}
@@
 \section{Policies}
-\begin{frame}
-	\frametitle{Overview}
-	\begin{itemize}
-		\item Login node usage policy
-		\item Scratch filesystem purge policy
-	\end{itemize}
-\end{frame}
+%% \begin{frame}
+%% 	\frametitle{Overview}
+%% 	\begin{itemize}
+%% 		\item Login node usage policy
+%% 		\item Scratch filesystem purge policy
+%% 	\end{itemize}
+%% \end{frame}
 
@@
 \begin{frame}
-\frametitle{Login Node Usage Policy}
+\frametitle{Login Node Usage Policy}\footnotesize
+The login nodes are a shared resource and are the only access to the cluster for hundreds of user and are meant to provide the following functionality for all users: 
+\begin{itemize}
+\item Providing ssh shell access 
+\item Facilitate the transfer files to and from the cluster:\\Globus, sftp, scp, rsync
+\item Launching and monitoring SLURM jobs (batch and interactive)
+\item Modifying text files with a text editor: vi/vim, emacs, nano
+\end{itemize}
+\bigskip
+\begin{itemize}
+\item[--] If we identify or are notified that a user is running computation on the login nodes, the application will be killed
+\item[--] If we determine that a user repeatly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
+\end{itemize}
 \btVFill
@@
 \frametitle{FAQ}
-\begin{itemize}
+\begin{itemize}\footnotesize
 \item Are files on the cluster backed up?
-	\begin{itemize}\footnotesize
+  \begin{itemize}\tiny
   \item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.
 It is up to you as the user to validate and maintain your own backups.
-The {\ci} team takes no responsibility for any data that is lost due to human or mechanical error.
-	\end{itemize}
-\end{itemize}
+The {\citeam} takes no responsibility for any data that is lost due to human or mechanical error.
+  \end{itemize}
+
+\item Help I have been trying to compile an application without success, can the {\citeam} help?
+  \begin{itemize}\tiny
+  \item \textbf{Answer:} We are willing to help, just send us an email at \textit{uh-hpc-help@lists.hawaii.edu} with a description of what you have attempted as well a link to the software you are trying to compile, and/or where it is located on the cluster.
+  \end{itemize}
+
+\item HELP! My job needs more time and does not support checkpointing.  What can I do?
+  \begin{itemize}\tiny
+  \item \textbf{Answer:} We understand that jobs may not all fit within the 3 day timelimit or unforeseen issues during the execution can cause a job to run long.
+The {\citeam} does review and potentially approves requests to extend jobs in a one off fashion.
+A request to extend a job can be denied, since we need to take into account factor such as: how busy the cluster is, the number of time extension requests a user has made in recent history.\\If it is known before the job is submitted that more than 3 days is required, we would ask that you contact us first so we can verify that you truly will need more time.\\If a job is already running and just doesn't look like it will complete in time, please send us a request with the jobid number and what you want the job extended to e.g., please extend my jobs run time to 7 days.\\Please remember the {\citeam} is small and we are also human.
+Depending on when you make your request we may not be able to immediately respond or act.
+It is best to make your request at the first signs you need more time, in hopes that you can provide us with ample time to act.
+  \end{itemize}
+
+\end{itemize}
+
 \end{frame}

### 2016-01-20 matlab_dcs_101/paper.tex
all slides in hpc_101 now have text.  Chnaged the main slide on the matlab ones to state the correct group

@@
 	Information Technology Services \\
-	Department of Cyberinfrastructure \\
+	Cyberinfrastructure \\
 	University of Hawai\okina{}i \\ % Your institution for the title page

### 2016-01-21 hpc_101/paper.tex
Added slide simulating submission of example script removed some commented text, and re worked the Job scheduler definition.

@@
 			\item `module avail' -- list installed modules
+			\item `module show $<$module name$>$' -- Show what actions a module performs
 			\item `module load $<$module name$>$' -- Loads the named module
-			\item `module unload $<$module name$>$' -- Remove named module
 			\item `module purge' -- Unload all loaded modules
@@
 		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
-		%\begin{itemize}\footnotesize
-			%\item Two ways to install software on the cluster
-			%\begin{itemize}\tiny
-				%\item Download compatible binaries
-				%\item Compile software from source
-			%\end{itemize}
-		%\end{itemize}
 	\end{itemize}
@@
 	\frametitle{What is a Job Scheduler?}
-	\begin{definition}{\footnotesize
-		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
-		}
-	\end{definition}
 	Basic features expected of a job scheduler include:
 	\begin{itemize}\footnotesize
-
-        \item The ability for the user to interact with the schduler to define workflows and/or define job dependencies
-        \item[--] Interfaces which help to define workflows and/or job dependencies
-
-        \item The ability to automatically manage the scheduling of user jobs
-        \item[--] Automatic submission of executions
-
-        \item The functionality to allow users to monitor submitted jobs
-        \item[--] Interfaces to monitor the executions
-
         \item The capability to control and prioritize the execution order of unrelated jobs
-        \item[--] Priorities and/or queues to control the execution order of unrelated jobs
-	\end{itemize}
+        \item Allocates exclusive and/or non-exclusive access to resources to users for some duration of time so they can perform work
+				\item Provides a framework for starting, executing, and monitoring work on the set of allocated nodes
+				\item Arbitrate contention for resources by managing a queue of pending work
+	\end{itemize}
+	\bigskip
 	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
@@
 \end{block}
-%The following slides that cover slurm will cover
-%\begin{enumerate}
-%\item Several useful SLURM commands
-%\item Resource partitions
-%\item example interactive jobs and batch job scripts
-%\end{enumerate}
 \end{frame}
@@
 
+\subsection{Interactive Job -- srun}
 \frametitle{SLURM Job Scripts}
@@
 
+\subsection{MPI Batch Job Example Script}
 \begin{frame}[fragile]
@@
 
+\subsection{Non-MPI Batch Job Example Script}
 \begin{frame}[fragile]
@@
 
+\subsection{Batch Job Example Script Using Job Array}
 \begin{frame}[fragile]
@@
 
+\subsection{Executing and Monitoring Batch Jobs}
+\begin{frame}[fragile]
+\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
+To execute a submission script you use the \textbf{\texttt{sbatch}} command
+\begin{block}{Example -- Non Job Array}
+\begin{semiverbatim}\tiny
+[login lus]\$ sinfo
+PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
+community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
+[login lus]\$ sbatch hello_world.slurm
+sbatch: Submitted batch job 469
+[login lus]\$ squeue
+JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
+469   community.q  example  user99  R  00:01 1      compute-0001
+\end{semiverbatim}
+\end{block}
+\begin{block}{Example -- Job Array}
+\begin{semiverbatim}\tiny
+[login lus]\$ sinfo
+PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
+community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
+[login lus]\$ sbatch {\ddash}array=0-2 job_array.slurm
+sbatch: Submitted batch job 470
+[login lus]\$ squeue
+JOBID     PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
+470[0-2]  community.q  example  user99  PD   00:00   1      (resource)
+470_0     community.q  example  user99  R    00:01   1      compute-0001
+470_1     community.q  example  user99  R    00:05   1      compute-0002
+\end{semiverbatim}
+\end{block}
+
+\end{frame}
+
 \section{Cluster Etiquette}

### 2016-01-21 hpc_101/paper.tex
Updated the layout image.  Changed the description of how priorities are assigned, so that it just lists potential factors

@@
 	\frametitle{{\craycs} -- Layout}
-	\includegraphics[width=0.95\textwidth]{images/high_level}
+	\includegraphics[width=1\textwidth]{images/layout}
 \end{frame}
@@
         \item The capability to control and prioritize the execution order of unrelated jobs
-        \item Allocates exclusive and/or non-exclusive access to resources to users for some duration of time so they can perform work
-				\item Provides a framework for starting, executing, and monitoring work on the set of allocated nodes
+        \item Allocate exclusive and/or non-exclusive access to resources to users for some duration of time so they can perform work
+	\item Provides a framework for starting, executing, and monitoring work on the set of allocated resources
 	\item Arbitrate contention for resources by managing a queue of pending work
@@
 	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
-	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
 	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
@@
 \begin{block}{How are jobs scheduled?}
-User submitted jobs are assigned a priority using a fairshare algorithm.
-Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job
+User submitted jobs are assigned a priority using a fairshare algorithm.\\
+Factors such as the following are all used to assign a priority to a given job: 
+\begin{itemize}
+\item Runtime
+\item Resource usage request
+\item Age of job
+\item Amount of core hours a user has used in recent history
+\end{itemize}
+
 \end{block}

### 2016-01-22 hpc_101/cheatsheet/paper.tex
added a cheatsheet

\documentclass[11pt,twocolumn]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{etoolbox}
\usepackage{lmodern}% http://ctan.org/pkg/lm                                                                                                                                                                                                                                                
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{anyfontsize}
\usepackage{graphicx} % Allows including images                                                                                                                                                                                                                                             
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables                                                                                                                                                                                                      
\usepackage{calc}
\usepackage{xstring}
\usepackage{titling}

\newcommand{\lft}{$<$}
\newcommand{\rht}{$>$}

\newcommand{\pth}[1]{{\lft}#1{\rht}}

\newlength{\okinalen}
\setlength{\okinalen}{\widthof{'}}
\newcommand{\okina}{\hbox to.666\okinalen{\hss`\hss}}
\newcommand{\ctilde}{{\fontfamily{ptm}\selectfont\texttildelow}}
\newcommand{\ddash}{-{}-}
\newcommand{\hawaii}{Hawai{\okina}i}
\newcommand{\ci}{Cyberinfrastructure}

\newcommand{\desc}[1]{\small \begin{itemize}\item[]#1\end{itemize}}

\begin{document}
\setlength{\droptitle}{-10em}
\title{HPC--101 -- Cheatsheet \vspace{-2ex}}
\date{\today}
\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\University of {\hawaii}\\Information Technology Services -- {\ci}}
\maketitle
\section{Directory Navigation}
\begin{itemize}
\item[] cd {\pth{Directory}}\desc{Change to directory specified}
\item[] ls {\pth{Path}}\desc{List the contents of the specified path}
\item[] ls -l {\pth{Path}}\desc{This flag list the paths content in long form}
\item[] ls -lh {\pth{Path}}\desc{Make the filesize humanreable}
\item[] ls -a {\pth{Path}}\desc{Show hidden directories and files}
%\item[] pwd \desc{Print working directory}
\end{itemize}

\section{SSH/SCP}
\begin{itemize}
\item ssh {\lft}user{\rht}@{\lft}host{\rht}
\item ssh -Y {\lft}user{\rht}@{\lft}host{\rht}\desc{Enable X11 forwarding}
\item scp~{\lft}file{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}\desc{Copy a file to a remote location}
\item scp~-r~{\lft}dir{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}\desc{Recursivly copy a directory to a remote location}
\end{itemize}

\section{SLURM}
\subsection{Management And Monitoring}
\begin{itemize}
\item[] sbatch \pth{slurm\_script}\desc{Submit a slurm script to the scheduler for batch processing}
\item[] sbatch~--array=[\#-\#]~\pth{slurm\_script}\desc{Submit a job array to the scheduler}
\item[] srun \desc{Used in submitting an interactive job, or also  used within a slurm\_script}
\item[] srun.x11 \desc{Used in submitting an interactive job, that also forwards the X, if you logged in while enabling X forwarding}
\item{} squeue \desc{Display all jobs running or waiting on the cluster}
\item{} squeue -u{\lft}user{\rht} \desc{Only display jobs for the specified user}
\item{} sinfo \desc{Display the partitions and nodes state}
\item{} scancel {\lft}jobid{\rht} \desc{Cancel the specified job}
\end{itemize}

\subsection{Details and Profiling}
\begin{itemize}
\item[] scontrol show jobid {\lft}jobid{\rht}\desc{Show job details of a waiting or running job}
\item[] scontrol show parition {\lft}partition name{\rht}\desc{Show details of the specified partition}
\item[] scontrol show node {\lft}node name{\rht}\desc{Show details of the specified node}
\item[] sacct -l -j {\lft}jobid{\rht}\desc{Show usage details of a job}
\item[] sacct -u {\lft}user{\rht}\desc{Show all jobs from a given user}
\end{itemize}

\section{Miscellaneous}
\subsection{Commands}
\begin{itemize}
\item[] less \pth{file} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, and pageup/pagedown keys\\  Press 'q' to exit out of less}
\item[] less -S \pth{file} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, and pageup/pagedown keys\\Press 'q' to exit out of less}

\end{itemize}

\subsection{Operators}
\begin{itemize}
\item[] | \desc{Pipe operator is used between two commands, in which you wish to pass the output of the 1st command to the input of the 2nd command\\Ex. \texttt{sacct -l -j 467 | less -S }}
\end{itemize}

\subsection{Hosts}
\begin{itemize}
\item[] login-0001 -- uhhpc1.its.hawaii.edu
\item[] login-0002 -- uhhpc2.its.hawaii.edu
\end{itemize}

\subsection{Globus Endpoints}
\begin{itemize}
\item[] login-0001 -- hawaii#UHHPC1
\item[] login-0002 -- hawaii#UHHPC2
\end{itemize}

\end{document}

### 2016-01-22 hpc_101/cheatsheet/paper.tex
minor edit

@@
 \begin{itemize}
-\item[] less \pth{file} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, and pageup/pagedown keys\\  Press 'q' to exit out of less}
-\item[] less -S \pth{file} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, and pageup/pagedown keys\\Press 'q' to exit out of less}
+\item[] less \pth{file} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\  Press 'q' to exit out of less}
+\item[] less -S \pth{file} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
 

### 2016-01-22 hpc_101/cheatsheet/paper.tex
style changes

@@
 \begin{itemize}
-\item[] cd {\pth{Directory}}\desc{Change to directory specified}
-\item[] ls {\pth{Path}}\desc{List the contents of the specified path}
-\item[] ls -l {\pth{Path}}\desc{This flag list the paths content in long form}
-\item[] ls -lh {\pth{Path}}\desc{Make the filesize humanreable}
-\item[] ls -a {\pth{Path}}\desc{Show hidden directories and files}
+\item[] \texttt{cd {\pth{Directory}}}\desc{Change to directory specified}
+\item[] \texttt{ls {\pth{Path}}}\desc{List the contents of the specified path}
+\item[] \texttt{ls -l {\pth{Path}}}\desc{This flag list the paths content in long form}
+\item[] \texttt{ls -lh {\pth{Path}}}\desc{Make the filesize humanreable}
+\item[] \texttt{ls -a {\pth{Path}}}\desc{Show hidden directories and files}
 %\item[] pwd \desc{Print working directory}
@@
 \begin{itemize}
-\item ssh {\lft}user{\rht}@{\lft}host{\rht}
-\item ssh -Y {\lft}user{\rht}@{\lft}host{\rht}\desc{Enable X11 forwarding}
-\item scp~{\lft}file{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}\desc{Copy a file to a remote location}
-\item scp~-r~{\lft}dir{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}\desc{Recursivly copy a directory to a remote location}
+\item \texttt{ssh {\lft}user{\rht}@{\lft}host{\rht}}
+\item \texttt{ssh -Y {\lft}user{\rht}@{\lft}host{\rht}}\desc{Enable X11 forwarding}
+\item \texttt{scp~{\lft}file{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Copy a file to a remote location}
+\item \texttt{scp~-r~{\lft}dir{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Recursivly copy a directory to a remote location}
 \end{itemize}
@@
 \begin{itemize}
-\item[] sbatch \pth{slurm\_script}\desc{Submit a slurm script to the scheduler for batch processing}
-\item[] sbatch~--array=[\#-\#]~\pth{slurm\_script}\desc{Submit a job array to the scheduler}
-\item[] srun \desc{Used in submitting an interactive job, or also  used within a slurm\_script}
-\item[] srun.x11 \desc{Used in submitting an interactive job, that also forwards the X, if you logged in while enabling X forwarding}
-\item{} squeue \desc{Display all jobs running or waiting on the cluster}
-\item{} squeue -u{\lft}user{\rht} \desc{Only display jobs for the specified user}
-\item{} sinfo \desc{Display the partitions and nodes state}
-\item{} scancel {\lft}jobid{\rht} \desc{Cancel the specified job}
+\item[] \texttt{sbatch \pth{slurm\_script}}\desc{Submit a slurm script to the scheduler for batch processing}
+\item[] \texttt{sbatch~--array=[\#-\#]~\pth{slurm\_script}}\desc{Submit a job array to the scheduler}
+\item[] \texttt{srun} \desc{Used in submitting an interactive job, or also  used within a slurm\_script}
+\item[] \texttt{srun.x11} \desc{Used in submitting an interactive job, that also forwards the X, if you logged in while enabling X forwarding}
+\item{} \texttt{squeue} \desc{Display all jobs running or waiting on the cluster}
+\item{} \texttt{squeue -u{\lft}user{\rht}} \desc{Only display jobs for the specified user}
+\item{} \texttt{sinfo} \desc{Display the partitions and nodes state}
+\item{} \texttt{scancel {\lft}jobid{\rht}} \desc{Cancel the specified job}
 \end{itemize}
@@
 \begin{itemize}
-\item[] scontrol show jobid {\lft}jobid{\rht}\desc{Show job details of a waiting or running job}
-\item[] scontrol show parition {\lft}partition name{\rht}\desc{Show details of the specified partition}
-\item[] scontrol show node {\lft}node name{\rht}\desc{Show details of the specified node}
-\item[] sacct -l -j {\lft}jobid{\rht}\desc{Show usage details of a job}
-\item[] sacct -u {\lft}user{\rht}\desc{Show all jobs from a given user}
+\item[] \texttt{scontrol show jobid {\lft}jobid{\rht}}\desc{Show job details of a waiting or running job}
+\item[] \texttt{scontrol show parition {\lft}partition name{\rht}}\desc{Show details of the specified partition}
+\item[] \texttt{scontrol show node {\lft}node name{\rht}}\desc{Show details of the specified node}
+\item[] \texttt{sacct -l -j {\lft}jobid{\rht}}\desc{Show usage details of a job}
+\item[] \texttt{sacct -u {\lft}user{\rht}}\desc{Show all jobs from a given user}
 \end{itemize}
@@
 \begin{itemize}
-\item[] less \pth{file} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\  Press 'q' to exit out of less}
-\item[] less -S \pth{file} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
-
+\item[] \texttt{less \pth{file}} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\  Press 'q' to exit out of less}
+\item[] \texttt{less -S \pth{file}} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
 \end{itemize}
-
+\bigskip
 \subsection{Operators}
 \begin{itemize}
-\item[] | \desc{Pipe operator is used between two commands, in which you wish to pass the output of the 1st command to the input of the 2nd command\\Ex. \texttt{sacct -l -j 467 | less -S }}
+\item[] \texttt{|} \desc{Pipe operator is used between two commands, in which you wish to pass the output of the 1st command to the input of the 2nd command\\Ex. \texttt{sacct -l -j 467 | less -S }}
 \end{itemize}
@@
 \begin{itemize}
-\item[] login-0001 -- uhhpc1.its.hawaii.edu
-\item[] login-0002 -- uhhpc2.its.hawaii.edu
+\item[] login-0001 -- \texttt{uhhpc1.its.hawaii.edu}
+\item[] login-0002 -- \texttt{uhhpc2.its.hawaii.edu}
 \end{itemize}
@@
 \begin{itemize}
-\item[] login-0001 -- hawaii#UHHPC1
-\item[] login-0002 -- hawaii#UHHPC2
+\item[] login-0001 -- \texttt{hawaii#UHHPC1}
+\item[] login-0002 -- \texttt{hawaii#UHHPC2}
 \end{itemize}

### 2016-01-22 hpc_101/cheatsheet/paper.tex
minor tweeks

@@
 \begin{itemize}
-\item \texttt{ssh {\lft}user{\rht}@{\lft}host{\rht}}
-\item \texttt{ssh -Y {\lft}user{\rht}@{\lft}host{\rht}}\desc{Enable X11 forwarding}
-\item \texttt{scp~{\lft}file{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Copy a file to a remote location}
-\item \texttt{scp~-r~{\lft}dir{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Recursivly copy a directory to a remote location}
+\item[] \texttt{ssh {\lft}user{\rht}@{\lft}host{\rht}}
+\item[] \texttt{ssh -Y {\lft}user{\rht}@{\lft}host{\rht}}\desc{Enable X11 forwarding}
+\item[] \texttt{scp~{\lft}file{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Copy a file to a remote location}
+\item[] \texttt{scp~-r~{\lft}dir{\rht}~{\lft}user{\rht}@{\lft}host{\rht}:{\lft}remote~path{\rht}}\desc{Recursivly copy a directory to a remote location}
 \end{itemize}
@@
 \item[] \texttt{srun.x11} \desc{Used in submitting an interactive job, that also forwards the X, if you logged in while enabling X forwarding}
-\item{} \texttt{squeue} \desc{Display all jobs running or waiting on the cluster}
-\item{} \texttt{squeue -u{\lft}user{\rht}} \desc{Only display jobs for the specified user}
-\item{} \texttt{sinfo} \desc{Display the partitions and nodes state}
-\item{} \texttt{scancel {\lft}jobid{\rht}} \desc{Cancel the specified job}
+\item[] \texttt{squeue} \desc{Display all jobs running or waiting on the cluster}
+\item[] \texttt{squeue -u{\lft}user{\rht}} \desc{Only display jobs for the specified user}
+\item[]\texttt{sinfo} \desc{Display the partitions and nodes state}
+\item[] \texttt{scancel {\lft}jobid{\rht}} \desc{Cancel the specified job}
 \end{itemize}
@@
 \item[] \texttt{scontrol show jobid {\lft}jobid{\rht}}\desc{Show job details of a waiting or running job}
-\item[] \texttt{scontrol show parition {\lft}partition name{\rht}}\desc{Show details of the specified partition}
+\item[] \texttt{scontrol show parition {\lft}partition{\rht}}\desc{Show details of the specified partition}
 \item[] \texttt{scontrol show node {\lft}node name{\rht}}\desc{Show details of the specified node}
@@
 \begin{itemize}
-\item[] login-0001 -- \texttt{hawaii#UHHPC1}
-\item[] login-0002 -- \texttt{hawaii#UHHPC2}
+\item[] login-0001 -- \texttt{hawaii\#UHHPC1}
+\item[] login-0002 -- \texttt{hawaii\#UHHPC2}
 \end{itemize}

### 2016-01-22 hpc_101/cheatsheet/paper.tex
addition of module commands

@@
 
+\section{Modules}
+\begin{itemize}
+\item[] \texttt{module avail} \desc{List all available software modules}
+\item[] \texttt{module load \pth{module name}} \desc{Load the specified module into your environment}
+\item[] \texttt{module show \pth{module name}} \desc{Shows you what actions the module performs to your environment}
+\item[] \texttt{module purge} \desc{unload all loaded modules}
+\end{itemize}
+
 \section{Miscellaneous}
@@
 \begin{itemize}
-\item[] \texttt{less \pth{file}} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\  Press 'q' to exit out of less}
+\item[] \texttt{less \pth{file}} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
 \item[] \texttt{less -S \pth{file}} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
+\item[] \texttt{man \pth{command}} \desc{Show the manual entry (if one exists) for the given command\\Press 'q' to exit out of less}
 \end{itemize}
-\bigskip
+
 \subsection{Operators}

### 2016-01-22 hpc_101/paper.tex
Update to partition table

@@
 \hline
-lm.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
+lm.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
 \hline
-sb.q & \begin{tabular}{l r}\textbf{Default} & 0-00:05:00~\\ \textbf{Max} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+sb.q & \begin{tabular}{l r}\textbf{Default} & 0-00:05:00~\\ \textbf{Max} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \bottomrule 

### 2016-01-22 hpc_101/paper.tex
corrected slurm expanded

@@
 	\bigskip
-	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
+	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility for \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
 	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}

### 2016-01-22 hpc_101/paper.tex
Defined Job Schedulers and last slides reordering

--- 2016-01-22 hpc_101/paper.tex
@@
 \begin{frame}
-	\frametitle{What is a Job Scheduler?}
-	Basic features expected of a job scheduler include:
+  \frametitle{Job Schedulers}
+  \begin{block}{Purpose}To control and prioritize the execution order of unrelated jobs\end{block}
+	Basic features expected of a job scheduler:
 	\begin{itemize}\footnotesize
-        \item The capability to control and prioritize the execution order of unrelated jobs
-        \item Allocate exclusive and/or non-exclusive access to resources to users for some duration of time so they can perform work
-	\item Provides a framework for starting, executing, and monitoring work on the set of allocated resources
-	\item Arbitrate contention for resources by managing a queue of pending work
+        \item Provides a user interface for users to request resources and monitor work 
+        \item Allocates access to resources for the requested duration of time
+	\item Starts, monitors and terminates work on allocated resources
+	\item Arbitrates contention for resources by managing queues of pending work
 	\end{itemize}
 	\bigskip
-	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
+        The {\craycs} uses the \textbf{S}imple \textbf{L}inux \textbf{U}tility for \textbf{R}esource \textbf{M}anagement~\\ or simply known as the \emph{SLURM scheduler}
 	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
@@
 	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
-	\item \emph{\textbf{srun.x11}} -- Used to submit an interactive job with x11 support
+	\item \emph{\textbf{srun.x11}} -- Used to submit an interactive job with X11 support
 	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
@@
 
-\section{Frequently Asked Questions}
+\begin{frame}
+\Huge{\centerline{Questions?}}
+\end{frame}
+
+%\section{Frequently Asked Questions}
 \begin{frame}
@@
 
-\begin{frame}
-\Huge{\centerline{Questions?}}
-\end{frame}
-
 \end{document}

### 2016-01-22 hpc_101/paper.tex
Fixed some spelling

@@
  \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
- \item Deilvered in October of 2014
+ \item Delivered in October of 2014
  \item Accepted in December of 2014 
@@
 			\item Acquiring software
-			\item Compiliers
+			\item Compilers
 			}}
@@
 \midrule
-community.q & \begin{tabular}{l r} \textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \hline
-exclusive.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min}& \padfrom{\ \ \ }{1}~\\ \textbf{Max}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \hline
-kill.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+exclusive.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:}& \padfrom{\ \ \ }{1}~\\ \textbf{Max:}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \hline
-htc.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \hline
-lm.q & \begin{tabular}{l r}\textbf{Default} & 0-00:10:00~\\ \textbf{Max} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
+kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \hline
-sb.q & \begin{tabular}{l r}\textbf{Default} & 0-00:05:00~\\ \textbf{Max} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min} & \padfrom{\ \ \ }{1}~\\ \textbf{Max} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+htc.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+\hline
+\hline
+lm.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
+\hline
+\hline
+sb.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:05:00~\\ \textbf{Max:} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
 \bottomrule 
@@
 	\begin{block}{Interactive session (no X11)}\tiny
-	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00 \ddash{}pty /bin/bash
+	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
 	\end{block}
@@
 		$[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
-		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00	~\\
+		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
 		$[$compute-0001 \ctilde$]$\$ xterm	~\\
@@
 \item[--] If we identify or are notified that a user is running computation on the login nodes, the application will be killed
-\item[--] If we determine that a user repeatly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
+\item[--] If we determine that a user repeatedly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
 \end{itemize}

### 2016-01-22 hpc_101/cheatsheet/paper.tex
update to some wording in a few command description

@@
 \item[] \texttt{module load \pth{module name}} \desc{Load the specified module into your environment}
-\item[] \texttt{module show \pth{module name}} \desc{Shows you what actions the module performs to your environment}
+\item[] \texttt{module show \pth{module name}} \desc{Shows you what actions the module performs on your environment}
 \item[] \texttt{module purge} \desc{unload all loaded modules}
@@
 \begin{itemize}
-\item[] \texttt{less \pth{file}} \desc{Show contents of file\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
-\item[] \texttt{less -S \pth{file}} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, spacebar, or pageup/pagedown keys\\Press 'q' to exit out of less}
-\item[] \texttt{man \pth{command}} \desc{Show the manual entry (if one exists) for the given command\\Press 'q' to exit out of less}
+\item[] \texttt{less \pth{file}} \desc{Show contents of file\\Navigate using up/down arrow keys, or spacebar\\Press 'q' to exit out of less}
+\item[] \texttt{less -S \pth{file}} \desc{Show contents of file without line wrap\\Navigate using up/down arrow keys, or spacebar\\Press 'q' to exit out of less}
+\item[] \texttt{man \pth{command}} \desc{Show the manual entry (if one exists) for the given command\\Uses the same controls as less\\Press 'q' to exit out of man}
 \end{itemize}

### 2016-01-22 hpc_101/paper.tex
missing word in faq

@@
   \begin{itemize}\tiny
-  \item \textbf{Answer:} We are willing to help, just send us an email at \textit{uh-hpc-help@lists.hawaii.edu} with a description of what you have attempted as well a link to the software you are trying to compile, and/or where it is located on the cluster.
+  \item \textbf{Answer:} We are willing to help, just send us an email at \textit{uh-hpc-help@lists.hawaii.edu} with a description of what you have attempted as well as a link to the software you are trying to compile, and/or where it is located on the cluster.
   \end{itemize}

### 2016-01-23 hpc_101/paper.tex
Clarification on time format Removal of confusing cd in the example slurm scripts

@@
 			\item[--] rsync (rsync protocol with SSH transport)
-			\item[--] SFTP (SSH FTP protocol)
+			\item[--] SFTP (SSH FTP protocol) -- Filezilla, Cyberduck
 			\item[--] \href{https://www.globus.org/}{Globus} (Grid FTP protocol)
@@
 \#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00 
+\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
 \#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
@@
 
-cd \$SLURM\_SUBMIT\_DIR
 srun  -n \$\{SLURM\_NTASKS\}  ./hello\_mpi.intel 
@@
 \#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00
+\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
 \#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
@@
 
-cd \$SLURM\_SUBMIT\_DIR
 ./hello\_world
@@
 
-cd \$SLURM\_SUBMIT\_DIR
 ./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.flower -o output\_\$\{SLURM_ARRAY_TASK_ID\}.honey
@@
 \end{center}
-
 \end{frame}
@@
 
-\item HELP! My job needs more time and does not support checkpointing.  What can I do?
+\item HELP! My job needs more time and does not support check-pointing.  What can I do?
   \begin{itemize}\tiny

### 2016-02-05 hpc_101/paper.tex
removed the cd action in example slides
fixed typo in globus endpoint
forgot to have the "source ~/.bash_profile" in the in slide examples...

@@
 \begin{itemize}\tiny
- \item Each of the login nodes is also an endpoint: \textbf{\emph{hawaii\#UHHPC1}} \& \textbf{\emph{hawaii\#UHHPC1}}
+ \item Each of the login nodes is also an endpoint: \textbf{\emph{hawaii\#UHHPC1}} \& \textbf{\emph{hawaii\#UHHPC2}}
 \end{itemize}
@@
 
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
+
 \#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
@@
 
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
+
 \#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-
 export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
@@
 \#SBATCH \ddash{}mail-user=user@test.org
+
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
 
@@
 
-%\section{Frequently Asked Questions}
+\section{Frequently Asked Questions}
 \begin{frame}

### 2016-02-22 hpc_101/paper.tex
added queue to initial free listing
Added a slide about additional servcies, and listed MATLAB

@@
 
+\newcommand{\trademark}{\fontsize{5}{6}\selectfont \textsuperscript{\texttrademark}}
 \newcommand{\regtrademark}{\fontsize{5}{6}\selectfont \textsuperscript{\textregistered}}
@@
 		\item The initial investment in the cluster provides resources for all faculty, staff, and students affiliated with the University of {\hawaii}
-		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, lm.q, sb.q, kill.q, htc.q}
+		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, sb.q, kill.q, htc.q}
 		\item For some users, the publicly available resources may not be enough~\ldots
@@
 
+\section{Related Services}
+\begin{frame}
+\frametitle{Services}
+\begin{itemize}
+\item MATLAB Distributed Computing Server{\trademark}
+  \begin{itemize}
+    \item Requires user to already have a license for MATLAB
+    \item User must also have access to at least the Parallel Computing Toolbox{\trademark}
+    \item Provides users access to up to 64 workers
+    \item Requires some additional configuration
+    \item Email us to request access to this additional service
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
 \section{Frequently Asked Questions}

### 2016-02-22 hpc_101/paper.tex
changes to srun example and sbatch examples

@@
 	\begin{block}{Interactive session (no X11)}\tiny
-	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
+	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
+          ~\\
+          ~\\
+          \hrule
+          ~\\
+	$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash
+
 	\end{block}
@@
 		$[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
-		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
+		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
 		$[$compute-0001 \ctilde$]$\$ xterm	~\\
@@
 \#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes
-\#SBATCH \ddash{}error=hello-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\#SBATCH \ddash{}output=hello-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid.
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid.
 \#\# Useful for remote notification

### 2016-03-02 matlab_dcs_101/paper.tex
added info on the optional parameter

@@
           \only<6>{\footnotesize{Each of the functions in the \textbf{SUBMIT FUNCTIONS} section, have the follow format:}\\
-            \textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}} } 
+            \textbf{\tiny{\{$<$func$>$,`$<$server addr$>$', `$<$remote storage dir$>$', `$<$partition$>$', `$<$timelimit$>$', [$<$mem usage in MB$>$] \} } }
           }
-          \only<7>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}}}\\Replace the remote storage dir with the path that matches your users lus folder location i.e., /home/$<$username$>$/lus/}}
-          \only<8>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$slurm partition$>$', `$<$slurm timelimit$>$'\}}}\\Adjust the slurm partition and slurm timelimit to fit the partition you wish to run on, and finally click \textbf{done}}}
+          \only<7>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$partition$>$', `$<$timelimit$>$', [$<$mem usage in MB$>$] \}}}\\Replace the remote storage dir with the path that matches your users lus folder location i.e., /home/$<$username$>$/lus/}}
+          \only<8>{\normalsize{\textbf{\tiny{\{`$<$func$>$',`$<$server addr$>$', `$<$remote storage dir$>$', `$<$partition$>$', `$<$timelimit$>$', [$<$mem usage in MB$>$] \}}}\\Adjust the slurm partition and slurm timelimit to fit the partition you wish to run on, and finally click \textbf{done}}}
 					\only<9>{\normalsize{The profile should now be configured for your user on \CNAME }}

### 2016-03-02 hpc_101/paper.tex
update to slides. moved x11 to aux sectio.  swtiched to --memn

@@
 	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
-	\item \emph{\textbf{srun.x11}} -- Used to submit an interactive job with X11 support
 	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
+	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
 	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
 	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
-	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
 	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
@@
 \frametitle{Interactive Job with SLURM}
-	\begin{block}{Interactive session (no X11)}\tiny
+	\begin{block}{Interactive session}\tiny
 	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
@@
 	\end{block}
-
-	\begin{block}{Interactive session (with X11)}\tiny
-		\begin{enumerate}
-			\item Connect via SSH using the -Y option, X11 forwarding enabled
-			\item run srun.x11 to start a session on a node
-		\end{enumerate}
-		~\\
-		~\\
-		$[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
-		$[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
-		$[$compute-0001 \ctilde$]$\$ xterm	~\\
-	\end{block}
 \end{frame}
@@
 
-\#!/bin/sh
+\#!/bin/bash
 \#SBATCH \ddash{}job-name=MPI\_example
@@
 \#SBATCH \ddash{}cpus-per-task=1
-\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid.
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid.
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
 \#\# Useful for remote notification
@@
 
-\#!/bin/sh
+\#!/bin/bash
 \#SBATCH \ddash{}job-name=example
@@
 \#SBATCH \ddash{}cpus-per-task=5
-\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid.
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid.
+\#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
 \#\# Useful for remote notification
@@
 \#SBATCH \ddash{}cpus-per-task=5
-\#SBATCH \ddash{}mem-per-cpu=6400 \#\# max of 6400 for standard nodes, max of 26214 for large memory nodes
+\#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
 \#SBATCH \ddash{}error=buzz-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
@@
    \begin{itemize}
-   \item \textbf{Answer:} Current frequency is every other month, but the frequency many increase or decrease based on the fill rate of the {\lustre} filesystem 
+   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
    \end{itemize}
@@
 
+\begin{frame}	
+\frametitle{FAQ}
+\begin{itemize}
+  \item Can I run a display program on the cluster?
+    \begin{itemize}
+      \item \textbf{Answer:} Yes, but not from the login nodes.  To follow cluster usage policies, please run X11 applications from a compute node.  Please see the following steps.
+\end{itemize}
+\end{itemize}
+ \begin{block}{Interactive session with X11}\tiny  
+       \begin{enumerate}
+      \item Connect via SSH using the -Y option, X11 forwarding enabled
+      \item run srun.x11 to start a session on a node
+      \end{enumerate}
+      \
+      ~\\
+      ~\\
+      $[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
+      $[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
+      $[$compute-0001 \ctilde$]$\$ xterm	~\\
+ \end{block}
+
+\end{frame}
+
 \end{document}

### 2016-03-22 hpc_101/paper.tex
Introduction of a GPU submission script to the slide deck

@@
 
+\subsection{MPI Batch Job Example Script}
+\begin{frame}[fragile]
+\frametitle{SLURM sbatch -- Submission Script File -- GPU}
+\begin{semiverbatim}\tiny
+[login lus]\$ cat gpu.slurm
+
+\#!/bin/bash
+\#SBATCH \ddash{}job-name=GPU\_example
+\#SBATCH \ddash{}partition=gpu.q
+\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
+\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
+\#SBATCH \ddash{}nodes=1
+\#SBATCH \ddash{}tasks-per-node=1
+\#SBATCH \ddash{}cpus-per-task=10
+\#SBATCH \ddash{}gres=gpu:NV-K40:2  \#\# request both GPUs in the GPU node.
+\#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
+\#\# Useful for remote notification
+\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
+\#SBATCH \ddash{}mail-user=user@test.org
+
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
+module load GPGPU/cuda/samples/7.5
+
+\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
+export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
+
+bandwidthTest; cudaOpenMP; deviceQuery; simpleMultiGPU
+\end{semiverbatim}
+\end{frame}
+
 \subsection{Executing and Monitoring Batch Jobs}

### 2016-03-29 hpc_101/paper.tex
fixed cray CS3000 -> CS300
addition of owner nodes

@@
  \item Opened for general use in April 2015
- \item As of October 2015, more than 200 users have been granted access to the cluster
+ \item March 2016, 92 new nodes added to cluster, raising total core count to 5,876 cores
+ \item As of March 2016, more than 240 users have been granted access to the cluster
 \end{itemize}
@@
 	\begin{itemize}
-		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
-		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
-			\begin{itemize}
-				\item Two 10 core {\intel} processors (\emph{20 cores total})
-				\item Diskless -- Some RAM is used for the Operating System
-				\item 128GB of physical RAM
-				\item $\approx110$GB of useable RAM
-			\end{itemize}
-		\item 6 large memory nodes
-			\begin{itemize}
-				\item Four 10 core {\intel} processors (\emph{40 cores total})
-				\item Diskless -- Some RAM is used for the Operating System
-				\item 1TB of physical RAM
-				\item $\approx1008$GB of useable RAM
+        \item All nodes are diskless with some RAM used for the Operating System
+	\item 178 Standard nodes
+	  \begin{itemize}
+            {\footnotesize
+	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
+	    \item 128GB of physical RAM with $\approx110$GB of useable RAM
+            }
+	  \end{itemize}
+
+	\item 6 Large memory nodes
+	  \begin{itemize}
+            {\footnotesize
+	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
+	    \item 1TB of physical RAM with $\approx1008$GB of useable RAM
+            }
+	  \end{itemize}			
+        \item 1 GPU node
+          \begin{itemize}
+            {\footnotesize
+	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
+	    \item 128GB of physical RAM with $\approx110$GB of useable RAM
+            \item 2 Nvidia Tesla K40 GPUs
+            }
+	  \end{itemize}
+        \item 91 Owner nodes
+          \begin{itemize}
+            {\footnotesize
+            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
+            \item 58 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
+            }
           \end{itemize}

### 2016-04-05 hpc_101/paper.tex
updated onboarding sildes to have our directory listed.  Added in the GPU partition to sections we list our partitions out

@@
 
-\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
+\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\Director\\Gwen Jacobs Ph.D}
 \institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
@@
 \begin{frame}
-	\frametitle{{\craycs}0 -- Compute Nodes}
+	\frametitle{{\craycs} -- Compute Nodes}
 	\begin{itemize}
@@
 		\item The initial investment in the cluster provides resources for all faculty, staff, and students affiliated with the University of {\hawaii}
-		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, sb.q, kill.q, htc.q}
+		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
 		\item For some users, the publicly available resources may not be enough~\ldots
@@
 		{\large \textbf{Service Unit Price Card}}\\
-		\begin{tabular}{l c r }
+		\begin{tabular}{l || c || r}
 			\toprule
@@
 			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
+                        \item \href{http://mobaxterm.mobatek.net/}{MobaXterm}
 		\end{itemize}
@@
 \begin{itemize}
-	\item The {\craycs} currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
+	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
 	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
@@
 \midrule
-community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
 \hline
-exclusive.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:}& \padfrom{\ \ \ }{1}~\\ \textbf{Max:}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+exclusive.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:}& \padfrom{\ \ \ }{1}~\\ \textbf{Max:}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
 \hline
-kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+gpu.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
 \hline
-htc.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
 \hline
-lm.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ }{262100}\end{tabular}\\
+htc.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
 \hline
-sb.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:05:00~\\ \textbf{Max:} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ }{6500}\end{tabular}\\
+lm.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
+\hline
+\hline
+sb.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:05:00~\\ \textbf{Max:} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \bottomrule 

### 2016-04-07 hpc_101/cheatsheet/paper.tex
update of author list to include director

@@
 \date{\today}
-\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\University of {\hawaii}\\Information Technology Services -- {\ci}}
+\author{Gwen Jacobs Ph.D,\\Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\University of {\hawaii}\\Information Technology Services -- {\ci}}
 \maketitle

### 2016-04-07 hpc_101/paper.tex
update of author list to include director

@@
 
-\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\Director\\Gwen Jacobs Ph.D}
+\author{Gwen Jacobs Ph.D,\\Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.}
 \institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
@@
 	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx110$GB of useable RAM
+	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
             }
@@
 	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
-	    \item 1TB of physical RAM with $\approx1008$GB of useable RAM
+	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
             }
@@
 	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx110$GB of useable RAM
+	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
             \item 2 Nvidia Tesla K40 GPUs
@@
 		\item $500$TB of scale out storage
-		\item Currently only available for purchase by cluster users and only accessible via the login nodes
-		\item ValueStorage will eventually be available for purchase by everyone
-		\item ValueStorage owners will potentially be able to mount as a network drive (CIFS) on laptop, workstations, servers
+		\item Is only available for purchase by cluster users~\\and only accessible via the login nodes
 		\item Purchased in 0.5TB increments

### 2016-04-22 hpc_101/paper.tex
added problem reporting and fixed a typo

@@
 
+\section{Problem Reporting}
+\begin{frame}
+\frametitle{Problem Reporting}
+\begin{itemize}
+\item Always email UH-HPC-Help@lists.hawaii.edu
+\item For batch jobs...
+\begin{itemize}
+\item[--] Job ID, path to submission script, submission command, error file location, output files
+\end{itemize} 
+\item For other problems...
+\begin{itemize}
+\item[--] State the problem, command issued, host, directory, remote host, error messages
+\end{itemize}
+\end{itemize}
+\end{frame}
+
 \section{Cluster Etiquette}
@@
 \item Try to request only the resources that your job needs.  Leaving unneeded resources available, allows other users to use them
-\item Test what your application in an interactive session before scheduling a long running job
+\item Test your application in an interactive session before scheduling a long running job
 \begin{itemize}

### 2016-05-11 hpc_101/paper.tex
switched to latex elipsis
updates to the text on several slides.  Offset commands during testing of user accounts to show commands in blue

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Beamer Presentation
@@
 \usepackage{xstring}
+\usepackage{color}
 \mode<presentation> {
@@
  \item Opened for general use in April 2015
- \item March 2016, 92 new nodes added to cluster, raising total core count to 5,876 cores
- \item As of March 2016, more than 240 users have been granted access to the cluster
+ \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
+ \item As of March 2016, more than 257 users have been granted access to the cluster
 \end{itemize}
@@
 		\item {\lustre}
+                  \begin{itemize}
+                  \item High performance parallel filesystem 
+                  \item Available on all the cluster nodes
+                    \item Freely available to all users
+                  \end{itemize}
 		\item ValueStorage
+                  \begin{itemize}
+                  \item Network attached scale out storage
+                  \item Only available from the login nodes
+                  \item Available as a for fee service
+                  \end{itemize}
 	\end{enumerate}
@@
 	\begin{itemize}
-		\item {\lustre} is a high performance parallel filesystem
 		\item The {\craycs} has $\approx582$TB of storage space
-		\item Shared between all compute nodes and login nodes
+
 		\item Primarily used as scratch space for jobs (Input and Output)
@@
 	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
+        ~\\
 	\begin{itemize}
 		\item $500$TB of scale out storage
-		\item Is only available for purchase by cluster users~\\and only accessible via the login nodes
-		\item Purchased in 0.5TB increments
+		\item Purchased annually in 0.5TB increments 
 	\end{itemize}
@@
 	\textbf{ValueStorage Pricing}~\\ 
-		\resizebox{0.75\textwidth}{!}{%
-		\begin{tabular}{l r ||l r }
+		\resizebox{0.50\textwidth}{!}{%
+		\begin{tabular}{l r }
 			\toprule
-			\textbf{Product} & \textbf{Annual Cost}~~ & ~~\textbf{Product} & \textbf{Annual Cost}\\
+			\textbf{Product} & \textbf{Annual Cost}~~ \\
 	                \midrule
 			\midrule
-			0.5TB & \$65.00~~ & ~~0.5TB + Replication  & \$130.00 \\
+			0.5TB & \$65.00\\
+                        0.5TB + Replication  & \$130.00 \\
 			\bottomrule
@@
 	\begin{itemize}
-		\item The initial investment in the cluster provides resources for all faculty, staff, and students affiliated with the University of {\hawaii}
+		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
 		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
@@
 	\item[] \begin{itemize}
-		\item To connect to the cluster, we utilize a client which communicates using the Secure Shell (\emph{SSH}) protocol 
-		\item Linux and MacOSX, typically have a SSH client already installed
-		\item Windows typically does not come with an SSH client installed
-		\item Windows 10 may come pre-installed with a SSH client, but it might not be stable
+		\item To connect to the cluster, an SSH client is required
+		\item Linux and MacOS X, typically have a SSH client already installed
+		\item Windows typically needs a 3rd party SSH client
 		\item Suggested SSH clients for Windows include:
@@
 	\end{description}
+        ~\\~\\
 	{\large Let's attempt to login!}
@@
 	\begin{enumerate}
-		\item Test that you can list files in your home: `\texttt{ls -la}'    
-		\item Test making a file in your home: `\texttt{touch test.txt}'    
-		\item Go into \ctilde{}/lus: `\texttt{cd \ctilde{}/lus/}'  
-		\item Test making a file in your lus directory: `\texttt{touch test.txt}'   
-		\item Go into \ctilde{}/apps:  `\texttt{cd \ctilde{}/apps/}'     
-		\item Test making a file in your apps directory: `\texttt{touch test.txt}'   
+		\item Test that you can list files in your home: `\textcolor{blue}{\texttt{ls -la}}'    
+		\item Test making a file in your home: `\textcolor{blue}{\texttt{touch test.txt}}'    
+		\item Go into \ctilde{}/lus: `\textcolor{blue}{\texttt{cd \ctilde{}/lus/}}'  
+		\item Test making a file in your lus directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
+		\item Go into \ctilde{}/apps:  `\textcolor{blue}{\texttt{cd \ctilde{}/apps/}}'     
+		\item Test making a file in your apps directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
 	\end{enumerate}
@@
 \item Always email UH-HPC-Help@lists.hawaii.edu
-\item For batch jobs...
+\item For batch jobs \ldots
 \begin{itemize}
@@
 \end{itemize} 
-\item For other problems...
+\item For other problems \ldots
 \begin{itemize}

### 2016-06-09 hpc_101/paper.tex
changed text regarding scheduler priority setting
Addition of the job array image

@@
 \item Runtime
-\item Resource usage request
+\item Amount of resources requested
 \item Age of job
@@
 \#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
-\#SBATCH \ddash{}error=buzz-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\#SBATCH \ddash{}output=buzz-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
+\#SBATCH \ddash{}error=myprog-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
+\#SBATCH \ddash{}output=myprog-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
 \#\# Useful for remote notification
@@
 
-./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.flower -o output\_\$\{SLURM_ARRAY_TASK_ID\}.honey
+./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.dat -o output\_\$\{SLURM_ARRAY_TASK_ID\}
 \end{semiverbatim}
@@
 community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
-[login lus]\$ sbatch {\ddash}array=0-2 job_array.slurm
+[login lus]\$ sbatch {\ddash}array=1-30 job_array.slurm
 sbatch: Submitted batch job 470
@@
 JOBID      PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
-470[0-2]  community.q  example  user99  PD   00:00   1      (resource)
-470_0     community.q  example  user99  R    00:01   1      compute-0001
-470_1     community.q  example  user99  R    00:05   1      compute-0002
+470[1-30]  community.q  example  user99  PD   00:00   1      (resource)
+470_1      community.q  example  user99  R    00:01   1      compute-0001
+470_2      community.q  example  user99  R    00:05   1      compute-0002
 \end{semiverbatim}
 \end{block}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Slurm Job Array}
+  \includegraphics[width=1\textwidth]{images/job_array}
 

### 2016-06-23 hpc_101/paper.tex
rearranged slides regarding batch submission.  Updated globus endpoint names

@@
 \begin{itemize}\tiny
- \item Each of the login nodes is also an endpoint: \textbf{\emph{hawaii\#UHHPC1}} \& \textbf{\emph{hawaii\#UHHPC2}}
+ \item Each of the login nodes is also an endpoint: \textbf{\emph{UHHPC1}} \& \textbf{\emph{UHHPC2}}
 \end{itemize}
@@
 
+\subsection{GPU Batch Job Example Script}
+\begin{frame}[fragile]
+\frametitle{SLURM sbatch -- Submission Script File -- GPU}
+\begin{semiverbatim}\tiny
+[login lus]\$ cat gpu.slurm
+
+\#!/bin/bash
+\#SBATCH \ddash{}job-name=GPU\_example
+\textcolor{blue}{\#SBATCH \ddash{}partition=gpu.q}
+\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
+\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
+\#SBATCH \ddash{}nodes=1
+\#SBATCH \ddash{}tasks-per-node=1
+\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=10}
+\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
+\textcolor{blue}{\#SBATCH \ddash{}gres=gpu:NV-K40:2 } \#\# request both GPUs in the GPU node.
+\#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
+\#\# Useful for remote notification
+\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
+\#SBATCH \ddash{}mail-user=user@test.org
+
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
+module load GPGPU/cuda/samples/7.5
+
+\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
+export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
+
+bandwidthTest; cudaOpenMP; deviceQuery; simpleMultiGPU
+\end{semiverbatim}
+\end{frame}
+
 \subsection{MPI Batch Job Example Script}
@@
 \#SBATCH \ddash{}job-name=MPI\_example
-\#SBATCH \ddash{}partition=exclusive.q
+\textcolor{blue}{\#SBATCH \ddash{}partition=exclusive.q}
 \#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
@@
 \#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=4
-\#SBATCH \ddash{}tasks-per-node=20
-\#SBATCH \ddash{}cpus-per-task=1
+\textcolor{blue}{\#SBATCH \ddash{}nodes=4}
+\textcolor{blue}{\#SBATCH \ddash{}tasks-per-node=20}
+\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=1}
 \#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
@@
 [login lus]\$ cat hello_world.slurm
+
+\#!/bin/bash
+\#SBATCH \ddash{}job-name=example
+\textcolor{blue}{\#SBATCH \ddash{}partition=community.q}
+\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
+\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
+\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
+\#SBATCH \ddash{}nodes=1
+\#SBATCH \ddash{}tasks-per-node=1
+\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=5}
+\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
+\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
+\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
+\#\# Useful for remote notification
+\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
+\#SBATCH \ddash{}mail-user=user@test.org
+
+source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
+
+\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
+export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
+
+./hello\_world
+\end{semiverbatim}
+\end{frame}
+
+\subsection{Executing and Monitoring Batch Jobs}
+\begin{frame}[fragile]
+\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
+To execute a submission script you use the \textbf{\texttt{sbatch}} command
+\begin{block}{Example}
+\begin{semiverbatim}\tiny
+[login lus]\$ sinfo
+PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
+community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
+
+[login lus]\$ sbatch hello_world.slurm
+sbatch: Submitted batch job 469
+
+[login lus]\$ squeue
+JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
+469   community.q  example  user99  R  00:01 1      compute-0001
+\end{semiverbatim}
+\end{block}
+
+\btVFill
+\begin{center}What if I want to run many jobs with identical parameters, but on different inputs?~\\
+  SLURM provides a mechanism for this called a \textbf{Job Array}.\end{center}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Slurm Job Array}
+\begin{center}  
+  \includegraphics[width=0.90\textwidth]{images/job_array}
+\end{center}
+\end{frame}
+
+\subsection{Batch Job Example Script Using Job Array}
+\begin{frame}[fragile]
+\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
+\begin{semiverbatim}\tiny
+[login lus]\$ cat job_array.slurm
 
@@
 \#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
+\textcolor{blue}{\#SBATCH \ddash{}error=myprog-\%A\_\%a.err} \#\# \%A - filled with jobid. \%a - filled with job arrayid
+\textcolor{blue}{\#SBATCH \ddash{}output=myprog-\%A\_\%a.out} \#\# \%A - filled with jobid. \%a - filled with job arrayid
 \#\# Useful for remote notification
@@
 \#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-
-./hello\_world
-\end{semiverbatim}
-\end{frame}
-
-\subsection{Batch Job Example Script Using Job Array}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat job_array.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=example
-\#SBATCH \ddash{}partition=community.q
-\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=1
-\#SBATCH \ddash{}tasks-per-node=1
-\#SBATCH \ddash{}cpus-per-task=5
-\#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
-\#SBATCH \ddash{}error=myprog-\%A\_\%a.err \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\#SBATCH \ddash{}output=myprog-\%A\_\%a.out \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
 
@@
 
-\subsection{MPI Batch Job Example Script}
+\subsection{Executing and Monitoring Job Array Batch Jobs}
+
 \begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File -- GPU}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat gpu.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=GPU\_example
-\#SBATCH \ddash{}partition=gpu.q
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=1
-\#SBATCH \ddash{}tasks-per-node=1
-\#SBATCH \ddash{}cpus-per-task=10
-\#SBATCH \ddash{}gres=gpu:NV-K40:2  \#\# request both GPUs in the GPU node.
-\#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-module load GPGPU/cuda/samples/7.5
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-
-bandwidthTest; cudaOpenMP; deviceQuery; simpleMultiGPU
-\end{semiverbatim}
-\end{frame}
-
-\subsection{Executing and Monitoring Batch Jobs}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
-To execute a submission script you use the \textbf{\texttt{sbatch}} command
-\begin{block}{Example -- Non Job Array}
+\frametitle{SLURM sbatch -- Executing \& Monitoring Job Arrays}\footnotesize
+\begin{block}{Example}
 \begin{semiverbatim}\tiny
@@
 community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
-[login lus]\$ sbatch hello_world.slurm
-sbatch: Submitted batch job 469
-[login lus]\$ squeue
-JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
-469   community.q  example  user99  R  00:01 1      compute-0001
-\end{semiverbatim}
-\end{block}
-\begin{block}{Example -- Job Array}
-\begin{semiverbatim}\tiny
-[login lus]\$ sinfo
-PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
-community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
+
 [login lus]\$ sbatch {\ddash}array=1-30 job_array.slurm
 sbatch: Submitted batch job 470
+
 [login lus]\$ squeue
@@
 \end{block}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Slurm Job Array}
-  \includegraphics[width=1\textwidth]{images/job_array}
-
 \end{frame}

### 2016-07-29 hpc_101/paper.tex
fixed typo on globus endpoint and swtiched from tmi to shm:tmi

@@
 \begin{itemize}\tiny
- \item Each of the login nodes is also an endpoint: \textbf{\emph{UHHPC1}} \& \textbf{\emph{UHHPC2}}
+ \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
 \end{itemize}
@@
 	\end{block}
+\btVFill
+\begin{center}Interactive sessions terminate when the specified time has elapsed~\\or if you give the \textbf{exit} command \end{center}
+
 \end{frame}
@@
 \textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
-\textcolor{blue}{\#SBATCH \ddash{}gres=gpu:NV-K40:2 } \#\# request both GPUs in the GPU node.
+\textcolor{blue}{\#SBATCH \ddash{}gres=gpu:NV-K40:2 } \#\# request both GPUs in the GPU node
 \#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
@@
 \textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=1}
+\textcolor{blue}{\#\#SBATCH \ddash{}mem=11000} \# Memory should not be set for jobs in exclusive.q
 \#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
@@
 export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-export I\_MPI\_FABRICS=tmi  
+export I\_MPI\_FABRICS=shm:tmi  
 export I\_MPI\_PMI\_LIBRARY=/opt/local/slurm/default/lib64/libpmi.so
@@
       $[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
-      $[$login \ctilde$]$\$ srun.x11 \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
+      $[$login \ctilde$]$\$ srun.x11 \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
       $[$compute-0001 \ctilde$]$\$ xterm	~\\

### 2016-08-30 hpc_101/paper.tex
author line
Removed names from cheatsheet, and modified the array view to look correct compared to slurm squeue output

@@
 
-\author{Gwen Jacobs Ph.D,\\Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.}
+\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
 \institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
@@
 JOBID      PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
-470[1-30]  community.q  example  user99  PD   00:00   1      (resource)
+470[3-30]  community.q  example  user99  PD   00:00   1      (resource)
 470_1      community.q  example  user99  R    00:01   1      compute-0001

### 2017-01-09 hpc_101/paper.tex
update node count and user count

@@
 
-%\AtBeginSubsection[]
-%{
-%   \begin{frame}
-%       \frametitle{Outline}
-%       \tableofcontents[currentsection,currentsubsection]
-%   \end{frame}
-%}
+%% \AtBeginSubsection[]
+%% {
+%%   \begin{frame}
+%%       \frametitle{Outline}
+%%       \tableofcontents[currentsection,currentsubsection]
+%%   \end{frame}
+%% }
 
@@
  \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
- \item As of March 2016, more than 257 users have been granted access to the cluster
+ \item January 2017, 5 new nodes added to cluster, raising total core count from 5,876 to 5,992 cores
+ \item As of Nov 2016, more than 300 users have been granted access to the cluster
 \end{itemize}
@@
         \item All nodes are diskless with some RAM used for the Operating System
-	\item 178 Standard nodes
+	\item 178 Standard nodes {\tiny(\emph{community})}
 	  \begin{itemize}
@@
 	  \end{itemize}          
-
-	\item 6 Large memory nodes
+	\item 6 Large memory nodes {\tiny(\emph{community})}
 	  \begin{itemize}
@@
 	  \end{itemize}			
-        \item 1 GPU node
+        \item 1 GPU node {\tiny(\emph{community})}
           \begin{itemize}
@@
 	  \end{itemize}
-        \item 91 Owner nodes
+        \item 96 Owner nodes
           \begin{itemize}
@@
             \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
-            \item 58 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
+            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
+            \item 1 GPU node
             }

### 2017-02-04 hpc_101/paper.tex
Breakup of sections into tex file

@@
 
-\title[HPC--101]{HPC--101\\Onboarding} % The short title appears at the bottom of every slide, the full title is only on the title page
-
-\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
-\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
-{
-Information Technology Services \\
-{\ci} \\
-University of {\hawaii} \\ % Your institution for the title page
-\medskip
-\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
-\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
-}
-\date{\today} % Date, can be changed to a custom date
+\include{sections/00_title}
 
@@
 
-\section{Introduction}
-
-\subsection{Parallel Computing}
-\begin{frame}
-	\frametitle{Parallel Computing}
-	\begin{itemize}
-		\item High Performance Compute
-		\begin{itemize}
-			\item Each separate process can send and receive data amongst other processes (MPI \& OpenMP)
-			\item If processes must communicate for the overall program to proceed, high-speed networking is needed (only MPI)
-		\end{itemize}
-		\item High Throughput Compute
-		\begin{itemize}
-			\item \emph{Pleasantly Parallel} -- Processes are independent and no communication is necessary 
-		\end{itemize}
-	\end{itemize}
-\end{frame}
-
-\section{University of Hawai'i Cluster}
-
-\subsection{{\craycs}}
-\begin{frame}
-\frametitle{{\craycs} -- History}
-\begin{itemize}
- \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
- \item Delivered in October of 2014
- \item Accepted in December of 2014 
- \item Early adopter testing started in January 2015
- \item Opened for general use in April 2015
- \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
- \item January 2017, 5 new nodes added to cluster, raising total core count from 5,876 to 5,992 cores
- \item As of Nov 2016, more than 300 users have been granted access to the cluster
-\end{itemize}
-\end{frame}
-
-\subsubsection{Hardware}
-\begin{frame}
-	\frametitle{{\craycs} -- Compute Nodes}
-	\begin{itemize}
-        \item All nodes are diskless with some RAM used for the Operating System
-	\item 178 Standard nodes {\tiny(\emph{community})}
-	  \begin{itemize}
-            {\footnotesize
-	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
-            }
-	  \end{itemize}          
-	\item 6 Large memory nodes {\tiny(\emph{community})}
-	  \begin{itemize}
-            {\footnotesize
-	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
-	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
-            }
-	  \end{itemize}			
-        \item 1 GPU node {\tiny(\emph{community})}
-          \begin{itemize}
-            {\footnotesize
-	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
-            \item 2 Nvidia Tesla K40 GPUs
-            }
-	  \end{itemize}
-        \item 96 Owner nodes
-          \begin{itemize}
-            {\footnotesize
-            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
-            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
-            \item 1 GPU node
-            }
-          \end{itemize}
-	\item CentOS Linux
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{{\craycs} -- Storage}
-	Two storage options are currently available on the {\craycs}
-	\begin{enumerate}
-		\item {\lustre}
-                  \begin{itemize}
-                  \item High performance parallel filesystem 
-                  \item Available on all the cluster nodes
-                    \item Freely available to all users
-                  \end{itemize}
-		\item ValueStorage
-                  \begin{itemize}
-                  \item Network attached scale out storage
-                  \item Only available from the login nodes
-                  \item Available as a for fee service
-                  \end{itemize}
-	\end{enumerate}
-\end{frame}
-
-\begin{frame}
-	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
-	\begin{itemize}
-		\item The {\craycs} has $\approx582$TB of storage space
-
-		\item Primarily used as scratch space for jobs (Input and Output)
-		\item User do not have a usage quota (soft or hard)
-		\item Certain directories are subject to a 90 day purge policy
-		\item \textbf{Data is not backed up!  Users are responsible for their own data}
-                \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
-        ~\\
-	\begin{itemize}
-		\item $500$TB of scale out storage
-		\item Purchased annually in 0.5TB increments 
-	\end{itemize}
-	\btVFill
-
-	\begin{center}
-	\textbf{ValueStorage Pricing}~\\ 
-		\resizebox{0.50\textwidth}{!}{%
-		\begin{tabular}{l r }
-			\toprule
-			\textbf{Product} & \textbf{Annual Cost}~~ \\
-	                \midrule
-			\midrule
-			0.5TB & \$65.00\\
-                        0.5TB + Replication  & \$130.00 \\
-			\bottomrule
-		\end{tabular}
-		}
-		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
-		~\\	{\footnotesize \emph{All prices are subject to change}}
-	\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{{\craycs} -- Network}
-	\begin{itemize}
-		\item 40Gb Infiniband inter-connects (QDR)
-		\begin{itemize}
-			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
-                        \item low latency ($\approx1.3\mu$s)
-			\item Utilizes the \emph{fat tree network topology}
-			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
-		\end{itemize}
-		\item 10Gb login node internet connectivity
-		\begin{itemize}
-			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
-		\end{itemize}	
-	\end{itemize}
-\end{frame}
-
-\subsubsection{Layout}
-\begin{frame}
-	\frametitle{{\craycs} -- Layout}
-	\includegraphics[width=1\textwidth]{images/layout}
-\end{frame}
-
-\section{Sustainability}
-\subsection{Community Resources}
-\begin{frame}
-	\frametitle{Community Resources}
-	\begin{itemize}
-		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
-		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
-		\item For some users, the publicly available resources may not be enough~\ldots
-	\end{itemize}
-\end{frame}
-
-\subsection{Ownership Models}
-
-\subsubsection{Condo Model}
-\begin{frame}
-	\frametitle{Condo Model}
-	\begin{itemize}
-		\item The Condo model allows users to buy nodes (\emph{condos}) to incorporate into the cluster	
-		\item Node owners are provided with priority access to their purchased hardware
-		\item All nodes have a \emph{5 year warranty}
-		\begin{itemize}
-			\item Once a nodes warranty has expired, it will be removed from the cluster
-		\end{itemize}
-		\item Condo owners are given early access to purchased nodes
-		  \begin{itemize}
-                    \item Access is granted as soon as the funds land in our accounts
-		    \item The 5 year warranty will not begin until newly ordered nodes are installed
-		\end{itemize}
-		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
-	\end{itemize}
-\end{frame}
-
-\subsubsection{Service Units}
-\begin{frame}	
-	\frametitle{Service Units}
-	\begin{itemize}
-	 \item In some cases, users will not require owning a node, but still need priority access
-	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
-	 \item SUs come in two flavors:
-		 \begin{itemize}
-			 \item[--] Standard node units 
-				\begin{itemize}
-					\item 20 core hours, with access to 128GB of ram on a standard~node
-				\end{itemize}
-			 \item[--] Large memory node units
-				\begin{itemize}
-					\item 40 core hours, with access to 1TB of ram on a large~memory~node
-				\end{itemize}
-		 \end{itemize}
-	 \item A minimum order totaling \$500 is required
-	\end{itemize}
-\end{frame}
-
-\subsubsection{Price Card}
-\begin{frame}
-	\frametitle{Price Card}
-	\begin{center}
-	{\large \textbf{Condo Price Card}} \\
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l r ||l r}
-			\toprule
-			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
-			\midrule
-			\midrule
-			Standard & \$6,600.00  & Standard + 2 GPUs (Nvidia{\regtrademark} K40) & \$13,600.00 \\
-			\hline
-			Large Memory & \$33,900.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
-			\bottomrule
-		\end{tabular}
-		}
-	\end{center}
-	\smallskip
-	\begin{center}
-		{\large \textbf{Service Unit Price Card}}\\
-		\begin{tabular}{l || c || r}
-			\toprule
-			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
-			\midrule
-			\midrule
-			Standard node & \$0.50 per SU &  1,000~SU (\$500.00)\\
-			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
-			\bottomrule
-		\end{tabular}
-		\btVFill
-	{\footnotesize \emph{All prices are subject to change}}
-	\end{center}
-\end{frame}
-
-\section{Cluster Interaction}
-
-\begin{frame}
-	\frametitle{Overview -- Cluster Interaction}
-	\begin{itemize}
-		\item{Connecting to a cluster @ UH}
-			\begin{itemize}
-			\item Login to the cluster
-			\item Verify user permissions
-			\end{itemize}
-    {\semitransp[25]{\item User directories
-		\item Transferring files
-		\begin{itemize}{\semitransp[25]{
-			\item Globus
-		}}\end{itemize}
-		\item Software
-		\begin{itemize}{\semitransp[25]{
-			\item Modules
-			\item Acquiring software
-			\item Compilers
-			}}
-		\end{itemize}
-		\item Managing user jobs
-		\begin{itemize}{\semitransp[25]{
-			\item Job scheduler
-			\item Using SLURM
-			\item Partitions			
-			\item Submitting jobs (Examples)
-			}}
-		\end{itemize}			
-		}}
-	\end{itemize}
-\end{frame}
-
-\subsection{Connecting to a cluster @ UH}
-\begin{frame}
-	\frametitle{Connecting to a cluster @ UH}
-	\begin{description}[\setlength{\leftmargini}{0pt}]
-	\item[] \begin{itemize}
-		\item To connect to the cluster, an SSH client is required
-		\item Linux and MacOS X, typically have a SSH client already installed
-		\item Windows typically needs a 3rd party SSH client
-		\item Suggested SSH clients for Windows include:
-		\begin{itemize}
-			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
-			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
-                        \item \href{http://mobaxterm.mobatek.net/}{MobaXterm}
-		\end{itemize}
-	\end{itemize}
-	\item[] ~\\
-	\item[] \begin{itemize}
-		\item The {\craycs} has two login nodes:
-		\begin{itemize}
-			\item uhhpc1.its.hawaii.edu
-			\item uhhpc2.its.hawaii.edu
-		\end{itemize}
-	\end{itemize}
-	\end{description}
-        ~\\~\\
-	{\large Let's attempt to login!}
-\end{frame}
-
-\subsubsection{Login to the {\craycs}}
-\begin{frame}
-\frametitle{Login to the {\craycs}}
-	\begin{center}\textbf{Windows}\end{center}
-	\hrule~\\
-	\begin{itemize}
-		\item If SSH 3.2.9 installed (Lab PCs have it installed)
-		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
-		\item Click ``Quick Connect'' and enter the following information:
-			\begin{itemize}
-			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
-			\item[] \textbf{User Name:} Your UH User name e.g., user99
-			\item[] \textbf{Port:} 22
-			\end{itemize}
-		\item Press ``Connect''
-		\item Enter your UH user password when prompted and press the return key
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Login to the {\craycs}}
-	\begin{center}\textbf{Mac \& Linux}\end{center}
-	\hrule~\\
-	\begin{itemize}
-		\item Open a terminal window
-		\item Enter one of the following:
-		\begin{itemize}
-			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
-			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
-			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
-		\end{itemize}
-		\item Enter your UH user password when prompted and press the return key
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{On Initial Login~\ldots}
-	Validate that all system permissions are correct for your user~\\
-	\begin{enumerate}
-		\item Test that you can list files in your home: `\textcolor{blue}{\texttt{ls -la}}'    
-		\item Test making a file in your home: `\textcolor{blue}{\texttt{touch test.txt}}'    
-		\item Go into \ctilde{}/lus: `\textcolor{blue}{\texttt{cd \ctilde{}/lus/}}'  
-		\item Test making a file in your lus directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
-		\item Go into \ctilde{}/apps:  `\textcolor{blue}{\texttt{cd \ctilde{}/apps/}}'     
-		\item Test making a file in your apps directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
-	\end{enumerate}
-	\begin{alertblock}{Result}
-		\begin{center}Did you get any errors?  Let us know if you did\end{center}
-	\end{alertblock}
-	\btVFill
-
-	\small Notes:
-		\begin{itemize}\tiny
-		\item On login you are placed in /home/$<$username$>$/
-		\item By default, \ctilde{} is equivalent to /home/$<$username$>$/
-		\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Overview -- Cluster Interaction}
-	\begin{itemize}
-		\item {\semitransp[25]{Connecting to a cluster @ UH}}
-			\begin{itemize}{\semitransp[25]{
-			\item Login to the cluster
-			\item Verify user permissions
-			}}
-			\end{itemize}
-    \item User directories
-		{\semitransp[25]{\item Transferring files
-		\begin{itemize}{\semitransp[25]{
-			\item Globus
-		}}\end{itemize}
-		\item Software
-		\begin{itemize}{\semitransp[25]{
-			\item Modules
-			\item Acquiring software
-			\item Compilers
-			}}
-		\end{itemize}
-		\item Managing user jobs
-		\begin{itemize}{\semitransp[25]{
-			\item Job scheduler
-			\item Using SLURM
-			\item Partitions			
-			\item Submitting jobs (Examples)
-			}}
-		\end{itemize}			
-		}}
-	\end{itemize}
-\end{frame}
-
-\subsection{User Directories}
-\begin{frame}[fragile]
-\frametitle{User Directories}
-\begin{block}{Home}
-\begin{semiverbatim}\tiny \texttt
-[user99@login \ctilde]\$ ls -l 
-total 0
-lrwxrwxrwx 1 user99 user99 23 Jan 15 20:38 \textcolor{teal}{apps} -> /lus/scratch/usr/user99
-lrwxrwxrwx 1 user99 user99 19 Jan 15 20:38 \textcolor{teal}{lus} -> /lus/scratch/user99
-lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
-\end{semiverbatim}
-\end{block}
-	\begin{itemize}\footnotesize
-		\item \ctilde{}/ is not on the {\lustre} filesystem and \textbf{should not be used for job data!}
-		\item \ctilde{}/lus/ is a symlink to the {\lustre} scratch
-		\begin{itemize}\tiny
-			\item This is where all your job data files should live
-			\item Items in this directory \emph{\textbf{are}} subject to our 90 day purge policy
-		\end{itemize}
-		\item \ctilde{}/apps/ is a symlink to where programs should be stored
-		\begin{itemize}\tiny
-			\item Items in this directory \emph{\textbf{are not}} subject to our 90 day purge policy
-			\item Directory is monitored for abuse
-		\end{itemize}
-		\item \ctilde{}/purge/ is typically a dead symlink
-		\begin{itemize}\tiny
-			\item Symlink becomes active if the user has files that are part of the next automatic purge
-			\item When \ctilde{}/purge/ is active, the directory containing two files -- \emph{purge\_list.txt} \& \emph{totals.txt}
-			\item An email is sent to users if they have files that will be purged
-			\item Email notification is sent out 14 days before the purge takes place
-		\end{itemize}
-	\end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]
-\frametitle{User Directories}
-\begin{block}{Filesystems}
-\begin{semiverbatim}\tiny \texttt
-[user99@login \ctilde]\$ df -h
-Filesystem                                  Size  Used Avail Use% Mounted on
-10.10.0.3:/ha_cluster/home                  1.8T  888G  851G  52% /home
-10.12.0.51@o2ib:10.12.0.52@o2ib:/scratch    582T  429T  125T  78% /lus/scratch
-\end{semiverbatim}
-\end{block}
-	\begin{itemize}\footnotesize
-		\item /home/$<$username$>$ exists on a NFS mounted filesystem
-		\begin{itemize}\tiny
-			\item Only has 1.8TB of useable space
-			\item Using all this space may cause problems for the entire cluster
-			\item Not a high performance filesystem and small in size
-		\end{itemize}
-		\item /lus/scratch/ is the {\lustre} filesystem
-		\begin{itemize}\tiny
-			\item Has 582TB of useable space
-			\item \ctilde{}/apps/, \ctilde{}/lus/, and \ctilde{}/purge/ all point to directories on this filesystem
-			\item High performance and a lot more space for users to use
-			\item No hard or soft quotas are in place
-			\item Utilization is managed through the 90 day purge policy
-		\end{itemize}
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Overview -- Cluster Interaction}
-	\begin{itemize}
-		\item {\semitransp[25]{Connecting to a cluster @ UH}}
-			\begin{itemize}{\semitransp[25]{
-			\item Login to the cluster
-			\item Verify user permissions
-			}}
-			\end{itemize}
-    {\semitransp[25]{\item User directories}}
-		\item Transferring files
-		\begin{itemize}
-			\item Globus
-		\end{itemize}
-		{\semitransp[25]{
-		\item Software
-		\begin{itemize}{\semitransp[25]{
-			\item Modules
-			\item Acquiring software
-			\item Compilers
-			}}
-		\end{itemize}
-		\item Managing user jobs
-		\begin{itemize}{\semitransp[25]{
-			\item Job scheduler
-			\item Using SLURM
-			\item Partitions			
-			\item Submitting jobs (Examples)
-			}}
-		\end{itemize}			
-		}}
-	\end{itemize}
-\end{frame}
-
-\subsection{Transferring Files}
-\begin{frame}
-	\frametitle{Available File Transfer Protocols}
-	\begin{itemize}
-		\item The cluster has the following options for transferring files:
-		\begin{itemize}
-			\item[--] scp (RCP$+$SSH protocol)
-			\item[--] rsync (rsync protocol with SSH transport)
-			\item[--] SFTP (SSH FTP protocol) -- Filezilla, Cyberduck
-			\item[--] \href{https://www.globus.org/}{Globus} (Grid FTP protocol)
-		\end{itemize}
-		\item All options are widely used, and have clients that can be found for on most major operating systems
-%               \item Please see the the CI website for links to several different clients?                 
-	\end{itemize}
-	\btVFill
-	\begin{center}
-		SFTP, scp, and rsync are fairly common on Linux systems,~\\but Globus is not as common~\ldots
-	\end{center}
-
-\end{frame}
-
-\subsection{Globus}
-\begin{frame}
-	\frametitle{Globus}
-	\begin{block}{What is Globus?}\footnotesize
-	The Globus transfer service provides high-performance, secure, file transfer and synchronization between endpoints.~\\~\\
-	Globus handles all the difficult aspects of data transfer, allowing application users to easily start and manage transfers between endpoints, while automatically tuning parameters to maximize bandwidth usage, managing security configurations, providing automatic fault recovery, and notifying users of completion and problems.
-	\end{block}
-	\begin{definition}\tiny
-	An \textbf{\emph{endpoint}} is one of the two file transfer locations -- either the source or the destination -- between which files can move.~\\Once a resource (such as a server, cluster, storage system, laptop, or other system) is defined as an endpoint, it will be available to authorized users who can transfer files to or from this endpoint.
-	\end{definition}
-\numlessfootnotetxt{\tiny \url{https://www.globus.org/file-transfer}}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Globus}\footnotesize
-\begin{block}{How do I get Globus?}
-To utilize Globus, follow the following steps:
-\begin{itemize}\tiny
-\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
-\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
-\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
-\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
-\begin{itemize}\tiny
- \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
-\end{itemize}
-\item Once you select an endpoint, a login window will pop up.
-You can access the UHHPC endpoints by simply using your UH username and password.
-Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’.
-\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
-\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
-\end{itemize}
-\end{block}
-~\\
-These instructions can also be found on the {\ci} website:~\\
-{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Globus}
-        \begin{itemize}
-          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
-          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
-          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
-            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
-\end{itemize} 
-\end{frame}
-
-\begin{frame}
-	\frametitle{Overview -- Cluster Interaction}
-	\begin{itemize}
-		\item {\semitransp[25]{Connecting to a cluster @ UH}}
-			\begin{itemize}{\semitransp[25]{
-			\item Login to the cluster
-			\item Verify user permissions
-			}}
-			\end{itemize}
-   {\semitransp[25]{ \item User directories
-		\item Transferring files
-		\begin{itemize}{\semitransp[25]{
-			\item Globus
-		}}\end{itemize}
-		}}
-		\item Software
-		\begin{itemize}
-			\item Modules
-			\item Acquiring software
-			\item Compilers
-		\end{itemize}
-		{\semitransp[25]{
-		\item Managing user jobs
-		\begin{itemize}{\semitransp[25]{
-			\item Job scheduler
-			\item Using SLURM
-			\item Partitions
-			\item Submitting jobs (Examples)
-			}}
-		\end{itemize}			
-		}}
-	\end{itemize}
-\end{frame}
-
-\subsection{Software}
-\begin{frame}
-	\frametitle{Modules}
-	\begin{block}{Modules}\tiny
-	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark
-
-	\end{block}
-	\begin{itemize}
-		\item We globally install frequently requested software packages and create modules for all users to access
-		\item Access to modules is via the \textbf{module} command
-		\begin{itemize}\footnotesize
-			\item `module avail' -- list installed modules
-			\item `module show $<$module name$>$' -- Show what actions a module performs
-			\item `module load $<$module name$>$' -- Loads the named module
-			\item `module purge' -- Unload all loaded modules
-		\end{itemize}
-		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
-	\end{itemize}
-	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Acquiring Software -- Binaries and/or Source}
-	\begin{itemize}
-		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the {\craycs}
-		\begin{itemize}
-			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
-		\end{itemize}
-		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
-		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Compilers}
-	\begin{itemize}
-		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
-		\item Compiling must take place on a compute node
-		\begin{itemize}
-			\item Interactive sessions are useful for compiling software
-			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
-			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
-		\end{itemize}
-		\item {\intel} compilers are recommended for best performance
-		\begin{itemize}
-			\item {\intel} 2013 compilers:
-			\begin{itemize}
-				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
-				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
-			\end{itemize}
-			\item {\intel} 2016 compilers:
-			\begin{itemize}
-				\item We have 2 floating seats for {\intel} 2016 compiler
-				\item intel\_2016/ics 
-				\item intel\_2016/impi 
-			\end{itemize}
-		\end{itemize}			
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Overview -- Cluster Interaction}
-	\begin{itemize}
-		\item {\semitransp[25]{Connecting to a cluster @ UH}}
-			\begin{itemize}{\semitransp[25]{
-			\item Login to the cluster
-			\item Verify user permissions
-			}}
-			\end{itemize}
-   {\semitransp[25]{ \item User directories
-		\item Transferring files
-		\begin{itemize}{\semitransp[25]{
-			\item Globus
-		}}\end{itemize}
-		\item Software
-		\begin{itemize}{\semitransp[25]{
-			\item Modules
-			\item Acquiring software
-			\item Compilers
-			}}
-		\end{itemize}
-		}}
-		\item Managing user jobs
-		\begin{itemize}
-			\item Job scheduler
-			\item Using SLURM
-			\item Partitions
-			\item Submitting jobs (Examples)
-		\end{itemize}		
-	\end{itemize}
-\end{frame}
-
-\subsection{Managing User Jobs}
-\begin{frame}
-\frametitle{Managing User Jobs}
-User jobs all come in different shapes and sizes:  
-\begin{itemize}\footnotesize
-	\item Require multiple nodes working in concert towards a common goal (MPI)
-	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
-	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
-\end{itemize}
-\bigskip
-The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we impose order on this chaos?
-\btVFill
-\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Job Schedulers}
-  \begin{block}{Purpose}To control and prioritize the execution order of unrelated jobs\end{block}
-	Basic features expected of a job scheduler:
-	\begin{itemize}\footnotesize
-        \item Provides a user interface for users to request resources and monitor work 
-        \item Allocates access to resources for the requested duration of time
-	\item Starts, monitors and terminates work on allocated resources
-	\item Arbitrates contention for resources by managing queues of pending work
-	\end{itemize}
-	\bigskip
-        The {\craycs} uses the \textbf{S}imple \textbf{L}inux \textbf{U}tility for \textbf{R}esource \textbf{M}anagement~\\ or simply known as the \emph{SLURM scheduler}
-	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
-	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
-\end{frame}
-
-\subsection{SLURM}
-\begin{frame}
-\frametitle{SLURM}
-\begin{block}{How are jobs scheduled?}
-User submitted jobs are assigned a priority using a fairshare algorithm.\\
-Factors such as the following are all used to assign a priority to a given job: 
-\begin{itemize}
-\item Runtime
-\item Amount of resources requested
-\item Age of job
-\item Amount of core hours a user has used in recent history
-\end{itemize}
-
-\end{block}
-\end{frame}
-
-\begin{frame}
-\frametitle{SLURM commands}\footnotesize
-SLURM has a series of commands, each of which allow users to interact with the job scheduler
-\begin{itemize}\tiny
-	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
-	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
-	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
-	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
-	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
-	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
-	\item \emph{\textbf{scontrol}} -- The administrative tool used to view and/or modify Slurm state. Note that many scontrol commands can only be executed as user root
-\end{itemize}
-\begin{itemize}
-	\item[--] Examples usage of the SLURM commands can be seen on schedmd's \href{http://slurm.schedmd.com/quickstart.html}{quickstart}
-	\item[--] Each command should have a `man' page, or displays help when the -h flag is used
-\end{itemize}
-\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
-\end{frame}
-
-\subsection{SLURM Partitions}
-\begin{frame}
-\footnotesize
-\frametitle{Partitions}
-\begin{block}{What is a partition?}\footnotesize
-A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.
-Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc.
-Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
-\end{block}
-\begin{itemize}
-	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
-	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
-\end{itemize}
-\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
-\end{frame}
-
-\begin{frame}
-\frametitle{Partitions}
-\includegraphics[width=0.95\textwidth]{images/partitions}
-\end{frame}
-
-\begin{frame}
-\frametitle{Partition Details}
-\resizebox{0.95\textwidth}{!}{%
-\begin{tabular}{l || c || c || c || c || c || c}
-\toprule                                                                    
-\textbf{Partition} & \textbf{Time} & \textbf{Nodes per job} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt Mode} & \textbf{Memory per CPU (MB)} \\
-%\toprule                                                                    
-%\toprule                                                                    
-\midrule
-\midrule
-community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-exclusive.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:}& \padfrom{\ \ \ }{1}~\\ \textbf{Max:}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-gpu.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-htc.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-lm.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\hline
-\hline
-sb.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:05:00~\\ \textbf{Max:} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
-\bottomrule 
-\end{tabular}   
-}
-\\
-\bigskip
-Partition details also available on the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command:
-\\
-\texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
-\end{frame}
-
-\subsection{Interactive Job -- srun}
-\frametitle{SLURM Job Scripts}
-\begin{frame}
-\frametitle{Interactive Job with SLURM}
-	\begin{block}{Interactive session}\tiny
-	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
-          ~\\
-          ~\\
-          \hrule
-          ~\\
-	$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash
-
-	\end{block}
-\btVFill
-\begin{center}Interactive sessions terminate when the specified time has elapsed~\\or if you give the \textbf{exit} command \end{center}
-
-\end{frame}
-
-\subsection{GPU Batch Job Example Script}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File -- GPU}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat gpu.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=GPU\_example
-\textcolor{blue}{\#SBATCH \ddash{}partition=gpu.q}
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=1
-\#SBATCH \ddash{}tasks-per-node=1
-\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=10}
-\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
-\textcolor{blue}{\#SBATCH \ddash{}gres=gpu:NV-K40:2 } \#\# request both GPUs in the GPU node
-\#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-module load GPGPU/cuda/samples/7.5
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-
-bandwidthTest; cudaOpenMP; deviceQuery; simpleMultiGPU
-\end{semiverbatim}
-\end{frame}
-
-\subsection{MPI Batch Job Example Script}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat mpi.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=MPI\_example
-\textcolor{blue}{\#SBATCH \ddash{}partition=exclusive.q}
-\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\textcolor{blue}{\#SBATCH \ddash{}nodes=4}
-\textcolor{blue}{\#SBATCH \ddash{}tasks-per-node=20}
-\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=1}
-\textcolor{blue}{\#\#SBATCH \ddash{}mem=11000} \# Memory should not be set for jobs in exclusive.q
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-\#\# Intel MPI manual: \href{https://software.intel.com/en-us/mpi-refman-lin-html}{https://software.intel.com/en-us/mpi-refman-lin-html}
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-export I\_MPI\_FABRICS=shm:tmi  
-export I\_MPI\_PMI\_LIBRARY=/opt/local/slurm/default/lib64/libpmi.so
-
-srun  -n \$\{SLURM\_NTASKS\}  ./hello\_mpi.intel 
-\end{semiverbatim}
-\end{frame}
-
-\subsection{Non-MPI Batch Job Example Script}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat hello_world.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=example
-\textcolor{blue}{\#SBATCH \ddash{}partition=community.q}
-\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=1
-\#SBATCH \ddash{}tasks-per-node=1
-\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=5}
-\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
-\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
-\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-
-./hello\_world
-\end{semiverbatim}
-\end{frame}
-
-\subsection{Executing and Monitoring Batch Jobs}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
-To execute a submission script you use the \textbf{\texttt{sbatch}} command
-\begin{block}{Example}
-\begin{semiverbatim}\tiny
-[login lus]\$ sinfo
-PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
-community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
-
-[login lus]\$ sbatch hello_world.slurm
-sbatch: Submitted batch job 469
-
-[login lus]\$ squeue
-JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
-469   community.q  example  user99  R  00:01 1      compute-0001
-\end{semiverbatim}
-\end{block}
-
-\btVFill
-\begin{center}What if I want to run many jobs with identical parameters, but on different inputs?~\\
-  SLURM provides a mechanism for this called a \textbf{Job Array}.\end{center}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Slurm Job Array}
-\begin{center}  
-  \includegraphics[width=0.90\textwidth]{images/job_array}
-\end{center}
-\end{frame}
-
-\subsection{Batch Job Example Script Using Job Array}
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
-\begin{semiverbatim}\tiny
-[login lus]\$ cat job_array.slurm
-
-\#!/bin/bash
-\#SBATCH \ddash{}job-name=example
-\#SBATCH \ddash{}partition=community.q
-\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
-\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
-\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
-\#SBATCH \ddash{}nodes=1
-\#SBATCH \ddash{}tasks-per-node=1
-\#SBATCH \ddash{}cpus-per-task=5
-\#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
-\textcolor{blue}{\#SBATCH \ddash{}error=myprog-\%A\_\%a.err} \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\textcolor{blue}{\#SBATCH \ddash{}output=myprog-\%A\_\%a.out} \#\# \%A - filled with jobid. \%a - filled with job arrayid
-\#\# Useful for remote notification
-\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
-\#SBATCH \ddash{}mail-user=user@test.org
-
-source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
-
-\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
-
-export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
-
-./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.dat -o output\_\$\{SLURM_ARRAY_TASK_ID\}
-\end{semiverbatim}
-\end{frame}
-
-\subsection{Executing and Monitoring Job Array Batch Jobs}
-
-\begin{frame}[fragile]
-\frametitle{SLURM sbatch -- Executing \& Monitoring Job Arrays}\footnotesize
-\begin{block}{Example}
-\begin{semiverbatim}\tiny
-[login lus]\$ sinfo
-PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
-community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
-
-[login lus]\$ sbatch {\ddash}array=1-30 job_array.slurm
-sbatch: Submitted batch job 470
-
-[login lus]\$ squeue
-JOBID      PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
-470[3-30]  community.q  example  user99  PD   00:00   1      (resource)
-470_1      community.q  example  user99  R    00:01   1      compute-0001
-470_2      community.q  example  user99  R    00:05   1      compute-0002
-\end{semiverbatim}
-\end{block}
-\end{frame}
-
-\section{Problem Reporting}
-\begin{frame}
-\frametitle{Problem Reporting}
-\begin{itemize}
-\item Always email UH-HPC-Help@lists.hawaii.edu
-\item For batch jobs \ldots
-\begin{itemize}
-\item[--] Job ID, path to submission script, submission command, error file location, output files
-\end{itemize} 
-\item For other problems \ldots
-\begin{itemize}
-\item[--] State the problem, command issued, host, directory, remote host, error messages
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\section{Cluster Etiquette}
-\begin{frame}
-\frametitle{Cluster Etiquette}
-\begin{itemize}
-\item Never run anything on the login nodes
-\item Try to request only the resources that your job needs.  Leaving unneeded resources available, allows other users to use them
-\item Test your application in an interactive session before scheduling a long running job
-\begin{itemize}
-\item[--] This can help identify what resources your job will require, and potentially aid you in correctly sizing what resources you request.
-\end{itemize}
-\item Use the nodes in the sb.q to validate your slurm script runs as intended
-\begin{itemize}
-\item[--] This helps to reduce the chance of you having a small error that kills your job prematurely after waiting for resources 
-\end{itemize} 
-\item Always run jobs from within {\ctilde}/lus/ directory to ensure output is written there and does not fill up the system filesystem
-\end{itemize}
-\end{frame}
-
-\section{Policies}
-%% \begin{frame}
-%% 	\frametitle{Overview}
-%% 	\begin{itemize}
-%% 		\item Login node usage policy
-%% 		\item Scratch filesystem purge policy
-%% 	\end{itemize}
-%% \end{frame}
-
-\subsection{Login Node Usage Policy}
-\begin{frame}
-\frametitle{Login Node Usage Policy}\footnotesize
-The login nodes are a shared resource and are the only access to the cluster for hundreds of user and are meant to provide the following functionality for all users: 
-\begin{itemize}
-\item Providing ssh shell access 
-\item Facilitate the transfer files to and from the cluster:\\Globus, sftp, scp, rsync
-\item Launching and monitoring SLURM jobs (batch and interactive)
-\item Modifying text files with a text editor: vi/vim, emacs, nano
-\end{itemize}
-\bigskip
-\begin{itemize}
-\item[--] If we identify or are notified that a user is running computation on the login nodes, the application will be killed
-\item[--] If we determine that a user repeatedly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
-\end{itemize}
-\btVFill
-\begin{center}
-\footnotesize \textbf{\emph{Login node usage policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
-\end{center}
-\end{frame}
-
-\subsection{Lustre Filesystem Purge Policies}
-\begin{frame}
-\frametitle{{\lustre} Filesystem Purge Policies}
-Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.
-To accomplish this, we have currently implemented a purge policy, for any file that is older than 90 days.
-\begin{itemize}
- \item Which of my directories are subject to the purge policy?
-   \begin{itemize}
-   \item \textbf{Answer:} All files and folders found in \ctilde/lus/ are subject to the 90 day purge policy.  Symlinks are excluded from the purge, and are not followed by the purge bot.  
-   \end{itemize}
-\item How frequently are files for purging identified?
-   \begin{itemize}
-   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
-   \end{itemize}
-\end{itemize}
-\btVFill
-\begin{center}
-\footnotesize \textbf{\emph{Purge policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
-\end{center}
-\end{frame}
-
-\begin{frame}
-\Huge{\centerline{Questions?}}
-\end{frame}
-
-\section{Related Services}
-\begin{frame}
-\frametitle{Services}
-\begin{itemize}
-\item MATLAB Distributed Computing Server{\trademark}
-  \begin{itemize}
-    \item Requires user to already have a license for MATLAB
-    \item User must also have access to at least the Parallel Computing Toolbox{\trademark}
-    \item Provides users access to up to 64 workers
-    \item Requires some additional configuration
-    \item Email us to request access to this additional service
-  \end{itemize}
-\end{itemize}
-\end{frame}
-
-\section{Frequently Asked Questions}
-\begin{frame}
-\frametitle{FAQ}
-\begin{itemize}\footnotesize
-\item Are files on the cluster backed up?
-  \begin{itemize}\tiny
-  \item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.
-It is up to you as the user to validate and maintain your own backups.
-The {\citeam} takes no responsibility for any data that is lost due to human or mechanical error.
-  \end{itemize}
-
-\item Help I have been trying to compile an application without success, can the {\citeam} help?
-  \begin{itemize}\tiny
-  \item \textbf{Answer:} We are willing to help, just send us an email at \textit{uh-hpc-help@lists.hawaii.edu} with a description of what you have attempted as well as a link to the software you are trying to compile, and/or where it is located on the cluster.
-  \end{itemize}
-
-\item HELP! My job needs more time and does not support check-pointing.  What can I do?
-  \begin{itemize}\tiny
-  \item \textbf{Answer:} We understand that jobs may not all fit within the 3 day timelimit or unforeseen issues during the execution can cause a job to run long.
-The {\citeam} does review and potentially approves requests to extend jobs in a one off fashion.
-A request to extend a job can be denied, since we need to take into account factor such as: how busy the cluster is, the number of time extension requests a user has made in recent history.\\If it is known before the job is submitted that more than 3 days is required, we would ask that you contact us first so we can verify that you truly will need more time.\\If a job is already running and just doesn't look like it will complete in time, please send us a request with the jobid number and what you want the job extended to e.g., please extend my jobs run time to 7 days.\\Please remember the {\citeam} is small and we are also human.
-Depending on when you make your request we may not be able to immediately respond or act.
-It is best to make your request at the first signs you need more time, in hopes that you can provide us with ample time to act.
-  \end{itemize}
-
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}	
-\frametitle{FAQ}
-\begin{itemize}
-  \item Can I run a display program on the cluster?
-    \begin{itemize}
-      \item \textbf{Answer:} Yes, but not from the login nodes.  To follow cluster usage policies, please run X11 applications from a compute node.  Please see the following steps.
-\end{itemize}
-\end{itemize}
- \begin{block}{Interactive session with X11}\tiny  
-       \begin{enumerate}
-      \item Connect via SSH using the -Y option, X11 forwarding enabled
-      \item run srun.x11 to start a session on a node
-      \end{enumerate}
-      \
-      ~\\
-      ~\\
-      $[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
-      $[$login \ctilde$]$\$ srun.x11 \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
-      $[$compute-0001 \ctilde$]$\$ xterm	~\\
- \end{block}
-
-\end{frame}
+\include{sections/01_intro}
+\include{sections/02_cluster}
+%\include{sections/03_sustainability}
+\include{sections/04_cluster_interaction}
+\include{sections/05_reporting}
+\include{sections/06_ediquette}
+\include{sections/07_policies}
+\include{sections/08_end}
+%\include{sections/09_related_services}
+\include{sections/10_faq}
 

### 2017-02-04 hpc_101/sections/00_title.tex
Breakup of sections into tex file

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[HPC--101]{HPC--101\\Onboarding} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
{\ci} \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

### 2017-02-04 hpc_101/sections/01_intro.tex
Breakup of sections into tex file

\section{Introduction}

\subsection{Parallel Computing}
\begin{frame}
	\frametitle{Parallel Computing}
	\begin{itemize}
		\item High Performance Compute
		\begin{itemize}
			\item Each separate process can send and receive data amongst other processes (MPI \& OpenMP)
			\item If processes must communicate for the overall program to proceed, high-speed networking is needed (only MPI)
		\end{itemize}
		\item High Throughput Compute
		\begin{itemize}
			\item \emph{Pleasantly Parallel} -- Processes are independent and no communication is necessary 
		\end{itemize}
	\end{itemize}
\end{frame}

### 2017-02-04 hpc_101/sections/02_cluster.tex
Breakup of sections into tex file

\section{University of Hawai'i Cluster}

\subsection{{\craycs}}
\begin{frame}
\frametitle{{\craycs} -- History}
\begin{itemize}
 \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
 \item Delivered in October of 2014
 \item Accepted in December of 2014 
 \item Early adopter testing started in January 2015
 \item Opened for general use in April 2015
 \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
 \item January 2017, 5 new nodes added to cluster, raising total core count from 5,876 to 5,992 cores
 \item As of Nov 2016, more than 300 users have been granted access to the cluster
\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs} -- Compute Nodes}
	\begin{itemize}
        \item All nodes are diskless with some RAM used for the Operating System
	\item 178 Standard nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            }
	  \end{itemize}          
	\item 6 Large memory nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
            }
	  \end{itemize}			
        \item 1 GPU node {\tiny(\emph{community})}
          \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            \item 2 Nvidia Tesla K40 GPUs
            }
	  \end{itemize}
        \item 96 Owner nodes
          \begin{itemize}
            {\footnotesize
            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
            \item 1 GPU node
            }
          \end{itemize}
	\item CentOS Linux
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage}
	Two storage options are currently available on the {\craycs}
	\begin{enumerate}
		\item {\lustre}
                  \begin{itemize}
                  \item High performance parallel filesystem 
                  \item Available on all the cluster nodes
                    \item Freely available to all users
                  \end{itemize}
		\item ValueStorage
                  \begin{itemize}
                  \item Network attached scale out storage
                  \item Only available from the login nodes
                  \item Available as a for fee service
                  \end{itemize}
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item The {\craycs} has $\approx582$TB of storage space

		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
                \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
        ~\\
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Purchased annually in 0.5TB increments 
	\end{itemize}
	\btVFill

	\begin{center}
	\textbf{ValueStorage Pricing}~\\ 
		\resizebox{0.50\textwidth}{!}{%
		\begin{tabular}{l r }
			\toprule
			\textbf{Product} & \textbf{Annual Cost}~~ \\
	                \midrule
			\midrule
			0.5TB & \$65.00\\
                        0.5TB + Replication  & \$130.00 \\
			\bottomrule
		\end{tabular}
		}
		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
		~\\	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
                        \item low latency ($\approx1.3\mu$s)
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

\subsubsection{Layout}
\begin{frame}
	\frametitle{{\craycs} -- Layout}
	\includegraphics[width=1\textwidth]{images/layout}
\end{frame}

### 2017-02-04 hpc_101/sections/03_sustainability.tex
Breakup of sections into tex file

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

\subsection{Ownership Models}

\subsubsection{Condo Model}
\begin{frame}
	\frametitle{Condo Model}
	\begin{itemize}
		\item The Condo model allows users to buy nodes (\emph{condos}) to incorporate into the cluster	
		\item Node owners are provided with priority access to their purchased hardware
		\item All nodes have a \emph{5 year warranty}
		\begin{itemize}
			\item Once a nodes warranty has expired, it will be removed from the cluster
		\end{itemize}
		\item Condo owners are given early access to purchased nodes
		  \begin{itemize}
                    \item Access is granted as soon as the funds land in our accounts
		    \item The 5 year warranty will not begin until newly ordered nodes are installed
		\end{itemize}
		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
	\end{itemize}
\end{frame}

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In some cases, users will not require owning a node, but still need priority access
	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
	 \item SUs come in two flavors:
		 \begin{itemize}
			 \item[--] Standard node units 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
\end{frame}

\subsubsection{Price Card}
\begin{frame}
	\frametitle{Price Card}
	\begin{center}
	{\large \textbf{Condo Price Card}} \\
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{l r ||l r}
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
			\midrule
			\midrule
			Standard & \$6,600.00  & Standard + 2 GPUs (Nvidia{\regtrademark} K40) & \$13,600.00 \\
			\hline
			Large Memory & \$33,900.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
			\bottomrule
		\end{tabular}
		}
	\end{center}
	\smallskip
	\begin{center}
		{\large \textbf{Service Unit Price Card}}\\
		\begin{tabular}{l || c || r}
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
			\midrule
			\midrule
			Standard node & \$0.50 per SU &  1,000~SU (\$500.00)\\
			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
			\bottomrule
		\end{tabular}
		\btVFill
	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

### 2017-02-04 hpc_101/sections/04_cluster_interaction.tex
Breakup of sections into tex file

\section{Cluster Interaction}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item{Connecting to a cluster @ UH}
			\begin{itemize}
			\item Login to the cluster
			\item Verify user permissions
			\end{itemize}
    {\semitransp[25]{\item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Connecting to a cluster @ UH}
\begin{frame}
	\frametitle{Connecting to a cluster @ UH}
	\begin{description}[\setlength{\leftmargini}{0pt}]
	\item[] \begin{itemize}
		\item To connect to the cluster, an SSH client is required
		\item Linux and MacOS X, typically have a SSH client already installed
		\item Windows typically needs a 3rd party SSH client
		\item Suggested SSH clients for Windows include:
		\begin{itemize}
			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
                        \item \href{http://mobaxterm.mobatek.net/}{MobaXterm}
		\end{itemize}
	\end{itemize}
	\item[] ~\\
	\item[] \begin{itemize}
		\item The {\craycs} has two login nodes:
		\begin{itemize}
			\item uhhpc1.its.hawaii.edu
			\item uhhpc2.its.hawaii.edu
		\end{itemize}
	\end{itemize}
	\end{description}
        ~\\~\\
	{\large Let's attempt to login!}
\end{frame}

\subsubsection{Login to the {\craycs}}
\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Windows}\end{center}
	\hrule~\\
	\begin{itemize}
		\item If SSH 3.2.9 installed (Lab PCs have it installed)
		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
		\item Click ``Quick Connect'' and enter the following information:
			\begin{itemize}
			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
			\item[] \textbf{User Name:} Your UH User name e.g., user99
			\item[] \textbf{Port:} 22
			\end{itemize}
		\item Press ``Connect''
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Mac \& Linux}\end{center}
	\hrule~\\
	\begin{itemize}
		\item Open a terminal window
		\item Enter one of the following:
		\begin{itemize}
			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
		\end{itemize}
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{On Initial Login~\ldots}
	Validate that all system permissions are correct for your user~\\
	\begin{enumerate}
		\item Test that you can list files in your home: `\textcolor{blue}{\texttt{ls -la}}'    
		\item Test making a file in your home: `\textcolor{blue}{\texttt{touch test.txt}}'    
		\item Go into \ctilde{}/lus: `\textcolor{blue}{\texttt{cd \ctilde{}/lus/}}'  
		\item Test making a file in your lus directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
		\item Go into \ctilde{}/apps:  `\textcolor{blue}{\texttt{cd \ctilde{}/apps/}}'     
		\item Test making a file in your apps directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
	\end{enumerate}
	\begin{alertblock}{Result}
		\begin{center}Did you get any errors?  Let us know if you did\end{center}
	\end{alertblock}
	\btVFill

	\small Notes:
		\begin{itemize}\tiny
		\item On login you are placed in /home/$<$username$>$/
		\item By default, \ctilde{} is equivalent to /home/$<$username$>$/
		\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    \item User directories
		{\semitransp[25]{\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{User Directories}
\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Home}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ ls -l 
total 0
lrwxrwxrwx 1 user99 user99 23 Jan 15 20:38 \textcolor{teal}{apps} -> /lus/scratch/usr/user99
lrwxrwxrwx 1 user99 user99 19 Jan 15 20:38 \textcolor{teal}{lus} -> /lus/scratch/user99
lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item \ctilde{}/ is not on the {\lustre} filesystem and \textbf{should not be used for job data!}
		\item \ctilde{}/lus/ is a symlink to the {\lustre} scratch
		\begin{itemize}\tiny
			\item This is where all your job data files should live
			\item Items in this directory \emph{\textbf{are}} subject to our 90 day purge policy
		\end{itemize}
		\item \ctilde{}/apps/ is a symlink to where programs should be stored
		\begin{itemize}\tiny
			\item Items in this directory \emph{\textbf{are not}} subject to our 90 day purge policy
			\item Directory is monitored for abuse
		\end{itemize}
		\item \ctilde{}/purge/ is typically a dead symlink
		\begin{itemize}\tiny
			\item Symlink becomes active if the user has files that are part of the next automatic purge
			\item When \ctilde{}/purge/ is active, the directory containing two files -- \emph{purge\_list.txt} \& \emph{totals.txt}
			\item An email is sent to users if they have files that will be purged
			\item Email notification is sent out 14 days before the purge takes place
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Filesystems}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ df -h
Filesystem                                  Size  Used Avail Use% Mounted on
10.10.0.3:/ha_cluster/home                  1.8T  888G  851G  52% /home
10.12.0.51@o2ib:10.12.0.52@o2ib:/scratch    582T  429T  125T  78% /lus/scratch
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item /home/$<$username$>$ exists on a NFS mounted filesystem
		\begin{itemize}\tiny
			\item Only has 1.8TB of useable space
			\item Using all this space may cause problems for the entire cluster
			\item Not a high performance filesystem and small in size
		\end{itemize}
		\item /lus/scratch/ is the {\lustre} filesystem
		\begin{itemize}\tiny
			\item Has 582TB of useable space
			\item \ctilde{}/apps/, \ctilde{}/lus/, and \ctilde{}/purge/ all point to directories on this filesystem
			\item High performance and a lot more space for users to use
			\item No hard or soft quotas are in place
			\item Utilization is managed through the 90 day purge policy
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    {\semitransp[25]{\item User directories}}
		\item Transferring files
		\begin{itemize}
			\item Globus
		\end{itemize}
		{\semitransp[25]{
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Transferring Files}
\begin{frame}
	\frametitle{Available File Transfer Protocols}
	\begin{itemize}
		\item The cluster has the following options for transferring files:
		\begin{itemize}
			\item[--] scp (RCP$+$SSH protocol)
			\item[--] rsync (rsync protocol with SSH transport)
			\item[--] SFTP (SSH FTP protocol) -- Filezilla, Cyberduck
			\item[--] \href{https://www.globus.org/}{Globus} (Grid FTP protocol)
		\end{itemize}
		\item All options are widely used, and have clients that can be found for on most major operating systems
%               \item Please see the the CI website for links to several different clients?                 
	\end{itemize}
	\btVFill
	\begin{center}
		SFTP, scp, and rsync are fairly common on Linux systems,~\\but Globus is not as common~\ldots
	\end{center}

\end{frame}

\subsection{Globus}
\begin{frame}
	\frametitle{Globus}
	\begin{block}{What is Globus?}\footnotesize
	The Globus transfer service provides high-performance, secure, file transfer and synchronization between endpoints.~\\~\\
	Globus handles all the difficult aspects of data transfer, allowing application users to easily start and manage transfers between endpoints, while automatically tuning parameters to maximize bandwidth usage, managing security configurations, providing automatic fault recovery, and notifying users of completion and problems. 
	\end{block}
	\begin{definition}\tiny
	An \textbf{\emph{endpoint}} is one of the two file transfer locations -- either the source or the destination -- between which files can move.~\\Once a resource (such as a server, cluster, storage system, laptop, or other system) is defined as an endpoint, it will be available to authorized users who can transfer files to or from this endpoint.
	\end{definition}
\numlessfootnotetxt{\tiny \url{https://www.globus.org/file-transfer}}
\end{frame}

\begin{frame}
	\frametitle{Globus}\footnotesize
\begin{block}{How do I get Globus?}
To utilize Globus, follow the following steps:
\begin{itemize}\tiny
\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
\begin{itemize}\tiny
 \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
\end{itemize}
\item Once you select an endpoint, a login window will pop up. You can access the UHHPC endpoints by simply using your UH username and password. Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’. 
\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
\end{itemize}
\end{block}
~\\
These instructions can also be found on the {\ci} website:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Globus}
        \begin{itemize}
          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
\end{itemize} 
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		}}
		\item Software
		\begin{itemize}
			\item Modules
			\item Acquiring software
			\item Compilers
		\end{itemize}
		{\semitransp[25]{
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Software}
\begin{frame}
	\frametitle{Modules}
	\begin{block}{Modules}\tiny
	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module show $<$module name$>$' -- Show what actions a module performs
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

\begin{frame}
	\frametitle{Acquiring Software -- Binaries and/or Source}
	\begin{itemize}
		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the {\craycs}
		\begin{itemize}
			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
		\end{itemize}
		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Compilers}
	\begin{itemize}
		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
		\item Compiling must take place on a compute node
		\begin{itemize}
			\item Interactive sessions are useful for compiling software
			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
		\end{itemize}
		\item {\intel} compilers are recommended for best performance
		\begin{itemize}
			\item {\intel} 2013 compilers:
			\begin{itemize}
				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
			\end{itemize}
			\item {\intel} 2016 compilers:
			\begin{itemize}
				\item We have 2 floating seats for {\intel} 2016 compiler
				\item intel\_2016/ics 
				\item intel\_2016/impi 
			\end{itemize}
		\end{itemize}			
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		}}
		\item Managing user jobs
		\begin{itemize}
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
		\end{itemize}		
	\end{itemize}
\end{frame}

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}\footnotesize
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
\end{itemize}
\bigskip
The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we impose order on this chaos?
\btVFill
\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
\end{frame}

\begin{frame}
  \frametitle{Job Schedulers}
  \begin{block}{Purpose}To control and prioritize the execution order of unrelated jobs\end{block}
	Basic features expected of a job scheduler:
	\begin{itemize}\footnotesize
        \item Provides a user interface for users to request resources and monitor work 
        \item Allocates access to resources for the requested duration of time
	\item Starts, monitors and terminates work on allocated resources
	\item Arbitrates contention for resources by managing queues of pending work
	\end{itemize}
	\bigskip
        The {\craycs} uses the \textbf{S}imple \textbf{L}inux \textbf{U}tility for \textbf{R}esource \textbf{M}anagement~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.\\
Factors such as the following are all used to assign a priority to a given job: 
\begin{itemize}
\item Runtime
\item Amount of resources requested
\item Age of job
\item Amount of core hours a user has used in recent history
\end{itemize}

\end{block}
\end{frame}

\begin{frame}
\frametitle{SLURM commands}\footnotesize
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}\tiny
	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
	\item \emph{\textbf{scontrol}} -- The administrative tool used to view and/or modify Slurm state. Note that many scontrol commands can only be executed as user root
\end{itemize}
\begin{itemize}
	\item[--] Examples usage of the SLURM commands can be seen on schedmd's \href{http://slurm.schedmd.com/quickstart.html}{quickstart}
	\item[--] Each command should have a `man' page, or displays help when the -h flag is used
\end{itemize}
\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\begin{frame}
\frametitle{Partitions}
\includegraphics[width=0.95\textwidth]{images/partitions}
\end{frame}

\begin{frame}
\frametitle{Partition Details}
\resizebox{0.95\textwidth}{!}{%
\begin{tabular}{l || c || c || c || c || c || c}
\toprule                                                                    
\textbf{Partition} & \textbf{Time} & \textbf{Nodes per job} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt Mode} & \textbf{Memory per CPU (MB)} \\
%\toprule                                                                    
%\toprule                                                                    
\midrule
\midrule
community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
exclusive.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:}& \padfrom{\ \ \ }{1}~\\ \textbf{Max:}&\padfrom{\ \ \ }{20}\end{tabular} & 10 & EXCLUSIVE & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
gpu.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
htc.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 1 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
lm.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\hline
\hline
sb.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:05:00~\\ \textbf{Max:} & 0-01:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{2}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
\bottomrule 
\end{tabular}   
}
\\
\bigskip
Partition details also available on the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command:
\\
\texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
\end{frame}

\subsection{Interactive Job -- srun}
\frametitle{SLURM Job Scripts}
\begin{frame}
\frametitle{Interactive Job with SLURM}
	\begin{block}{Interactive session}\tiny
	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
          ~\\
          ~\\
          \hrule
          ~\\
	$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash

	\end{block}
\btVFill
\begin{center}Interactive sessions terminate when the specified time has elapsed~\\or if you give the \textbf{exit} command \end{center}

\end{frame}

\subsection{GPU Batch Job Example Script}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File -- GPU}
\begin{semiverbatim}\tiny
[login lus]\$ cat gpu.slurm

\#!/bin/bash
\#SBATCH \ddash{}job-name=GPU\_example
\textcolor{blue}{\#SBATCH \ddash{}partition=gpu.q}
\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=1
\#SBATCH \ddash{}tasks-per-node=1
\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=10}
\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
\textcolor{blue}{\#SBATCH \ddash{}gres=gpu:NV-K40:2 } \#\# request both GPUs in the GPU node
\#\#\# To request only 1 of the two GPUs in the node, you would do: gpu:NV-K40:1
\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile
module load GPGPU/cuda/samples/7.5

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}

bandwidthTest; cudaOpenMP; deviceQuery; simpleMultiGPU
\end{semiverbatim}
\end{frame}

\subsection{MPI Batch Job Example Script}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

\#!/bin/bash
\#SBATCH \ddash{}job-name=MPI\_example
\textcolor{blue}{\#SBATCH \ddash{}partition=exclusive.q}
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\textcolor{blue}{\#SBATCH \ddash{}nodes=4}
\textcolor{blue}{\#SBATCH \ddash{}tasks-per-node=20}
\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=1}
\textcolor{blue}{\#\#SBATCH \ddash{}mem=11000} \# Memory should not be set for jobs in exclusive.q
\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
\#\# Intel MPI manual: \href{https://software.intel.com/en-us/mpi-refman-lin-html}{https://software.intel.com/en-us/mpi-refman-lin-html}
export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}
export I\_MPI\_FABRICS=shm:tmi  
export I\_MPI\_PMI\_LIBRARY=/opt/local/slurm/default/lib64/libpmi.so

srun  -n \$\{SLURM\_NTASKS\}  ./hello\_mpi.intel 
\end{semiverbatim}
\end{frame}

\subsection{Non-MPI Batch Job Example Script}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

\#!/bin/bash
\#SBATCH \ddash{}job-name=example
\textcolor{blue}{\#SBATCH \ddash{}partition=community.q}
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=1
\#SBATCH \ddash{}tasks-per-node=1
\textcolor{blue}{\#SBATCH \ddash{}cpus-per-task=5}
\textcolor{blue}{\#SBATCH \ddash{}mem=11000} \#\# max amount of memory per node you require
\#SBATCH \ddash{}error=hello-\%A.err \#\# \%A - filled with jobid
\#SBATCH \ddash{}output=hello-\%A.out \#\# \%A - filled with jobid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}
export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}

./hello\_world
\end{semiverbatim}
\end{frame}

\subsection{Executing and Monitoring Batch Jobs}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
To execute a submission script you use the \textbf{\texttt{sbatch}} command
\begin{block}{Example}
\begin{semiverbatim}\tiny
[login lus]\$ sinfo
PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
community.q   up     3-00:00:00 2      idle  compute-[0001-0002]

[login lus]\$ sbatch hello_world.slurm
sbatch: Submitted batch job 469

[login lus]\$ squeue
JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
469   community.q  example  user99  R  00:01 1      compute-0001
\end{semiverbatim}
\end{block}

\btVFill
\begin{center}What if I want to run many jobs with identical parameters, but on different inputs?~\\
  SLURM provides a mechanism for this called a \textbf{Job Array}.\end{center}

\end{frame}

\begin{frame}
  \frametitle{Slurm Job Array}
\begin{center}  
  \includegraphics[width=0.90\textwidth]{images/job_array}
\end{center}
\end{frame}

\subsection{Batch Job Example Script Using Job Array}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm

\#!/bin/bash
\#SBATCH \ddash{}job-name=example
\#SBATCH \ddash{}partition=community.q
\#\# 3 day max run time for community.q, kill.q, exclusive.q, and htc.q.  1 Hour max run time for sb.q
\#SBATCH \ddash{}time=3-00:00:00 ## time format is DD-HH:MM:SS
\#\# task-per-node x cpus-per-task should not typically exceed core count on an individual node 
\#SBATCH \ddash{}nodes=1
\#SBATCH \ddash{}tasks-per-node=1
\#SBATCH \ddash{}cpus-per-task=5
\#SBATCH \ddash{}mem=11000 \#\# max amount of memory per node you require
\textcolor{blue}{\#SBATCH \ddash{}error=myprog-\%A\_\%a.err} \#\# \%A - filled with jobid. \%a - filled with job arrayid
\textcolor{blue}{\#SBATCH \ddash{}output=myprog-\%A\_\%a.out} \#\# \%A - filled with jobid. \%a - filled with job arrayid
\#\# Useful for remote notification
\#SBATCH \ddash{}mail-type=BEGIN,END,FAIL,REQUEUE,TIME\_LIMIT\_80
\#SBATCH \ddash{}mail-user=user@test.org

source \ctilde/.bash_profile \#if you want to use modules or need environment variables, source your bash profile

\#\# All options and environment variables found on schedMD site: \href{http://slurm.schedmd.com/sbatch.html}{http://slurm.schedmd.com/sbatch.html}

export OMP\_NUM\_THREADS=\$\{SLURM\_CPUS\_PER\_TASK\}

./worker\_bee -i input\_\$\{SLURM_ARRAY_TASK_ID\}.dat -o output\_\$\{SLURM_ARRAY_TASK_ID\}
\end{semiverbatim}
\end{frame}

\subsection{Executing and Monitoring Job Array Batch Jobs}

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Executing \& Monitoring Job Arrays}\footnotesize
\begin{block}{Example}
\begin{semiverbatim}\tiny
[login lus]\$ sinfo
PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
community.q   up     3-00:00:00 2      idle  compute-[0001-0002]

[login lus]\$ sbatch {\ddash}array=1-30 job_array.slurm
sbatch: Submitted batch job 470

[login lus]\$ squeue
JOBID      PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
470[3-30]  community.q  example  user99  PD   00:00   1      (resource)
470_1      community.q  example  user99  R    00:01   1      compute-0001
470_2      community.q  example  user99  R    00:05   1      compute-0002
\end{semiverbatim}
\end{block}
\end{frame}

### 2017-02-04 hpc_101/sections/06_ediquette.tex
Breakup of sections into tex file

\section{Cluster Etiquette}
\begin{frame}
  \frametitle{Cluster Etiquette}
  \begin{itemize}
  \item Never run anything on the login nodes
  \item Try to request only the resources that your job needs.  Leaving unneeded resources available, allows other users to use them
  \item Test your application in an interactive session before scheduling a long running job
    \begin{itemize}
    \item[--] This can help identify what resources your job will require, and potentially aid you in correctly sizing what resources you request.
    \end{itemize}
  \item Use the nodes in the sb.q to validate your slurm script runs as intended
    \begin{itemize}
    \item[--] This helps to reduce the chance of you having a small error that kills your job prematurely after waiting for resources
    \end{itemize}
  \item Always run jobs from within {\ctilde}/lus/ directory to ensure output is written there and does not fill up the system filesystem
  \end{itemize}
\end{frame}

### 2017-02-04 hpc_101/sections/07_policies.tex
Breakup of sections into tex file

\section{Policies}
%% \begin{frame}
%% 	\frametitle{Overview}
%% 	\begin{itemize}
%% 		\item Login node usage policy
%% 		\item Scratch filesystem purge policy
%% 	\end{itemize}
%% \end{frame}

\subsection{Login Node Usage Policy}
\begin{frame}
\frametitle{Login Node Usage Policy}\footnotesize
The login nodes are a shared resource and are the only access to the cluster for hundreds of user and are meant to provide the following functionality for all users: 
\begin{itemize}
\item Providing ssh shell access 
\item Facilitate the transfer files to and from the cluster:\\Globus, sftp, scp, rsync
\item Launching and monitoring SLURM jobs (batch and interactive)
\item Modifying text files with a text editor: vi/vim, emacs, nano
\end{itemize}
\bigskip
\begin{itemize}
\item[--] If we identify or are notified that a user is running computation on the login nodes, the application will be killed
\item[--] If we determine that a user repeatedly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
\end{itemize}
\btVFill
\begin{center}
\footnotesize \textbf{\emph{Login node usage policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
\end{center}
\end{frame}

\subsection{Lustre Filesystem Purge Policies}
\begin{frame}
\frametitle{{\lustre} Filesystem Purge Policies}
Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.  To accomplish this, we have currently implemented a purge policy, for any file that is older than 90 days.
\begin{itemize}
 \item Which of my directories are subject to the purge policy?
   \begin{itemize}
   \item \textbf{Answer:} All files and folders found in \ctilde/lus/ are subject to the 90 day purge policy.  Symlinks are excluded from the purge, and are not followed by the purge bot.  
   \end{itemize}
\item How frequently are files for purging identified?
   \begin{itemize}
   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
   \end{itemize}
\end{itemize}
\btVFill
\begin{center}
\footnotesize \textbf{\emph{Purge policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
\end{center}
\end{frame}

### 2017-02-04 hpc_101/sections/09_related_services.tex
Breakup of sections into tex file

\section{Related Services}
\begin{frame}
\frametitle{Services}
\begin{itemize}
\item MATLAB Distributed Computing Server{\trademark}
  \begin{itemize}
    \item Requires user to already have a license for MATLAB
    \item User must also have access to at least the Parallel Computing Toolbox{\trademark}
    \item Provides users access to up to 64 workers
    \item Requires some additional configuration
    \item Email us to request access to this additional service
  \end{itemize}
\end{itemize}
\end{frame}

### 2017-02-04 hpc_101/sections/10_faq.tex
Breakup of sections into tex file

\section*{Frequently Asked Questions}
\begin{frame}
\frametitle{FAQ}
\begin{itemize}\footnotesize
\item Are files on the cluster backed up?
  \begin{itemize}\tiny
  \item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.  It is up to you as the user to validate and maintain your own backups.  The {\citeam} takes no responsibility for any data that is lost due to human or mechanical error.  
  \end{itemize}

\item Help I have been trying to compile an application without success, can the {\citeam} help?
  \begin{itemize}\tiny
  \item \textbf{Answer:} We are willing to help, just send us an email at \textit{uh-hpc-help@lists.hawaii.edu} with a description of what you have attempted as well as a link to the software you are trying to compile, and/or where it is located on the cluster.
  \end{itemize}

\item HELP! My job needs more time and does not support check-pointing.  What can I do?
  \begin{itemize}\tiny
  \item \textbf{Answer:} We understand that jobs may not all fit within the 3 day timelimit or unforeseen issues during the execution can cause a job to run long.  The {\citeam} does review and potentially approves requests to extend jobs in a one off fashion.  A request to extend a job can be denied, since we need to take into account factor such as: how busy the cluster is, the number of time extension requests a user has made in recent history.\\If it is known before the job is submitted that more than 3 days is required, we would ask that you contact us first so we can verify that you truly will need more time.\\If a job is already running and just doesn't look like it will complete in time, please send us a request with the jobid number and what you want the job extended to e.g., please extend my jobs run time to 7 days.\\Please remember the {\citeam} is small and we are also human.  Depending on when you make your request we may not be able to immediately respond or act.  It is best to make your request at the first signs you need more time, in hopes that you can provide us with ample time to act.
  \end{itemize}

\end{itemize}

\end{frame}

\begin{frame}	
\frametitle{FAQ}
\begin{itemize}
  \item Can I run a display program on the cluster?
    \begin{itemize}
      \item \textbf{Answer:} Yes, but not from the login nodes.  To follow cluster usage policies, please run X11 applications from a compute node.  Please see the following steps.
\end{itemize}
\end{itemize}
 \begin{block}{Interactive session with X11}\tiny  
       \begin{enumerate}
      \item Connect via SSH using the -Y option, X11 forwarding enabled
      \item run srun.x11 to start a session on a node
      \end{enumerate}
      \
      ~\\
      ~\\
      $[$local \ctilde$]$\$ ssh -Y user99@uhhpc1.its.hawaii.edu	~\\
      $[$login \ctilde$]$\$ srun.x11 \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00	~\\
      $[$compute-0001 \ctilde$]$\$ xterm	~\\
 \end{block}

\end{frame}

### 2017-02-04 hpc_101/sections/02_cluster.tex
removed valuestorage and commented out stainability and related services
redo of history

@@
 \begin{itemize}
- \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
- \item Delivered in October of 2014
- \item Accepted in December of 2014 
- \item Early adopter testing started in January 2015
- \item Opened for general use in April 2015
- \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
- \item January 2017, 5 new nodes added to cluster, raising total core count from 5,876 to 5,992 cores
- \item As of Nov 2016, more than 300 users have been granted access to the cluster
+\item[] 2014
+  \begin{itemize}
+  \item Jul. -- University of {\hawaii} (UH) selects Cray's proposal
+  \item Oct. -- Delivered 184 nodes, 3,800 cores
+  \item Dec. -- Passed testing and accepted
+  \end{itemize}
+
+\item[] 2015
+  \begin{itemize}
+  \item Jan. -- Early adopter testing started (8 users)
+  \item Apr. -- Opened for general use
+  \end{itemize}
+\item[] 2016
+  \begin{itemize}
+  \item Mar. -- 92 nodes added, 5,876 cores total
+  \end{itemize}
+\item[] 2017
+  \begin{itemize}
+  \item Jan. -- 5 nodes added, 5,992 cores total
+  \item Jan. -- 370$+$ users have been granted access to the cluster
+  \end{itemize}
 \end{itemize}

### 2017-02-04 hpc_101/sections/02_cluster.tex
compressed storage slides

@@
 \begin{frame}
-	\frametitle{{\craycs} -- Storage}
-	Two storage options are currently available on the {\craycs}
-	\begin{enumerate}
-		\item {\lustre}
-                  \begin{itemize}
-                  \item High performance parallel filesystem 
-                  \item Available on all the cluster nodes
-                    \item Freely available to all users
-                  \end{itemize}
-		\item ValueStorage
-                  \begin{itemize}
-                  \item Network attached scale out storage
-                  \item Only available from the login nodes
-                  \item Available as a for fee service
-                  \end{itemize}
-	\end{enumerate}
-\end{frame}
-
-\begin{frame}
 	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
 	\begin{itemize}
+        \item Lustre is a high performance parallel filesystem 
+        \item Available on all the cluster nodes
+
 	\item The {\craycs} has $\approx582$TB of storage space
@@
                 \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
+
 	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
-        ~\\
-	\begin{itemize}
-		\item $500$TB of scale out storage
-		\item Purchased annually in 0.5TB increments 
-	\end{itemize}
-	\btVFill
-
-	\begin{center}
-	\textbf{ValueStorage Pricing}~\\ 
-		\resizebox{0.50\textwidth}{!}{%
-		\begin{tabular}{l r }
-			\toprule
-			\textbf{Product} & \textbf{Annual Cost}~~ \\
-	                \midrule
-			\midrule
-			0.5TB & \$65.00\\
-                        0.5TB + Replication  & \$130.00 \\
-			\bottomrule
-		\end{tabular}
-		}
-		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
-		~\\	{\footnotesize \emph{All prices are subject to change}}
-	\end{center}
 \end{frame}

### 2017-02-16 hpc_101/sections/02_cluster.tex
changed node description to table fixed up footnotes across multiple slides.  Also removed section about related services (matlab DCS)

--- 2017-02-04 hpc_101/sections/02_cluster.tex
@@
   \frametitle{{\craycs} -- Compute Nodes}
-	\begin{itemize}
-        \item All nodes are diskless with some RAM used for the Operating System
-	\item 178 Standard nodes {\tiny(\emph{community})}
-	  \begin{itemize}
-            {\footnotesize
-	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
+  \begin{table}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tabular}{|l|l|l|l|l|l|l|l|}
+        \multicolumn{8}{c}{ {\large {\textbf{ Community Nodes\footnote{\label{node_note}\tiny Nodes are diskless and run CentOS linux}} } } }\\ \hline
+        \textbf{Type} & \textbf{Cores} & \textbf{Ivy-Bridge} & \textbf{Haswell} & \textbf{Broadwell} &
+        \textbf{128 GB}\footnote{\label{mem_limit}\tiny Usable RAM: $\approx118$ GB of 128 GB, $\approx246$ GB of 256 GB, $\approx1003$ GB of 1 TB} & \textbf{256 GB}\textsuperscript{\ref{mem_limit}}  & \textbf{1 TB}\textsuperscript{\ref{mem_limit}}  \\ \hline
+        Standard & 20 & 178 &  &  & 178 &  &  \\ \hline
+        Large & 40 & 6 &  &  &  &  & 6 \\ \hline
+        GPU\footnote{\label{gpu_v1}\tiny 2 x Nvidia Tesla K40} & 20 &  & 1 &  & 1 &  & \\ \hline
+      \end{tabular}%
     }
-	  \end{itemize}          
-	\item 6 Large memory nodes {\tiny(\emph{community})}
-	  \begin{itemize}
-            {\footnotesize
-	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
-	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
+  \end{table}
+
+  \begin{table}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tabular}{|l|l|l|l|l|l|l|l|}
+        \multicolumn{8}{c}{ {\large {\textbf{ Condo Nodes}\textsuperscript{\ref{node_note}} } }} \\ \hline
+        \textbf{Type} & \textbf{Cores} & \textbf{Ivy-Bridge} & \textbf{Haswell} & \textbf{Broadwell} & \textbf{128 GB}\textsuperscript{\ref{mem_limit}} & \textbf{256 GB}\textsuperscript{\ref{mem_limit}} & \textbf{1 TB}\textsuperscript{\ref{mem_limit}} \\ \hline
+        Standard & 24 & & 62 & & 62 & & \\ \hline
+        Standard & 20 & & & (5)\footnote{\label{future}\tiny April 2017 additions} & (5)\textsuperscript{\ref{future}} & & \\ \hline
+        Custom & 20 &  & 33 & (4)\textsuperscript{\ref{future}} &  & 33 & (4)\textsuperscript{\ref{future}} \\ \hline
+        GPU\textsuperscript{\ref{gpu_v1}}     & 20 &  & 1 &  & 1 &  &  \\ \hline
+      \end{tabular}%
     }
-	  \end{itemize}			
-        \item 1 GPU node {\tiny(\emph{community})}
-          \begin{itemize}
-            {\footnotesize
-	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
-	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
-            \item 2 Nvidia Tesla K40 GPUs
-            }
-	  \end{itemize}
-        \item 96 Owner nodes
-          \begin{itemize}
-            {\footnotesize
-            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
-            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
-            \item 1 GPU node
-            }
-          \end{itemize}
-	\item CentOS Linux
-	\end{itemize}
+    \end{table}
   \end{frame}

### 2017-02-16 hpc_101/sections/04_cluster_interaction.tex
changed node description to table fixed up footnotes across multiple slides.  Also removed section about related services (matlab DCS)

@@
 	\begin{block}{Modules}\tiny
-	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark
+	  A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnote{\label{wiki_module}\tiny\
+             \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
 
@@
 	\end{itemize}
-	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
+
 \end{frame}
@@
 	\begin{itemize}
-		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
+		\item We have the {\intel} \& GNU (gcc, g$++$) compilers
 		\item Compiling must take place on a compute node
@@
 Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc.
-Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
+Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnote{\label{schedmd_qstart}\tiny \url{http://slurm.schedmd.com/quickstart.html}}
 \end{block}
@@
 \end{itemize}
-\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
 \end{frame}
@@
     ~\\
-          \hrule
     ~\\
-	$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash
+    \hrule\begin{semiverbatim}$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash\end{semiverbatim}
 

### 2017-06-01 hpc_101/sections/02_cluster.tex
corrections to early slides of hpc101, addition of new members, correction to node prices.

@@
   \item Jan. -- 5 nodes added, 5,992 cores total
-  \item Jan. -- 370$+$ users have been granted access to the cluster
+  \item May -- 5 nodes added, 6,092 cores total
+  \item Jun. -- 452$+$ users have been granted access to the cluster
   \end{itemize}
@@
         Standard & 24 & & 62 & & 62 & & \\ \hline
-        Standard & 20 & & & (5)\footnote{\label{future}\tiny April 2017 additions} & (5)\textsuperscript{\ref{future}} & & \\ \hline
-        Custom & 20 &  & 33 & (4)\textsuperscript{\ref{future}} &  & 33 & (4)\textsuperscript{\ref{future}} \\ \hline
+        Standard & 20 & & & 5 & 5 & & \\ \hline
+        Custom & 20 &  & 33 $+$ (4)\footnote{\label{future}\tiny June 2017 additions} & &  & 33 & (4)\textsuperscript{\ref{future}} \\ \hline
         GPU\textsuperscript{\ref{gpu_v1}}     & 20 &  & 1 &  & 1 &  &  \\ \hline

### 2017-06-01 hpc_101/sections/04_cluster_interaction.tex
corrections to early slides of hpc101, addition of new members, correction to node prices.

@@
 	\end{definition}
+{\tiny Please visit the {\ci} Globus quick start guide for instructions on how to use Globus:~\href{http://go.hawaii.edu/jMW}{http://go.hawaii.edu/jMW}}
+
         \numlessfootnotetxt{\tiny \url{https://www.globus.org/file-transfer}}
@@
 
-\begin{frame}
-	\frametitle{Globus}\footnotesize
-\begin{block}{How do I get Globus?}
-To utilize Globus, follow the following steps:
-\begin{itemize}\tiny
-\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
-\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
-\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
-\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
-\begin{itemize}\tiny
- \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
-\end{itemize}
-\item Once you select an endpoint, a login window will pop up.
+%% \begin{frame}
+%% 	\frametitle{Globus}\footnotesize
+%% \begin{block}{How do I get Globus?}
+%% To utilize Globus, follow the following steps:
+%% \begin{itemize}\tiny
+%% \item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
+%% \item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
+%% \item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
+%% \item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
+%% \begin{itemize}\tiny
+%%  \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
+%% \end{itemize}
+%% \item Once you select an endpoint, a login window will pop up.
 You can access the UHHPC endpoints by simply using your UH username and password.
 Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’.
-\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
-\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
-\end{itemize}
-\end{block}
-~\\
-These instructions can also be found on the {\ci} website:~\\
-{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
-\end{frame}
+%% \item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
+%% \item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
+%% \end{itemize}
+%% \end{block}
+%% ~\\
+%% \end{frame}
 
-\begin{frame}
-	\frametitle{Globus}
-        \begin{itemize}
-          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
-          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
-          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
-            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
-\end{itemize} 
-\end{frame}
+%% \begin{frame}
+%% 	\frametitle{Globus}
+%%         \begin{itemize}
+%%           \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
+%%           \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
+%%           \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
+%%             \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
+%% \end{itemize} 
+%% \end{frame}
 
@@
 \hline
-kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{12}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
+kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline

### 2017-07-18 hpc_101/sections/02_cluster.tex
update of what nodes are installed

@@
         Standard & 20 & & & 5 & 5 & & \\ \hline
-        Custom & 20 &  & 33 $+$ (4)\footnote{\label{future}\tiny June 2017 additions} & &  & 33 & (4)\textsuperscript{\ref{future}} \\ \hline
+        Custom & 20 &  & 37 & &  & 33 & 4 \\ \hline
         GPU\textsuperscript{\ref{gpu_v1}}     & 20 &  & 1 &  & 1 &  &  \\ \hline

### 2018-04-03 hpc_101/sections/02_cluster.tex
removed history, added lease option, fixed partition info

@@
 
-\subsection{{\craycs}}
-\begin{frame}
-\frametitle{{\craycs} -- History}
-\begin{itemize}
-\item[] 2014
-  \begin{itemize}
-  \item Jul. -- University of {\hawaii} (UH) selects Cray's proposal
-  \item Oct. -- Delivered 184 nodes, 3,800 cores
-  \item Dec. -- Passed testing and accepted
-  \end{itemize}
-
-\item[] 2015
-  \begin{itemize}
-  \item Jan. -- Early adopter testing started (8 users)
-  \item Apr. -- Opened for general use
-  \end{itemize}
-\item[] 2016
-  \begin{itemize}
-  \item Mar. -- 92 nodes added, 5,876 cores total
-  \end{itemize}
-\item[] 2017
-  \begin{itemize}
-  \item Jan. -- 5 nodes added, 5,992 cores total
-  \item May -- 5 nodes added, 6,092 cores total
-  \item Jun. -- 452$+$ users have been granted access to the cluster
-  \end{itemize}
-\end{itemize}
-\end{frame}
+%% \subsection{{\craycs}}
+%% \begin{frame}
+%% \frametitle{{\craycs} -- History}
+%% \begin{itemize}
+%% \item[] 2014 -- Accepted delivery of 184 nodes, 3,800 cores  
+%% \item[] 2015 -- Opened for general use
+%% \item[] 2016 -- 92 nodes added, 5,876 cores
+%% \item[] 2017 -- 14 nodes added, 6,172 cores 
+%% %\item[] 2018
+%% %  \begin{itemize}
+%% %  \item Apr. -- 549$+$ users have been granted access to the cluster
+%% %  \end{itemize}
+%% \end{itemize}
+%% \end{frame}
 
@@
 \begin{frame}
-  \frametitle{{\craycs} -- Compute Nodes}
+  \frametitle{{\craycs}}% -- Compute Nodes}
   \begin{table}
@@
       \begin{tabular}{|l|l|l|l|l|l|l|l|}
-        \multicolumn{8}{c}{ {\large {\textbf{ Community Nodes\footnote{\label{node_note}\tiny Nodes are diskless and run CentOS linux}} } } }\\ \hline
+        \multicolumn{8}{c}{ {\large {\textbf{ Community Nodes\footnote{\label{node_note}\tiny Nodes are diskless and run CentOS linux}} } } } \\ \hline
         \textbf{Type} & \textbf{Cores} & \textbf{Ivy-Bridge} & \textbf{Haswell} & \textbf{Broadwell} &
@@
                   \item Only available from the login nodes
-                  \item Available as a for fee service
+                  \item Available as a for fee service (\href{http://www.hawaii.edu/its/value-storage-pricing/}{Pricing})
                   \end{itemize}
@@
 		\item User do not have a usage quota (soft or hard)
-		\item Certain directories are subject to a 90 day purge policy
+		\item User scratch directories are subject to a 30 day purge policy (Begins May 2nd)
 		\item \textbf{Data is not backed up!  Users are responsible for their own data}
@@
 
-\begin{frame}
-	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
-        ~\\
-	\begin{itemize}
-		\item $500$TB of scale out storage
-		\item Purchased annually in 0.5TB increments 
-	\end{itemize}
-	\btVFill
+%% \begin{frame}
+%% 	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
+%%         ~\\
+%% 	\begin{itemize}
+%% 		\item $500$TB of scale out storage
+%% 		\item Purchased annually in 0.5TB increments 
+%% 	\end{itemize}
+%% 	\btVFill
 
-	\begin{center}
-	\textbf{ValueStorage Pricing}~\\ 
-		\resizebox{0.50\textwidth}{!}{%
-		\begin{tabular}{l r }
-			\toprule
-			\textbf{Product} & \textbf{Annual Cost}~~ \\
-	                \midrule
-			\midrule
-			0.5TB & \$65.00\\
-                        0.5TB + Replication  & \$130.00 \\
-			\bottomrule
-		\end{tabular}
-		}
-		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
-		~\\	{\footnotesize \emph{All prices are subject to change}}
-	\end{center}
-\end{frame}
+%% 	\begin{center}
+%% 	\textbf{ValueStorage Pricing}~\\ 
+%% 		\resizebox{0.50\textwidth}{!}{%
+%% 		\begin{tabular}{l r }
+%% 			\toprule
+%% 			\textbf{Product} & \textbf{Annual Cost}~~ \\
+%% 	                \midrule
+%% 			\midrule
+%% 			0.5TB & \$65.00\\
+%%                         0.5TB + Replication  & \$130.00 \\
+%% 			\bottomrule
+%% 		\end{tabular}
+%% 		}
+%% 		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
+%% 		~\\	{\footnotesize \emph{All prices are subject to change}}
+%% 	\end{center}
+%% \end{frame}
 

### 2018-04-03 hpc_101/sections/03_sustainability.tex
corrections to early slides of hpc101, addition of new members, correction to node prices.
removed history, added lease option, fixed partition info

@@
 		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
-		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
+		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, kill.gpu.q, htc.q}
 		\item For some users, the publicly available resources may not be enough~\ldots
@@
 		\item Node owners are provided with priority access to their purchased hardware
-		\item All nodes have a \emph{5 year warranty}
+%		\item All nodes have a \emph{5 year warranty}
+%		\begin{itemize}
+%			\item Once a nodes warranty has expired, it will be removed from the cluster
+%		\end{itemize}
+%		\item Condo owners are given early access to purchased nodes
+%		  \begin{itemize}
+%                    \item Access is granted as soon as the funds land in our accounts
+%		    \item The 5 year warranty will not begin until newly ordered nodes are installed
+%		\end{itemize}
+		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
+                \item {\textcolor{red}{\textbf{ Condo purchases currently suspended} }}
+	\end{itemize}
+        \btVFill
+        \begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
+\end{frame}
+
+\subsubsection{Node Leasing}
+\begin{frame}
+	\frametitle{Leasing}
 	\begin{itemize}
-			\item Once a nodes warranty has expired, it will be removed from the cluster
+        \item In some cases, users will not require owning a node for multiple years, but still need priority access
+	\item Leases have a 1 year term
+        \item Leases are $\$1,320$ per year, which is \( \frac{1}{5} \) the cost of a condo purchase
+        \item Node leasees are provided priority access to their leased hardware
+        \item Leases are not considered an equipment purchase
+        \item Total availabity limited to $\approx100$ community nodes
 	\end{itemize}
-		\item Condo owners are given early access to purchased nodes
-		  \begin{itemize}
-                    \item Access is granted as soon as the funds land in our accounts
-		    \item The 5 year warranty will not begin until newly ordered nodes are installed
-		\end{itemize}
-		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
-	\end{itemize}
+        \btVFill
+        \begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
 \end{frame}
@@
 	\begin{itemize}
-	 \item In some cases, users will not require owning a node, but still need priority access
-	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
+	 \item In other cases, for users that require a large burst of dedicated nodes, service units (\emph{SU}) is an option
 	 \item SUs come in two flavors:
 		 \begin{itemize}
-			 \item[--] Standard node units 
+			 \item[--] Standard node units -- (\$0.50 per SU) 
 				\begin{itemize}
@@
 				\end{itemize}
-			 \item[--] Large memory node units
+			 \item[--] Large memory node units -- (\$2.00 per SU)
 				\begin{itemize}
@@
 	\end{itemize}
+        \btVFill
+	\begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
 \end{frame}
 
-\subsubsection{Price Card}
-\begin{frame}
-	\frametitle{Price Card}
-	\begin{center}
-	{\large \textbf{Condo Price Card}} \\
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l r ||l r}
-			\toprule
-			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
-			\midrule
-			\midrule
-			Standard & \$6,600.00  & Standard + 2 GPUs (Nvidia{\regtrademark} K40) & \$13,600.00 \\
-			\hline
-			Large Memory & \$33,900.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
-			\bottomrule
-		\end{tabular}
-		}
-	\end{center}
-	\smallskip
-	\begin{center}
-		{\large \textbf{Service Unit Price Card}}\\
-		\begin{tabular}{l || c || r}
-			\toprule
-			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
-			\midrule
-			\midrule
-			Standard node & \$0.50 per SU &  1,000~SU (\$500.00)\\
-			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
-			\bottomrule
-		\end{tabular}
-		\btVFill
-	{\footnotesize \emph{All prices are subject to change}}
-	\end{center}
-\end{frame}
+%% \subsubsection{Price Card}
+%% \begin{frame}
+%% 	\frametitle{Price Card}
+%% 	\begin{center}
+%% 	{\large \textbf{Condo Price Card}} \\
+%% 	\resizebox{\textwidth}{!}{%
+%% 		\begin{tabular}{l r ||l r}
+%% 			\toprule
+%% 			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
+%% 			\midrule
+%% 			\midrule
+%% 			Standard & \$6,600.00  & GPU Node& Contact for pricing\\
+%% 			\hline
+%% 			Large Memory & \$20,000.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
+%% 			\bottomrule
+%% 		\end{tabular}
+%% 		}
+%% 	\end{center}
+%% 	\smallskip
+%% 	\begin{center}
+%% 		{\large \textbf{Service Unit Price Card}}\\
+%% 		\begin{tabular}{l || c || r}
+%% 			\toprule
+%% 			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
+%% 			\midrule
+%% 			\midrule
+%% 			Standard node &  &  1,000~SU (\$500.00)\\
+%% 			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
+%% 			\bottomrule
+%% 		\end{tabular}
+%% 		\btVFill
+%% 	{\footnotesize \emph{All prices are subject to change}}
+%% 	\end{center}
+%% \end{frame}

### 2018-04-03 hpc_101/sections/04_cluster_interaction.tex
removed history, added lease option, fixed partition info

@@
 	\item[] \begin{itemize}
-		\item The {\craycs} has two login nodes:
+		\item The {\craycs} has one active login node:
 		\begin{itemize}
 			\item uhhpc1.its.hawaii.edu
-			\item uhhpc2.its.hawaii.edu
+			%\item uhhpc2.its.hawaii.edu
 		\end{itemize}
@@
 			\begin{itemize}
-			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
+			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu %--OR-- uhhpc2.its.hawaii.edu
 			\item[] \textbf{User Name:} Your UH User name e.g., user99
@@
 			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
-			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
+%			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
 			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
@@
 lrwxrwxrwx 1 user99 user99 19 Jan 15 20:38 \textcolor{teal}{lus} -> /lus/scratch/user99
-lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
+
 \end{semiverbatim}
 \end{block}
+% lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
 \begin{itemize}\footnotesize
@@
 			\item This is where all your job data files should live
-			\item Items in this directory \emph{\textbf{are}} subject to our 90 day purge policy
+			\item Items in this directory \emph{\textbf{are}} subject to our purge policy
 		\end{itemize}
@@
 		\begin{itemize}\tiny
-			\item Items in this directory \emph{\textbf{are not}} subject to our 90 day purge policy
+			\item Items in this directory \emph{\textbf{are not}} subject to our purge policy
 			\item Directory is monitored for abuse
-		\end{itemize}
-		\item \ctilde{}/purge/ is typically a dead symlink
-		\begin{itemize}\tiny
-			\item Symlink becomes active if the user has files that are part of the next automatic purge
-			\item When \ctilde{}/purge/ is active, the directory containing two files -- \emph{purge\_list.txt} \& \emph{totals.txt}
-			\item An email is sent to users if they have files that will be purged
-			\item Email notification is sent out 14 days before the purge takes place
 		\end{itemize}
@@
 			\item No hard or soft quotas are in place
-			\item Utilization is managed through the 90 day purge policy
+			\item Utilization is managed through the  purge policy
 		\end{itemize}
@@
 			\item `module load $<$module name$>$' -- Loads the named module
+                        \item `module list -- Show what modules are loaded
 			\item `module purge' -- Unload all loaded modules
@@
 			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
-			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
+			\item Login nodes \textbf{do not} mirror the compute node environment
 		\end{itemize}
@@
 \begin{itemize}
-	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
+	\item The current public partitions on the {\craycs}:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{kill.gpu.q}, \textbf{htc.q}
 	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
@@
 \midrule
-community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
+community.q & \begin{tabular}{l r} \textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} & \begin{tabular}{c c} \textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} & \padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & OFF & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline
@@
 kill.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{20}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
+\hline
+\hline
+kill.gpu.q & \begin{tabular}{l r}\textbf{Default:} & 0-00:10:00~\\ \textbf{Max:} & 3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min:} & \padfrom{\ \ \ }{1}~\\ \textbf{Max:} &\padfrom{\ \ \ }{1}\end{tabular} & 10 & NO & REQUEUE & \begin{tabular}{l c}\textbf{Default:} & \padfrom{\ \ \ \ \ \ \ }{3250}~\\ \textbf{Max:} & \padfrom{\ \ \ \ \ \ \ \ }{$\infty$}\end{tabular}\\
 \hline

### 2018-04-03 hpc_101/sections/07_policies.tex
removed history, added lease option, fixed partition info

@@
 %% \end{frame}
+
+\subsection{Use and Management of IT Resources @ UH }
+\begin{frame}
+  \frametitle{Use and Management of IT Resources @ UH }
+  All usage of the {\craycs} must be in compliance with \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes} and is subject to University of Hawaii Executive Policy \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{E2.210}.
+\end{frame}
 
@@
 Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.
-To accomplish this, we have currently implemented a purge policy, for any file that is older than 90 days.
+To accomplish this, we have currently implemented a purge policy, for any file that is older than 30 days.
 \begin{itemize}
- \item Which of my directories are subject to the purge policy?
+ \item What in \ctilde/lus/ is subject to the purge policy?
    \begin{itemize}
-   \item \textbf{Answer:} All files and folders found in \ctilde/lus/ are subject to the 90 day purge policy.  Symlinks are excluded from the purge, and are not followed by the purge bot.  
+   \item \textbf{Answer:} Everything but directories and files between 1 byte and 32 Kilobytes are subject to the purge policy.
    \end{itemize}
-\item How frequently are files for purging identified?
-   \begin{itemize}
-   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
-   \end{itemize}
+%\item How frequently are files for purged?
+%   \begin{itemize}
+%   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
+%   \end{itemize}
 \end{itemize}

### 2018-04-23 hpc_101/sections/07_policies.tex
correction to purge policy

@@
 \frametitle{{\lustre} Filesystem Purge Policies}
-Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.
-To accomplish this, we have currently implemented a purge policy, for any file that is older than 30 days.
+Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.  To accomplish this, we have implemented a purge policy.
+
 \begin{itemize}
@@
    \begin{itemize}
-   \item \textbf{Answer:} Everything but directories and files between 1 byte and 32 Kilobytes are subject to the purge policy.
+   \item \textbf{Answer:} Everything but directories.
    \end{itemize}
+\item When will my files be purged?
+  \begin{itemize}  
+    \item \textbf{Answer:} Files greater than or equal to 1 Megabytein size  or are empty (0 bytes) that are 35 days old
+    \item \textbf{Answer:} Files between 1 byte and 1 Megabyte in size that are 120 days old
+  \end{itemize}
+\item What is the age based on?
+   \begin{itemize}
+   \item \textbf{Answer:} Age of file is based on creation time
+   \end{itemize}
+
 %\item How frequently are files for purged?

### 2018-05-08 hpc_101/sections/07_policies.tex
added age_check instructions

@@
 \end{frame}
+
+\begin{frame}
+  \frametitle{How to check file age for purge}
+  A tool is available on the cluster called age\_check.  This tool will list all files that are older than the defined date you set.
+\begin{semiverbatim}\tiny \texttt
+  $[$root@login-0001 \ctilde$]$\$ age\_check -h
+
+  Usage:
+  age\_check [-s] [-z] [-d] [-f] [-a <age cutoff>] [-p <path]
+
+  Options:
+
+  -s -- Sort output
+
+  -d -- display creation date per file
+
+  -z -- display size in bytes per file
+
+  -f -- Do not exclude files between 1 byte and 1 MB
+
+  -a age -- Lists files that were created at least age days ago (default is 20 days)
+
+  -p path -- Directory to check in (default is /lus/scratch/root)
+
+  Example:
+  age\_check -a 30
+
+\end{semiverbatim}
+
+\end{frame}

### 2019-05-25 hpc_101/paper.tex
uncomment of all sections
changed node description to table fixed up footnotes across multiple slides.  Also removed section about related services (matlab DCS)
switch over to new presentation for new cluster

@@
 %\documentclass[t]{beamer}
-\documentclass[t,hyperref={pdfpagelabels=false}]{beamer}
+\documentclass[t,hyperref={pdfpagelabels=false},9pt]{beamer}
 
@@
 \usepackage{color}
+\usepackage{makecell}
 \mode<presentation> {
@@
 }
+%\setbeamerfont{section in toc}{size=\small}
+%\setbeamerfont{subsection in toc}{size=\footnotesize}
+
+%\setbeamerfont{section in toc}{size=fontsize}
+
+%\setbeamerfont{subsection in toc}{size=fontsize}
+
+%\setbeamerfont{subsubsection in toc}{size=fontsize}
 
@@
 
-\newcommand{\craycs}{Cray~CS300}
+\newcommand{\uhhpc}{UH-HPC}
 
@@
 
-\AtBeginSection[]
-{
-   \begin{frame}
-       \frametitle{Outline}
-       \tableofcontents[currentsection, hideallsubsections]
+%%
+%\setbeamertemplate{part page}{
+%\begin{beamercolorbox}[sep=8pt,center,wd=\textwidth]{part title}
+%\usebeamerfont{part title}\insertpart\par
+%\end{beamercolorbox}}
+
+%\newcommand{\OUTLINE}[1]{   \begin{frame}\frametitle{Outline}\tableofcontents[currentsection]\end{frame}
+%\newcommand{\OUTLINE}{
+	%\begin{frame}
+%%       \frametitle{Outline}
+       %\frametitle{{}}	
+			 %\partpage
+       %\tableofcontents[currentsection]
+       %%\tableofcontents[currentsection, hideallsubsections]
+   %\end{frame}
+%}
+\makeatletter
+\AtBeginPart{
+  \beamer@tocsectionnumber=0\relax
+  \setcounter{section}{0}
+}
+\makeatother
+
+\AtBeginSection[]{
+	\begin{frame}\
+       \frametitle{\insertpart~--~Overview}	
+       \tableofcontents[currentsection]
+       %\tableofcontents[currentsection, hideallsubsections]
    \end{frame}
@@
 \begin{frame}
-\frametitle{Outline} % Table of contents slide, comment this block out to remove it
-\tableofcontents[hideallsubsections] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
+\frametitle{Overview} % Table of contents slide, comment this block out to remove it
+Part I -- Introduction:
+\tableofcontents[part=1,hideallsubsections]
+Part II -- Using the UH-HPC:
+\tableofcontents[part=2,hideallsubsections]
+Part III -- High Level View of Policies:
+\tableofcontents[part=3,hideallsubsections]
+
+\tableofcontents[] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
 \end{frame}
@@
 \include{sections/01_intro}
-\include{sections/02_cluster}
-%\include{sections/03_sustainability}
-\include{sections/04_cluster_interaction}
-\include{sections/05_reporting}
-\include{sections/06_ediquette}
-\include{sections/07_policies}
+\include{sections/02_interaction}
+\include{sections/03_policies}
+%\include{sections/04_cluster_interaction}
+%\include{sections/05_reporting}
+%\include{sections/06_ediquette}
+%\include{sections/07_policies}
 \include{sections/08_end}
 %\include{sections/09_related_services}
-\include{sections/10_faq}
+%\include{sections/10_faq}
 

### 2019-05-25 hpc_101/sections/01_intro.tex
switch over to new presentation for new cluster

@@
-\section{Introduction}
+\part{Introduction}
+\begin{frame}
+			 \partpage
+\end{frame}
 
-\subsection{Parallel Computing}
+\section[Terminology]{Terminology}
+
 \begin{frame}
-	\frametitle{Parallel Computing}
+	\frametitle{Terminology}
 	\begin{itemize}
-		\item High Performance Compute
+	\item \textbf{Node} -- Another name for a server or computer
+        \item \textbf{Login node} -- A specialized node that users connect to in order to submit work to a computer cluster
+        \item \textbf{Computer cluster} -- A set of loosely or tightly connected nodes that work together so that, in many respects, they can be viewed as a single system\footnote{\label{wiki_ccluster}\tiny\url{https://en.wikipedia.org/wiki/Computer_cluster}}
+        \item \textbf{Data transfer node (DTN)} -- Specialized nodes that minimize the impedance on the network to access the full capability of the network
+	\item \textbf{Science DMZ (SciDMZ)} --  A portion of the network configured with equipment and security policies in order to optimize for high-performance scientific applications rather than for general-purpose business systems or “enterprise” computing
+	\item \textbf{Multi-factor Authentication (MFA)} -- An authentication method in which a computer user is granted access only after successfully presenting two or more pieces of evidence or factors to an authentication mechanism, e.g., DUO
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Terminology}
 	\begin{itemize}
-			\item Each separate process can send and receive data amongst other processes (MPI \& OpenMP)
-			\item If processes must communicate for the overall program to proceed, high-speed networking is needed (only MPI)
+	\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
+	\item \textbf{Command-line interface (CLI)} -- A text-based user interface used to view and manage computer files 
+	\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
+	\item \textbf{High Performance Compute (HPC)} -- A computing paradigm in which applications are typically a tightly coupled parallel job that benefit from a low-latency interconnect
+	\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
+%	      \item Pleasantly Parallel -- Processes are independent and no communication is necessary 
 	\end{itemize}
-		\item High Throughput Compute
+\end{frame}
+
+\section[University of Hawai'i High Performance Compute Cluster]{University of Hawai'i High Performance Compute Cluster}
+\begin{frame}
+    \frametitle{University of {\hawaii} High Performance Compute Cluster}
     \begin{itemize}
-			\item \emph{Pleasantly Parallel} -- Processes are independent and no communication is necessary 
+    \item The University~of~{\hawaii} High Performance Compute Cluster~({\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
+    \item Community acquired nodes are equally accessible to all users
+    \item Nodes purchased through the {\textbf{condo program}} are shared with the community, but priority is given to the node owner and their agents
+		\item Nodes may be {\textbf{leased}} from the community pool
+		\item Additional permanent storage can be leased with up to a five year contract by faculty \& staff
+    \end{itemize}
+		\begin{block}{UH-HPC Resource Summary}
+  \begin{table}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tabular}{l||l||l||l||l||l||l||l}
+      %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\ 
+			\toprule
+
+              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & Storage for lease \\
+							\midrule  \midrule %\hline
+        \thead{\textbf{Total}} & 297                    & 6,308                      & 50 TB                   &  56                   & 80 TB                             &  700 TB                        & 1 PB              \\ %\hline
+				\bottomrule
+      \end{tabular}%
+    }
+  \end{table}
+	\end{block}
+\end{frame}
+
+\subsection{Condo Program \& Leasing}
+\begin{frame}
+  \frametitle{What is the Condo Program \& leasing?}
+	\begin{block}{Condo Program}\scriptsize
+		\begin{itemize}
+		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with the UH-HPC
+		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
+		\item Condo nodes are managed and maintained by ITS staff
+		\item No maintenance fee will be assessed until the node is off warranty (typically five years)
+		\end{itemize}
+	\end{block}
+	\begin{block}{Leasing}\scriptsize
+		\begin{itemize}
+		\item Nodes leased from the community can have a contracts period from one month, up to one year
+		\item Node lessees are provided priority access to their leased hardware
+		\item Leases are not considered an equipment purchase
+		\end{itemize}
+	\end{block}
+\end{frame}
+
+\subsection{Storage}
+\begin{frame}
+  \frametitle{Storage}
+  The UH-HPC has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
+	\begin{columns}
+		\begin{column}{0.46\textwidth}
+		\begin{block}{Free Storage}
+			\begin{enumerate}
+			\item Permanent Storage
+				\begin{itemize}
+				\item Home Storage
+				\item Group/Lab Storage
+				\end{itemize}
+			\item Scratch Storage
+				\begin{itemize}
+				\item {\lustre (Aug 2019)}
+				\item Network File System (NFS)
+				\end{itemize}
+		\end{enumerate}
+		\end{block}
+		\end{column}
+		\begin{column}{0.46\textwidth}
+		\begin{block}{For Fee Storage}
+			\begin{enumerate}
+				\item ValueStorage
+				\item Long Term Storage (LTS)
+		\end{enumerate}
+		\end{block}
+		\end{column}
+	\end{columns}      
+\end{frame}
+
+\begin{frame}
+  \frametitle{Permanent Storage}
+  \begin{enumerate}
+  \item Home Storage
+    \begin{itemize}
+    \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on the UH-HPC
+    \end{itemize}
+  \item Group/Lab Storage
+    \begin{itemize}
+    \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on the UH-HPC
+    \end{itemize}
+  \end{enumerate}
+~\\
+  All permanent storage options on the UH-HPC have the following attributes:
+  \begin{itemize}
+  \item 50 GB default quota with a max quota of 300 GB
+  \item Quota increases are re-evaluated annually
+  \item Available on all nodes
+  \item Freely available to users
+  \end{itemize}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Scratch Storage}
+  \begin{enumerate}
+    \item {\lustre (Aug 2019)}
+      \begin{itemize}
+      \item High performance parallel file system 
+      \item 50 TB quota
+      \end{itemize}
+    \item NFS
+      \begin{itemize}
+      \item Not a parallel file system
+      \item Potentially better performance for random I/O patterns
+      \item 5 TB quota
+      \end{itemize}
+  \end{enumerate}
+~\\
+  All scratch file systems on the UH-HPC have the following attributes:
+  \begin{itemize}
+  \item \textbf{Purge policy} -- 10 days based on file modify time
+  \item Available on all nodes
+  \item Freely available to users
+  \end{itemize}
+
+\end{frame}
+
+%\begin{frame}
+  %\frametitle{For Fee Storage}
+  %\begin{enumerate}
+    %\item ValueStorage
+      %\begin{itemize}
+      %\item Enterprise class equipment 
+      %\item 10 Gbit connected to the UH-HPC
+      %\item Currently only accessible on the DTNs and login nodes
+      %\end{itemize}
+    %\item LTS  
+      %\begin{itemize}
+      %\item Cheap-n-Deep
+      %\item 100 Gbit connected to the UH-HPC
+      %\item Accessible on all nodes
+      %\end{itemize}
+  %\end{enumerate}
+%\end{frame}
+
+\subsection{Networking}
+\begin{frame}
+  \frametitle{Networking}
+	\begin{columns}
+	\begin{column}{0.46\textwidth}
+		\begin{block}{Internal Networks}
+		\begin{itemize}
+			\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
+				\begin{itemize}
+					\item 40 Gbit
+					\item Older compute nodes
+					\item low latency ($\approx1.3\mu$s)
+					\item non-blocking
+				\end{itemize}
+				~\\
+			\item 25/100 Gbit Ethernet
+				\begin{itemize}
+					\item Newer compute nodes
+					\item Nodes connected @ 25~Gbit
+					\item non-blocking
 				\end{itemize}
 	        \end{itemize}
+                \end{block}
+	\end{column}
+	\begin{column}{0.46\textwidth}
+		\begin{block}{External Networks}
+		\begin{itemize}
+			\item 100 Gbit SciDMZ
+				\begin{itemize}
+					\item Login \& compute nodes\\connected via a firewall
+					\item DTNs are directly connected
+				\end{itemize}
+		\end{itemize}
+                \end{block}
+	\end{column}
+	\end{columns}
 \end{frame}
+
+\subsection{User Support}
+\begin{frame}
+  \frametitle{User Support}
+
+  \begin{block}{Online docuements \& FAQ}
+    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeated questions and we try to capture them in our FAQ~\\
+    \href{https://www.hawaii.edu/its/ci/xcat/}{xCAT cluster information, policies \& FAQ}
+  \end{block}
+  \begin{block}{Contact Information}
+    If your questions is not answered in our online documentation, please contact us at: UH-HPC-Help@lists.hawaii.edu
+
+    \begin{itemize}
+    \item For batch jobs \ldots
+      \begin{itemize}
+      \item[--] Job ID, path to submission script, submission command, error file location, output files
+      \end{itemize} 
+    \item For other problems \ldots
+      \begin{itemize}
+      \item[--] State the problem, command issued, host, directory, remote host, error messages
+      \end{itemize}
+    \end{itemize}
+  \end{block}
+\end{frame}

### 2019-05-25 hpc_101/sections/03_policies.tex
switch over to new presentation for new cluster

\part{High Level View of Policies}
\begin{frame}
			 \partpage
\end{frame}

%\section[Resource and Data Governance Policies]{Resource and Data Governance Policies}

%% \begin{frame}
%% 	\frametitle{Overview}
%% 	\begin{itemize}
%% 		\item Login node usage policy
%% 		\item Scratch filesystem purge policy
%% 	\end{itemize}
%% \end{frame}

\section[UH Policies]{UH Policies}
\begin{frame}
  \frametitle{UH Policies}
	\begin{block}{Chapter 708, Hawaii Revised Statutes}
	\begin{itemize}
	\item Access only by authorized people
	\item Should not be used in the act of committing a crime
	\end{itemize}
	\end{block}
	\begin{block}{University of Hawaii Executive Policy E2.210} 
	\begin{itemize}
	\item Protect your password (we will never ask or require your password)
	\item Computer resources should not be used to test or compromise systems without prior authorization
	\item University resources are intended to be used for institutional purposes and may not be used for private gain
	\end{itemize}
	\end{block}
\end{frame}

\section[UH-HPC Usage Policies]{UH-HPC Usage Policies}

\subsection{Login nodes \& DTN nodes}
\begin{frame}
\frametitle{Login nodes \& DTN nodes}
\begin{block}{}
The following actions are acceptable on the shared systems in the UH-HPC:
\begin{enumerate}
\item File/Directory management [ Login \& DTN nodes ]
\item Text editing with a text editor: vi/vim, emacs, nano [ Login \& DTN nodes ]
\item Transferring files to and from the cluster (scp, rsync, SFTP, globus, aspera, lftp, etc.) [ Login \& DTN nodes ]
\item Shall be used to submit batch and interactive jobs [Login nodes]
\item SSH shell access [ Login \& DTN nodes ]
\end{enumerate}
\end{block}

\begin{block}{}
All other action shall take place on a compute node using an interactive session.  
If any actions outside of the sanctioned activities are detected, the following escalation will take place:
\begin{enumerate}
\item The process will be killed without notice
\item If the user continues this action, we will notify the user of violating the policy
\item If the user continues to ignore our warnings, the user may be banned from the cluster for a duration of time
\end{enumerate}
\end{block}
\end{frame}

\subsection{Storage Policy}
\begin{frame}
\frametitle{Storage Policy}
\begin{enumerate}
\item ITS is not responsible for any data that is deleted or loss due to user error, hardware failure, administrator error.  Users are responsible for their own data and are highly encouraged to backup data that is important to them at other storage locations
\item Files located on the scratch file systems are subject to a \textbf{10 day} purge policy that is based on the file/directory modified timestamp.  \textbf{Files that are purge cannot be recovered}.  Users are encouraged to copy their results off the scratch file system as soon as possible upon completion of their job
\item  Home storage currently snapshots once a day per user in case of accidental file deletion
\end{enumerate}
\end{frame}
%\btVFill
%\begin{center}
%\footnotesize \textbf{\emph{Login node usage policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
%\end{center}

\subsection{User Account Life Cycle}
\begin{frame}
\frametitle{User Account Life Cycle}
\begin{enumerate}
\item Accounts that are idle for 120 days will be locked
\item Accounts that have been locked can be unlocked upon request with no further action then emailing us a request to unlock their account
\item Accounts that are idle for a total of 180 days will be purged from the cluster
\item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use the UH-HPC
\item Users that become \textbf{Ohana} (graduated or no longer affiliated with UH) will be allowed up to one year on the UH-HPC before their accounts will be purged  
\item Ohana accounts can be purged sooner based on:
\begin{itemize}
\item Standard idle lockout/purge policy   
\item A faculty or staff requesting a graduated student's account be removed sooner than the one year grace period
\end{itemize}
\item Potential users whom are UH Ohana at the time of account request/creation may not have an account created on the UH-HPC without confirmation from an active faculty member or staff
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Where to find all UH-HPC specific policies}
\begin{enumerate}
\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes}
\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{University of Hawaii Executive Policy E2.210}
\item \href{http://go.hawaii.edu/GSY}{Common systems \& Storage}
\item \href{http://go.hawaii.edu/0SG}{User Account life cycle}
\item \href{http://go.hawaii.edu/WSG}{Security}
\item \href{http://go.hawaii.edu/YKG}{For lease storage}
\item \href{http://go.hawaii.edu/GK0}{Condo Program \& Off Warranty condo nodes}
\item \href{https://docs.google.com/document/d/1mGcCAsmzGT_hWXLcUNfT2LUwiggCnhqrln32k0IGZZI/}{Special allocations}
\end{enumerate}
\end{frame}
%
%\begin{frame}
%\frametitle{Login Node Usage Policy}\footnotesize
%The login nodes are a shared resource and are the only access to the cluster for hundreds of user and are meant to provide the following functionality for all users: 
%\begin{itemize}
%\item Providing ssh shell access 
%\item Facilitate the transfer files to and from the cluster:\\Globus, sftp, scp, rsync
%\item Launching and monitoring SLURM jobs (batch and interactive)
%\item Modifying text files with a text editor: vi/vim, emacs, nano
%\end{itemize}
%\bigskip
%\begin{itemize}
%\item[--] If we identify or are notified that a user is running computation on the login nodes, the application will be killed
%\item[--] If we determine that a user repeatedly violates the login node policy, even after being warned, the repeat offender can have their cluster account disabled
%\end{itemize}
%\btVFill
%\begin{center}
%\footnotesize \textbf{\emph{Login node usage policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
%\end{center}
%\end{frame}
%
%
%\subsection{Lustre Filesystem Purge Policies}
%\begin{frame}
%\frametitle{{\lustre} Filesystem Purge Policies}
%Due to users not having any quotas on the {\lustre} filesystem, we need some policy in place which removes older files from the system.  To accomplish this, we have implemented a purge policy.
%
%\begin{itemize}
%\item What in \ctilde/lus/ is subject to the purge policy?
   %\begin{itemize}
   %\item \textbf{Answer:} Everything but directories.
   %\end{itemize}
%\item When will my files be purged?
  %\begin{itemize}  
    %\item \textbf{Answer:} Files greater than or equal to 1 Megabyte in size or are empty (0 bytes) that are 35 days old
    %\item \textbf{Answer:} Files between 1 byte and 1 Megabyte in size that are 120 days old
  %\end{itemize}
%\item What is the age based on?
   %\begin{itemize}
   %\item \textbf{Answer:} Age of file is based on creation time
   %\end{itemize}
   %
%%\item How frequently are files for purged?
%%   \begin{itemize}
%%   \item \textbf{Answer:} Taking into account the fill rate of the {\lustre} filesystem, a purge is performend as usage approaches 80\%.
%%   \end{itemize}
%\end{itemize}
%\btVFill
%\begin{center}
%\footnotesize \textbf{\emph{Purge policy is subject to change~\\Users will be notified via email prior to changes taking effect}}
%\end{center}
%\end{frame}
%
%
%
%\begin{frame}
  %\frametitle{How to check file age for purge}
  %A tool is available on the cluster called age\_check.  This tool will list all files that are older than the defined date you set.
%\begin{semiverbatim}\tiny \texttt
  %$[$root@login-0001 \ctilde$]$\$ age\_check -h
  %
  %Usage:
  %age\_check [-s] [-z] [-d] [-f] [-a <age cutoff>] [-p <path]
%
%
  %Options:
%
  %-s -- Sort output
  %
  %-d -- display creation date per file
  %
  %-z -- display size in bytes per file
  %
  %-f -- Do not exclude files between 1 byte and 1 MB
  %
  %-a age -- Lists files that were created at least age days ago (default is 20 days)
  %
  %-p path -- Directory to check in (default is /lus/scratch/root)
%
  %
  %Example:
  %age\_check -a 30  
%
%\end{semiverbatim}
%
%\end{frame}

### 2019-06-13 hpc_101/sections/01_intro.tex
corrections

@@
 
-  \begin{block}{Online docuements \& FAQ}
-    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeated questions and we try to capture them in our FAQ~\\
+  \begin{block}{Online documents \& FAQ}
+    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ~\\
     \href{https://www.hawaii.edu/its/ci/xcat/}{xCAT cluster information, policies \& FAQ}
@@
   \begin{block}{Contact Information}
-    If your questions is not answered in our online documentation, please contact us at: UH-HPC-Help@lists.hawaii.edu
+    If your question is not answered in our online documentation, please contact us at: UH-HPC-Help@lists.hawaii.edu
 

### 2019-06-15 hpc_101/sections/01_intro.tex
revised slides to correct wording and reduce number of slides

@@
 	\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
-	\item \textbf{Command-line interface (CLI)} -- A text-based user interface used to view and manage computer files 
+	\item \textbf{Command-line interface/interpreter (CLI)} -- A text-based user interface used to view and manage computer files 
 	\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
@@
 	\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
+        \item \textbf{Modified time} -- The last time the file was modified (content has been modified)\footnote{\label{Types of timestamps}\tiny\url{https://unix.stackexchange.com/a/2465}}
+        \item \textbf{Shell script (script)} -- A computer program designed to be run by a CLI\footnote{\label{Shell script}\tiny\url{https://en.wikipedia.org/wiki/Shell_script}}
 %        \item Pleasantly Parallel -- Processes are independent and no communication is necessary 
 	\end{itemize}
+
 \end{frame}

### 2019-06-25 hpc_101/cheatsheet/paper.tex
fixed typo on globus endpoint and swtiched from tmi to shm:tmi
Removed names from cheatsheet, and modified the array view to look correct compared to slurm squeue output
comment out 2ndary login node
updates and corrections

--- 2016-01-22 hpc_101/cheatsheet/paper.tex
@@
 \date{\today}
-\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.\\~\\University of {\hawaii}\\Information Technology Services -- {\ci}}
+\author{University of {\hawaii}\\Information Technology Services -- {\ci}}
 \maketitle
@@
 \item[] \texttt{srun} \desc{Used in submitting an interactive job, or also  used within a slurm\_script}
-\item[] \texttt{srun.x11} \desc{Used in submitting an interactive job, that also forwards the X, if you logged in while enabling X forwarding}
 \item[] \texttt{squeue} \desc{Display all jobs running or waiting on the cluster}
@@
 \begin{itemize}
-\item[] login-0001 -- \texttt{uhhpc1.its.hawaii.edu}
-\item[] login-0002 -- \texttt{uhhpc2.its.hawaii.edu}
+\item[] login node -- \texttt{uhhpc.its.hawaii.edu}
+\item[] dtn 1 node -- \texttt{hpc-dtn1.its.hawaii.edu}
+\item[] dtn 2 node -- \texttt{hpc-dtn2.its.hawaii.edu}
+
+  %% \item[] login-0002 -- \texttt{uhhpc2.its.hawaii.edu}
 \end{itemize}
@@
 \begin{itemize}
-\item[] login-0001 -- \texttt{hawaii\#UHHPC1}
-\item[] login-0002 -- \texttt{hawaii\#UHHPC2}
+\item[] \texttt{UH-HPC}
+%% \item[] login-0002 -- \texttt{UH-HPC2}
+\end{itemize}
+
+\subsection{Downtime}
+The UH-HPC has a downtime scheduled for the first Wednesday of each month
+%% \item[] login-0002 -- \texttt{UH-HPC2}
+
+\subsection{Updated Example Scripts}
+Users may obtain updated example submission scripts from ``/mnt/config/examples/''
+
+%% \item[] login-0002 -- \texttt{UH-HPC2}
+
+\subsection{UH-HPC links}
+\begin{itemize}
+\item[]  \href{https://www.hawaii.edu/its/ci/xcat/}{General Information}
+\item[] \href{http://go.hawaii.edu/jdG}{FAQ}
 \end{itemize}

### 2020-01-16 hpc_101/sections/04_security.tex
starting of security and data governance section

\part{Data Governance \& Security}
\begin{frame}
			 \partpage
\end{frame}

\section[Data Governance]{Data Governance}

\begin{frame}
  \frametitle{What is data governance}
  %        \begin{definition}\footnotesize
  %          Data governance is a data management concept concerning the capability that enables an organization to ensure that high data quality exists throughout the complete lifecycle of the data. The key focus areas of data governance include availability, usability, consistency, data integrity and data security and includes establishing processes to ensure effective data management throughout the enterprise such as accountability for the adverse effects of poor data quality and ensuring that the data which an enterprise has can be used by the entire organization.
  %        \footnote[1,frame]{\tiny \href{https://en.wikipedia.org/wiki/Data\_governance}{https://en.wikipedia.org/wiki/Data\_governance}}
  %        \end{definition}
  \begin{itemize}
  \item Defines how data are collected, stored, and used
  \item Defines who can access data, when, and under what conditions
  \item Establishes decision rights
  \item Establishes clear lines of accountability
  \item Gives a voice to all appropriate parties
  \item Provides a mechanism for conflict resolutions involving data
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Goals of data governance @ UH}  
  \begin{itemize}
  \item Protect the privacy and security of ``Protected Data''
    \begin{itemize}
    \item All non-public data
    \item Institutional data
    \item Research data
    \end{itemize}
  \item Produce higher quality data for informed decision making
  \item Promote efficient use of resources
  \item Increase transparency and accountability
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Data Classifications @ UH}
  \begin{table}
    \centering
    \resizebox{\textwidth}{!}{%
      \begin{tabular}{l||p{6cm}||p{6cm}}
      %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\ 
	\toprule                                                                    
        \thead{\large\textbf{Category}} & \thead{\large\textbf{Definition}} & \thead{\large\textbf{Examples}} \\ 
	\midrule  \midrule %\hline
        \textbf{Public} & Access is not restricted and is subject to open records requests  & Student directory information, employee’s business contact info \\
	\midrule  \midrule %\hline
        \multicolumn{3}{c}{ {\large {\textbf{Protected Data} } } } \\ 
	\midrule  \midrule %\hline
        \textbf{Restricted} & Used for UH business only; will not be distributed to external parties; released externally only under the terms of a written MOA or contract & Student contact information, UH ID number \\
	\midrule  
        \textbf{Sensitive} & Data subject to privacy considerations  & Date of birth, job applicant records, salary/payroll information, most student information \\
	\midrule  
        \textbf{Regulated} & Inadvertent disclosure or inappropriate access requires a breach notification by law or is subject to financial fines & FN or first initial/LN in combination with SSN, driver license number, or bank information; credit card, HIPAA, or financial aid information \\            
    \bottomrule
      \end{tabular}%
    }
  \end{table}
\end{frame}  
	%% \begin{columns}
	%% \begin{column}{0.46\textwidth}
	%% 	\begin{block}{}
	%% 	\begin{itemize}
	%% 		\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
	%% 			\begin{itemize}
	%% 				\item 40 Gbit
	%% 				\item Older compute nodes
	%% 				\item low latency ($\approx1.3\mu$s)
	%% 				\item non-blocking
	%% 			\end{itemize}
	%% 			~\\
	%% 		\item 25/100 Gbit Ethernet
	%% 			\begin{itemize}
	%% 				\item Newer compute nodes
	%% 				\item Nodes connected @ 25~Gbit
	%% 				\item non-blocking
	%% 			\end{itemize}
	%%         \end{itemize}
        %%         \end{block}
	%% \end{column}
	%% \begin{column}{0.46\textwidth}
	%% 	\begin{block}{External Networks}
	%% 	\begin{itemize}
	%% 		\item 100 Gbit SciDMZ
	%% 			\begin{itemize}
	%% 				\item Login \& compute nodes\\connected via a firewall
	%% 				\item DTNs are directly connected
	%% 			\end{itemize}
	%% 	\end{itemize}
        %%         \end{block}
	%% \end{column}
	%% \end{columns}

%% \begin{frame}
%% 	\frametitle{Terminology}
%% 	\begin{itemize}
%% 	\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
%% 	\item \textbf{Command-line interface/interpreter (CLI)} -- A text-based user interface used to view and manage computer files 
%% 	\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
%% 	\item \textbf{High Performance Compute (HPC)} -- A computing paradigm in which applications are typically a tightly coupled parallel job that benefit from a low-latency interconnect
%% 	\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
%%         \item \textbf{Modified time} -- The last time the file was modified (content has been modified)\footnote{\label{Types of timestamps}\tiny\url{https://unix.stackexchange.com/a/2465}}
%%         \item \textbf{Shell script (script)} -- A computer program designed to be run by a CLI\footnote{\label{Shell script}\tiny\url{https://en.wikipedia.org/wiki/Shell_script}}
%% %        \item Pleasantly Parallel -- Processes are independent and no communication is necessary 
%% 	\end{itemize}

%% \end{frame}

%% \section[University of Hawai'i High Performance Compute Cluster]{University of Hawai'i High Performance Compute Cluster}
%% \begin{frame}
%%     \frametitle{University of {\hawaii} High Performance Compute Cluster}
%%     \begin{itemize}
%%     \item The University~of~{\hawaii} High Performance Compute Cluster~({\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
%%     \item Community acquired nodes are equally accessible to all users
%%     \item Nodes purchased through the {\textbf{condo program}} are shared with the community, but priority is given to the node owner and their agents
%% 		\item Nodes may be {\textbf{leased}} from the community pool
%% 		\item Additional permanent storage can be leased with up to a five year contract by faculty \& staff
%%     \end{itemize}
%% 		\begin{block}{UH-HPC Resource Summary}
%%   \begin{table}
%%     \centering
%%     \resizebox{\textwidth}{!}{%
%%       \begin{tabular}{l||l||l||l||l||l||l||l}
%%       %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\ 
%% 			\toprule

%%               {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & Storage for lease \\ 
%% 							\midrule  \midrule %\hline
%%         \thead{\textbf{Total}} & 297                    & 6,308                      & 50 TB                   &  56                   & 80 TB                             &  700 TB                        & 1 PB              \\ %\hline
%% 				\bottomrule
%%       \end{tabular}%
%%     }
%%   \end{table}
%% 	\end{block}
%% \end{frame}

%% \subsection{Condo Program \& Leasing}
%% \begin{frame}
%%   \frametitle{What is the Condo Program \& leasing?}
%% 	\begin{block}{Condo Program}\scriptsize
%% 		\begin{itemize}
%% 		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with the UH-HPC
%% 		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
%% 		\item Condo nodes are managed and maintained by ITS staff
%% 		\item No maintenance fee will be assessed until the node is off warranty (typically five years)
%% 		\end{itemize}
%% 	\end{block}
%% 	\begin{block}{Leasing}\scriptsize
%% 		\begin{itemize}
%% 		\item Nodes leased from the community can have a contract period from one month, up to one year
%% 		\item Node lessees are provided priority access to their leased hardware
%% 		\item Leases are not considered an equipment purchase
%% 		\end{itemize}
%% 	\end{block}
%% \end{frame}

%% \subsection{Storage}
%% \begin{frame}
%%   \frametitle{Storage}
%%   The UH-HPC has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
%% 	\begin{columns}
%% 		\begin{column}{0.46\textwidth}
%% 		\begin{block}{Free Storage}
%% 			\begin{enumerate}
%% 			\item Permanent Storage
%% 				\begin{itemize}
%% 				\item Home Storage
%% 				\item Group/Lab Storage
%% 				\end{itemize}
%% 			\item Scratch Storage
%% 				\begin{itemize}
%% %				\item {\lustre}(TBD)
%% 				\item Network File System (NFS)
%% 				\end{itemize}
%% 		\end{enumerate}
%% 		\end{block}
%% 		\end{column}
%% 		\begin{column}{0.46\textwidth}
%% 		\begin{block}{For Fee Storage}
%% 			\begin{enumerate}
%% 				\item Long Term Storage (LTS)
%% 		\end{enumerate}
%% 		\end{block}
%% 		\end{column}
%% 	\end{columns}      
%% \end{frame}

%% \begin{frame}
%%   \frametitle{Permanent Storage}
%%   \begin{enumerate}
%%   \item Home Storage
%%     \begin{itemize}
%%     \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on the UH-HPC
%%     \end{itemize}
%%   \item Group/Lab Storage
%%     \begin{itemize}
%%     \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on the UH-HPC
%%     \end{itemize}
%%   \end{enumerate}
%% ~\\
%%   All permanent storage options on the UH-HPC have the following attributes:
%%   \begin{itemize}
%%   \item 50 GB default quota with a max quota of 300 GB
%%   \item Quota increases are re-evaluated annually
%%   \item Available on all nodes
%%   \item Freely available to users
%%   \end{itemize}

%% \end{frame}

%% \begin{frame}
%%   \frametitle{Scratch Storage}
%%   \begin{enumerate}
%% %    \item {\lustre (TBD)}
%% %      \begin{itemize}
%% %      \item High performance parallel file system 
%% %      \item 50 TB quota
%% %      \end{itemize}
%%     \item NFS
%%       \begin{itemize}
%% %      \item Not a parallel file system
%%       \item 5 TB quota
%%       \end{itemize}
%%   \end{enumerate}
%% ~\\
%%   All scratch file systems on the UH-HPC have the following attributes:
%%   \begin{itemize}
%%   \item \textbf{Purge policy} -- 10 days based on file modify time
%%   \item Available on all nodes
%%   \item Freely available to users
%%   \end{itemize}

%% \end{frame}

%% %\begin{frame}
%%   %\frametitle{For Fee Storage}
%%   %\begin{enumerate}
%%     %\item ValueStorage
%%       %\begin{itemize}
%%       %\item Enterprise class equipment 
%%       %\item 10 Gbit connected to the UH-HPC
%%       %\item Currently only accessible on the DTNs and login nodes
%%       %\end{itemize}
%%     %\item LTS  
%%       %\begin{itemize}
%%       %\item Cheap-n-Deep
%%       %\item 100 Gbit connected to the UH-HPC
%%       %\item Accessible on all nodes
%%       %\end{itemize}
%%   %\end{enumerate}
%% %\end{frame}

%% \subsection{Networking}
%% \begin{frame}
%%   \frametitle{Networking}
%% 	\begin{columns}
%% 	\begin{column}{0.46\textwidth}
%% 		\begin{block}{Internal Networks}
%% 		\begin{itemize}
%% 			\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
%% 				\begin{itemize}
%% 					\item 40 Gbit
%% 					\item Older compute nodes
%% 					\item low latency ($\approx1.3\mu$s)
%% 					\item non-blocking
%% 				\end{itemize}
%% 				~\\
%% 			\item 25/100 Gbit Ethernet
%% 				\begin{itemize}
%% 					\item Newer compute nodes
%% 					\item Nodes connected @ 25~Gbit
%% 					\item non-blocking
%% 				\end{itemize}
%% 	        \end{itemize}
%%                 \end{block}
%% 	\end{column}
%% 	\begin{column}{0.46\textwidth}
%% 		\begin{block}{External Networks}
%% 		\begin{itemize}
%% 			\item 100 Gbit SciDMZ
%% 				\begin{itemize}
%% 					\item Login \& compute nodes\\connected via a firewall
%% 					\item DTNs are directly connected
%% 				\end{itemize}
%% 		\end{itemize}
%%                 \end{block}
%% 	\end{column}
%% 	\end{columns}
%% \end{frame}

%% \subsection{User Support}
%% \begin{frame}
%%   \frametitle{User Support}

%%   \begin{block}{Online documents \& FAQ}
%%     Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ~\\
%%     \href{https://www.hawaii.edu/its/ci/xcat/}{xCAT cluster information, policies \& FAQ}
%%   \end{block}
%%   \begin{block}{Contact Information}
%%     If your question is not answered in our online documentation, please contact us at: UH-HPC-Help@lists.hawaii.edu

%%     \begin{itemize}
%%     \item For batch jobs \ldots
%%       \begin{itemize}
%%       \item[--] Job ID, path to submission script, submission command, error file location, output files
%%       \end{itemize} 
%%     \item For other problems \ldots
%%       \begin{itemize}
%%       \item[--] State the problem, command issued, host, directory, remote host, error messages
%%       \end{itemize}
%%     \end{itemize}
%%   \end{block}
%% \end{frame}

### 2020-01-17 hpc_101/sections/04_security.tex
addition of more secruity slides

@@
   \begin{itemize}
-  \item Defines how data are collected, stored, and used
+  \item Defines how data is collected, stored, and used
   \item Defines who can access data, when, and under what conditions
@@
 
+\subsection{Data Classifications @ UH}
 \begin{frame}
   \frametitle{Data Classifications @ UH}
-  \begin{table}
-    \centering
+  \vspace{-5pt}
+  \makebox[\linewidth][c]{
     \resizebox{\textwidth}{!}{%
@@
       \end{tabular}%
+    }    }
+  \vspace{-5pt}
+  \begin{block}{Protected Data}
+    Collecting, managing, sharing or using  data that can be classified in a protected data category the Data Governance Process (DGP) is required 
+  \end{block}
+\end{frame}
+
+\section{Regulations}
+
+\begin{frame}
+  \frametitle{Regulations}
+\begin{block}{}
+  The following lists of regulations are not exahustive lists that may apply to your data at the university (personal \& research).
+  It is important that you know and understand what regulations the data you are working with is subject to.
+
+\end{block}
+\end{frame}
+
+\subsection{Personally Identifiable Information (PII)  \& Financial Regulations}
+\begin{frame}
+  \frametitle{Regulations}
+  \begin{block}{Personally Identifiable Information (PII)  \& Financial Regulations}
+    \begin{itemize}
+    \item \href{https://studentprivacy.ed.gov/ferpa-regulations}{Family Educational Rights and Privacy Act (FERPA)} [UH Policy \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ap\&policyChapter=7\&policyNumber=022\&menuView=closed}{AP 7.022}]
+      \begin{itemize}
+      \item Federal law that protects the privacy of student education records
+      \end{itemize}
+
+    \item Higher Education Act (HEA)
+      \begin{itemize}
+      \item Federal law that protects the federal financial aid information
+      \end{itemize}
+
+    \item \href{https://www.ftc.gov/tips-advice/business-center/privacy-and-security/gramm-leach-bliley-act}{Gramm-Leach-Bliley Act (GLBA)}
+      \begin{itemize}
+      \item Federal law that requires finanical institutions to explain how they share \& protect customer's data
+      \end{itemize}
+
+    \item \href{https://gdpr-info.eu/}{General Data Protection Regulation (GDPR)}
+      \begin{itemize}
+      \item A European Union (EU) consumer protection law that applies to companies collecting PII as part of delivering goods and services
+      \end{itemize}
+
+    \item \href{Health Insurance Portability and Accountability Act (https://www.hhs.gov/hipaa/for-professionals/security/laws-regulations/index.html}{HIPPA}) [UH Policy \href{https://www.hawaii.edu/policy/index.php?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=217\&menuView=closed}{EP 2.217}]
+      \begin{itemize}
+      \item Federal law that protects the privacy of health information
+        \item \href{https://www.hawaii.edu/infosec/hipaa/}{HIPPA @ UH}
+      \end{itemize}
+    \end{itemize}  
+  \end{block}
+\end{frame}
+
+\subsection{Federal Regulations}
+\begin{frame}
+  \frametitle{Regulations}
+  \vspace{-5pt}
+  \begin{block}{Federal Regulations}
+    \begin{itemize}
+    \item \href{https://csrc.nist.gov/publications/detail/sp/800-171/rev-1/final}{National Institute of Standards and Technology Special Programs (NIST) 800-171}
+      \begin{itemize}
+      \item Federal Department of Defense (DoD) standards aimed at safeguarding Controlled Unclassified Information (CUI)
+      \item May apply to grants from DoD and possibly other federal agencies
+      \end{itemize}
+
+    \item \href{https://www.hsdl.org/?abstract\&did=461297}{National Industrial Security Program [DoD Directive 5220.22-M]}
+      \begin{itemize}
+      \item Classified data subject to regulation
+      \end{itemize}
+
+    \item \href{https://www.pmddtc.state.gov/ddtc_public?id=ddtc_public_portal_itar_landing\#tab-itar}{Export Control \& International Traffic in Arms Regulations (ITAR)}
+      \begin{itemize}
+      \item Federal regulations that impose access, dissemination or participation restrictions on the use and/or transfer of commodities, technical data, or the provision of services subject to United States (US) export controls for reasons of national security, foreign policy, anti-terrorism or non-proliferation
+      \end{itemize}
+
+    \item \href{https://www.bis.doc.gov/index.php/regulations/export-administration-regulations-ear}{Export Administration Regulations (EAR)}
+      \begin{itemize}
+      \item EAR govern whether a person may export a thing from the U.S., reexport the thing from a foreign country, or transfer a thing from one person to another in a foreign country.
+The EAR apply to physical things (sometimes referred to as "commodities") as well as technology and software.\footnote{\label{EAR on Wikipedia}\tiny\url{https://en.wikipedia.org/wiki/Export_Administration_Regulations}}
+      \end{itemize}
+    \end{itemize}
+\end{block}
+\end{frame}
+
+\subsection{State Regulations}
+\begin{frame}
+  \frametitle{Regulations}
+  \vspace{-5pt}
+  \begin{block}{State Regulations}
+    \begin{itemize}
+    \item \href{https://www.capitol.hawaii.gov/hrscurrent/Vol11_Ch0476-0490/HRS0487N/HRS_0487N-.htm}{{\hawaii} Revised Statues (HRS) 487N}
+      \begin{itemize}
+      \item State law that defines the breach notification to the legislature
+      \item Examples: First name or First initial and Last name in combination with
+        \begin{itemize}
+        \item Social Security Number (SSN)
+        \item Driver license or State ID \#
+        \item Person's finanical account information
+          \end{itemize}
+      \end{itemize}
+
+    \item \href{https://www.capitol.hawaii.gov/hrscurrent/Vol02_Ch0046-0115/HRS0092F/HRS_0092F-.htm}{Uniform Information Practices Act (UIPA)[HRS Chapter 92F]}
+      \begin{itemize}
+      \item Governs open records requests
+      \end{itemize}
+    \end{itemize}
+\end{block}
+\end{frame}
+
+\section{UH Data Related Policies \& Procedures}
+\begin{frame}
+  \frametitle{UH Data Related Policies \& Procedures}
+  \vspace{-7pt}
+\makebox[\linewidth][c]{
+    \resizebox{\textwidth}{!}{%
+      \begin{tabular}{l||l}
+      %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\
+        \toprule
+        \thead{\large\textbf{Policy}} & \thead{\large\textbf{Description}} \\
+        \midrule  \midrule %\hline
+        \multicolumn{2}{c}{ {\large {\textbf{Administrative Procedure} } } } \\  
+        \midrule \midrule
+        \textbf{AP2.215} & Mandatory Training \& Continuing Education Requirements \\
+        \midrule
+        \textbf{AP7.022} & FERPA \\
+        \midrule \midrule
+        \multicolumn{2}{c}{ {\large {\textbf{Executive Policy} } } } \\  
+        \midrule \midrule
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=210\&menuView=closed}{\textbf{EP2.210}} & Use and Management of Information Technology Resources \\
+        \midrule
+        \href{}{\textbf{EP2.213}} & System and Campus Wide Electronic Channels for Communicating with Students\\
+        \midrule        
+        \textbf{EP2.214} & Data Classification Categories \& Information Security Guidelines \\
+        \midrule
+        \textbf{EP2.215} & Institutional Data Governance \\
+        \midrule
+        \textbf{EP2.216} & Institutional Records Management \& Electronic Approvals/Signatures (\textit{Pending Approval}) \\
+        \midrule
+        \textbf{EP2.217} & HIPPA (\textit{To be Revised}) \\
+        \midrule
+        \textbf{EP2.218} & Online Approvals of Internal University Transactions\\
+        \midrule
+        \textbf{EP2.2xx} & Student Online Data Protection Requirements (\textit{Draft}) \\
+        \midrule
+        \textbf{EP2.215} & Institutional Data Governance \\
+        \midrule
+        \textbf{EP8.200} & Policy on Contracts \& Signing Authority \\
+        \bottomrule
+      \end{tabular}%
       }
-  \end{table}
+}
+\vspace{-5pt}
+
 \end{frame}
+
+  %% \item General Data Protection Regulation (GDPR)
+  %%   \begin{itemize}
+  %%   \item A European Union (EU) consumer protection law that applies to companies collecting personally identifiable information (PII) as part of delivering goods and services
+  %%   \end{itemize}
+
+  %% \item General Data Protection Regulation (GDPR)
+  %%   \begin{itemize}
+  %%   \item A European Union (EU) consumer protection law that applies to companies collecting personally identifiable information (PII) as part of delivering goods and services
+  %%   \end{itemize}
+
+  %% \item General Data Protection Regulation (GDPR)
+  %%   \begin{itemize}
+  %%   \item A European Union (EU) consumer protection law that applies to companies collecting personally identifiable information (PII) as part of delivering goods and services
+  %%   \end{itemize}
+
+  %% \item General Data Protection Regulation (GDPR)
+  %%   \begin{itemize}
+  %%   \item A European Union (EU) consumer protection law that applies to companies collecting personally identifiable information (PII) as part of delivering goods and services
+  %%   \end{itemize}
+
+  %% \end{itemize}
+
+  %% \begin{table}
+  %%   \centering
+  %%   \resizebox{\textwidth}{!}{%
+  %%     \begin{tabular}{l||p{6cm}||p{6cm}}
+  %%     %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\ 
+  %%       \toprule                                                                    
+  %%       \thead{\large\textbf{Regulation}} & \thead{\large\textbf{Description}} & \thead{\large\textbf{}} \\ 
+  %%       \midrule  \midrule %\hline
+  %%       \textbf{Public} & Access is not restricted and is subject to open records requests  & Student directory information, employee’s business contact info \\
+  %%       \midrule  \midrule %\hline
+  %%       \multicolumn{3}{c}{ {\large {\textbf{Protected Data} } } } \\ 
+  %%       \midrule  \midrule %\hline
+  %%       \textbf{Restricted} & Used for UH business only; will not be distributed to external parties; released externally only under the terms of a written MOA or contract & Student contact information, UH ID number \\
+  %%       \midrule  
+  %%       \textbf{Sensitive} & Data subject to privacy considerations  & Date of birth, job applicant records, salary/payroll information, most student information \\
+  %%       \midrule  
+  %%       \textbf{Regulated} & Inadvertent disclosure or inappropriate access requires a breach notification by law or is subject to financial fines & FN or first initial/LN in combination with SSN, driver license number, or bank information; credit card, HIPAA, or financial aid information \\
+  %%   \bottomrule
+  %%     \end{tabular}%
+  %%   }
+  %% \end{table}
+
 %% \begin{columns}

### 2020-01-18 hpc_101/sections/03_policies.tex
comment out a policy
Addition of a security section

@@
 \begin{enumerate}
-\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes}
-\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{University of Hawaii Executive Policy E2.210}
+%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes}
+%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{University of Hawaii Executive Policy E2.210}
 \item \href{http://go.hawaii.edu/GSY}{Common systems \& Storage}
@@
 \item \href{http://go.hawaii.edu/GK0}{Condo Program \& Off Warranty condo nodes}
-\item \href{https://docs.google.com/document/d/1mGcCAsmzGT_hWXLcUNfT2LUwiggCnhqrln32k0IGZZI/}{Special allocations}
+%% \item \href{https://docs.google.com/document/d/1mGcCAsmzGT_hWXLcUNfT2LUwiggCnhqrln32k0IGZZI/}{Special allocations}
 \end{enumerate}

### 2020-01-18 hpc_101/sections/04_security.tex
Addition of a security section

@@
   \frametitle{Goals of data governance @ UH}
+  \begin{block}{}
+    \href{https://www.hawaii.edu/offices/vp-academic-planning-policy/uh-institutional-data-governance/}{The Data Governance Office at the University of {\hawaii}} creates policies and procedures focused on the privacy and security of data under the University’s care.
+  \end{block}
+  \begin{block}{Goals}
   \begin{itemize}
@@
     \end{itemize}
-  \item Produce higher quality data for informed decision making
+%  \item Produce higher quality data for informed decision making
   \item Promote efficient use of resources
@@
   \end{itemize}
+  \end{block}
 \end{frame}
@@
     }    }
-  \vspace{-5pt}
-  \begin{block}{Protected Data}
-    Collecting, managing, sharing or using  data that can be classified in a protected data category the Data Governance Process (DGP) is required 
-  \end{block}
+  \vfill
+  \centering
+    The \href{https://datagov.intranet.hawaii.edu/dgp/}{Data Governance Process (DGP)} applies to collecting, managing, sharing or using data that can be classified as protected data
 \end{frame}
@@
 
-    \item \href{Health Insurance Portability and Accountability Act (https://www.hhs.gov/hipaa/for-professionals/security/laws-regulations/index.html}{HIPPA}) [UH Policy \href{https://www.hawaii.edu/policy/index.php?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=217\&menuView=closed}{EP 2.217}]
+    \item \href{https://www.hhs.gov/hipaa/for-professionals/security/laws-regulations/index.html}{Health Insurance Portability and Accountability Act (HIPPA}) [UH Policy \href{https://www.hawaii.edu/policy/index.php?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=217\&menuView=closed}{EP 2.217}]
       \begin{itemize}
@@
       \item State law that defines the breach notification to the legislature
-      \item Examples: First name or First initial and Last name in combination with
+      \item Examples: First name or First initial and Last name in combination with at least one of the following
         \begin{itemize}
@@
 
-\section{UH Data Related Policies \& Procedures}
+\subsection{UH Data Related Policies \& Procedures}
 \begin{frame}
@@
         \midrule \midrule
-        \textbf{AP2.215} & Mandatory Training \& Continuing Education Requirements \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ap\&policyChapter=2\&policyNumber=215\&menuView=closed}{\textbf{AP2.215}} & Mandatory Training \& Continuing Education Requirements \\
         \midrule
-        \textbf{AP7.022} & FERPA \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ap\&policyChapter=7\&policyNumber=022\&menuView=closed}{\textbf{AP7.022}} & FERPA \\
         \midrule \midrule
@@
         \midrule
-        \href{}{\textbf{EP2.213}} & System and Campus Wide Electronic Channels for Communicating with Students\\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=213\&menuView=closed}{\textbf{EP2.213}} & System and Campus Wide Electronic Channels for Communicating with Students\\
         \midrule        
-        \textbf{EP2.214} & Data Classification Categories \& Information Security Guidelines \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=214\&menuView=closed}{\textbf{EP2.214}} & Data Classification Categories \& Information Security Guidelines \\
         \midrule
-        \textbf{EP2.215} & Institutional Data Governance \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=215\&menuView=closed}{\textbf{EP2.215}} & Institutional Data Governance \\
         \midrule
-        \textbf{EP2.216} & Institutional Records Management \& Electronic Approvals/Signatures (\textit{Pending Approval}) \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=216\&menuView=closed}{\textbf{EP2.216}} & Institutional Records Management \& Electronic Approvals/Signatures (\textit{Pending Approval}) \\
         \midrule
-        \textbf{EP2.217} & HIPPA (\textit{To be Revised}) \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=217\&menuView=closed}{\textbf{EP2.217}} & HIPPA (\textit{To be Revised}) \\
         \midrule
-        \textbf{EP2.218} & Online Approvals of Internal University Transactions\\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=218\&menuView=closed}{\textbf{EP2.218}} & Online Approvals of Internal University Transactions\\
         \midrule
@@
         \midrule
-        \textbf{EP2.215} & Institutional Data Governance \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=2\&policyNumber=215\&menuView=closed}{\textbf{EP2.215}} & Institutional Data Governance \\
         \midrule
-        \textbf{EP8.200} & Policy on Contracts \& Signing Authority \\
+        \href{https://www.hawaii.edu/policy/?action=viewPolicy\&policySection=ep\&policyChapter=8\&policyNumber=200\&menuView=closed}{\textbf{EP8.200}} & Policy on Contracts \& Signing Authority \\
         \bottomrule
@@
 
+\end{frame}
+
+\section[Security]{Security}
+
+\subsection{Information Security Program}
+\begin{frame}
+  \frametitle{Information Security Program}
+
+  The \href{https://www.hawaii.edu/infosec/infosecprogram/}{University of {\hawaii} Information Security Program} is comprised of the following strategic areas:
+
+  \begin{block}{Strategic Areas}
+    \begin{itemize}
+  \item Data Governance \& Oversight
+  \item Information Security Audits \& Risk Assessments
+  \item Information Security Policies \& Procedures
+  \item Identity Management \& Access Controls
+  \item Information Security Training \& Awareness
+  \end{itemize}
+  \end{block}
+\end{frame}
+
+\subsection{Data Protection}
+\begin{frame}
+  \frametitle{Protect Your Data!}
+  \vspace{-10pt}
+  \begin{block}{Know which files have sensitive information}
+    \begin{itemize}
+    \item Use \href{http://www.hawaii.edu/askus/1297}{Spirion} (formerly Identity Finder) to scan for sensitive information
+    \end{itemize}
+  \end{block}
+  \vspace{-3pt}  
+  \begin{block}{Protect sensitive/regulated data as required by UH policy}
+  \begin{itemize}
+  \item UH Policy \href{https://www.hawaii.edu/policy/EP2.214}{EP 2.214}
+  \end{itemize}
+  \end{block}
+  \vspace{-3pt}  
+  \begin{block}{Data protection life cycle}
+    \begin{itemize}
+    \item Back up your data regularly
+    \item Transmit and share sensitive data in a secure manner, e.g., SFTP, \href{https://www.hawaii.edu/filedrop/}{UH FileDrop}
+    \item Protect sensitive/regulared data with encryption
+    \item Securely delete any sensitive data that is no longer needed
+    \end{itemize}
+  \end{block}    
+  \vspace{-3pt}  
+  \begin{block}{Physical Security of data repository with sensitive data}
+    \begin{itemize}
+    \item Paper documents should be stored in a secured location, not accessible by unauthorized individuals e.g., a locked file cabinent
+    \item External hard drives should be stored in a secured location e.g., a locked file cabinent
+    \end{itemize}
+  \end{block}    
+\end{frame}
+
+\subsection{Device Protection}
+\begin{frame}
+  \frametitle{Protect Your Computers!}
+  \begin{block}{Operating System (OS) }
+    \begin{itemize}
+    \item Apply updates to your OS and applications frequently and on a regular basis
+    \end{itemize}
+  \end{block}
+
+  \begin{block}{Protective Software}
+  \begin{itemize}
+  \item Install and keep up to date protective software such as \href{http://www.hawaii.edu/askus/1254}{anti-virus}
+  \end{itemize}
+  \end{block}
+
+  \begin{block}{Physical Access}
+    \begin{itemize}
+    \item Never leave your devices logged-in \& unattended
+    \item Control physical access to your machines
+    \item When you step away from your machine, lock your screen
+    \item Require password authentication when unlocking or accessing your machine
+    \end{itemize}
+  \end{block}    
+\end{frame}
+
+\subsection{Storing and transmitting data with Google products}
+\begin{frame}
+  \frametitle{Storing and transmitting data with Google products}
+  \onslide<1->{\begin{block}{Question}
+    \begin{columns}[T ]
+      \begin{column}{0.15\textwidth}\hfill\includegraphics[scale=0.05]{question_mark.png}\end{column}
+      \begin{column}{0.80\textwidth}
+        Is it safe to store and transmit sensitive or regulated data using the different Google products offerred at the University?
+        e.g., Google Docs, Google Drive, Gmail
+      \end{column}
+    \end{columns}
+    \end{block}
+    }
+  \vfill
+  \onslide<2->{\begin{block}{Answer}
+%https://www.kissclipart.com/stop-hand-clipart-stop-sign-clip-art-pgay2w/
+\begin{columns}
+\begin{column}{0.15\textwidth}\hfill\includegraphics[scale=0.05]{stop_hand.png}\end{column}
+\begin{column}{0.70\textwidth}
+    \centering
+    {\HUGE \textbf{\textit{NO!}}}\Large~\\Do \textbf{NOT} store or transmit sensitive or regulated information in \emph{Google Docs}, \emph{Google Drive} or \emph{Email}!
+\end{column}
+\begin{column}{0.15\textwidth}\includegraphics[scale=0.05]{stop_hand.png}\end{column}
+\end{columns}
+    \end{block}}
+\end{frame}
+
+\subsection{Account protection}
+\begin{frame}
+  \frametitle{Protect Your Account!}
+\begin{columns}
+\begin{column}{0.45\textwidth}
+\begin{block}{}
+  \begin{itemize}
+  \item User strong passwords and a strong password management strategy
+  \item \textbf{DO NOT RE-USE PASSWORDS!}
+  \item Consider using a passphrase
+    \begin{itemize}
+    \item lower/uppercase
+    \item numbers
+    \item special characters
+    \item character substitutions, e.g.,
+      \begin{itemize}
+      \item o $\Longrightarrow$ 0
+      \item a $\Longrightarrow$ @
+      \item i $\Longrightarrow$ 1
+      \end{itemize}
+    \end{itemize}
+  \item Check your password strength: \href{https://lastpass.com/howsecure.php}{https://lastpass.com/howsecure.php}
+  \item Monitor your accounts for suspicious activity
+  \end{itemize}
+  \end{block}
+\end{column}
+\begin{column}{0.49\textwidth}
+\begin{block}{}
+  \includegraphics[scale=0.25]{password_strength.png}
+  \numlessfootnotetxt{\label{xkcd_pasword_strnght}\tiny\url{https://xkcd.com/936/}}
+\end{block}
+\end{column}
+\end{columns}
+\end{frame}
+
+\section*{More Information}
+\begin{frame}
+  \frametitle{More Information}
+  You can find more information about Data Governance and Security at UH at the following links:
+  \begin{itemize}
+  \item \href{https://datagov.intranet.hawaii.edu/}{Data Governance internal site}
+  \item \href{https://datagov.intranet.hawaii.edu/training/}{Data Governance \& Security trainings/presentations}
+  \item \href{https://datagov.intranet.hawaii.edu/institutional-data-classification-levels/}{Institutional data classifications}
+  \item \href{https://www.hawaii.edu/infosec/}{Information Security (InfoSec) at UH}
+  \item \href{https://www.hawaii.edu/its/alerts/}{Security notices and other alerts from UH Information Technology Services (ITS)}
+  \item \href{https://www.hawaii.edu/its/itacw/}{IT All Campus Workshop \ddash Yearly during the summer}
+  \end{itemize}
 \end{frame}

### 2020-01-28 hpc_101/sections/04_security.tex
addition of uh filedrop

@@
     \centering
-    {\HUGE \textbf{\textit{NO!}}}\Large~\\Do \textbf{NOT} store or transmit sensitive or regulated information in \emph{Google Docs}, \emph{Google Drive} or \emph{Email}!
+    {\HUGE \textbf{\textit{NO!}}}\Large~\\Do \textbf{NOT} store or transmit sensitive or regulated information in \emph{Google Docs}, \emph{Google Drive} or \emph{Email}~\\~\\Use the \href{https://www.hawaii.edu/filedrop/}{UH FileDrop} to securely transmit files under 1000 megabytes inside
 \end{column}

### 2020-05-27 hpc_101/sections/01_intro.tex
updates and corrections
minor fixes
removed mention of valuestorage
starting of security and data governance section
update to slides for stale information

@@
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 297                    & 6,308                      & 50 TB                   &  56                   & 80 TB                             &  700 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 297                    & 6,308                      & 50 TB                   &  56                   & 80 TB                             &  150 TB                        & 1 PB              \\ %\hline
 				\bottomrule
@@
 		\begin{itemize}
-		\item Nodes leased from the community can have a contracts period from one month, up to one year
+		\item Nodes leased from the community can have a contract period from one month, up to one year
 		\item Node lessees are provided priority access to their leased hardware
@@
 				\begin{itemize}
-				\item {\lustre (Aug 2019)}
+%				\item {\lustre}(TBD)
 				\item Network File System (NFS)
@@
 			\begin{enumerate}
-				\item ValueStorage
 				\item Long Term Storage (LTS)
@@
 ~\\
-  All permanent storage options on the UH-HPC have the following attributes:
+  Permanent storage options on the UH-HPC have the following attributes:
   \begin{itemize}
@@
   \begin{enumerate}
-    \item {\lustre (Aug 2019)}
-      \begin{itemize}
-      \item High performance parallel file system 
-      \item 50 TB quota
-      \end{itemize}
+%    \item {\lustre (TBD)}
+%      \begin{itemize}
+%      \item High performance parallel file system 
+%      \item 50 TB quota
+%      \end{itemize}
     \item NFS
       \begin{itemize}
-      \item Not a parallel file system
-      \item Potentially better performance for random I/O patterns
+%      \item Not a parallel file system
       \item 5 TB quota
@@
 ~\\
-  All scratch file systems on the UH-HPC have the following attributes:
+  Scratch file systems on the UH-HPC have the following attributes:
   \begin{itemize}

### 2020-05-27 hpc_101/sections/03_policies.tex
update to slides for stale information

@@
 \item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use the UH-HPC
-\item Users that become \textbf{Ohana} (graduated or no longer affiliated with UH) will be allowed up to one year on the UH-HPC before their accounts will be purged  
-\item Ohana accounts can be purged sooner based on:
-\begin{itemize}
-\item Standard idle lockout/purge policy   
-\item A faculty or staff requesting a graduated student's account be removed sooner than the one year grace period
-\end{itemize}
-\item Potential users whom are UH Ohana at the time of account request/creation may not have an account created on the UH-HPC without confirmation from an active faculty member or staff
+\item Users that will be leaving UH or are no longer active at UH, will need to have an active UH faculty/staff member in good standing, request a UH-HPC account extension.
 \end{enumerate}

### 2020-09-10 OOD/sections/00_title.tex
addition of open ondemand slides

--- 2017-02-04 hpc_101/sections/00_title.tex
@@
 
-\title[HPC--101]{HPC--101\\Onboarding} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Open OnDemand]{Open OnDemand} % The short title appears at the bottom of every slide, the full title is only on the title page
 
-\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
+\author{}
+%\author{Director: Gwen Jacobs\\Sean Cleveland, Adriana Comerford, Jennifer Geis,\\Ron Merrill, David Schanzenbach}
 \institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
@@
 \textbf{\url{https://www.hawaii.edu/its/ci/}}\\
+\textbf{\url{http://datascience.hawaii.edu/}}\\
 \textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address

### 2020-09-10 OOD/sections/01_intro.tex
addition of open ondemand slides

\part{Background}
\begin{frame}
			 \partpage
\end{frame}

\section[What]{What is it?}

\begin{frame}
	\frametitle{What is Open OnDemand \& who is using it?}
%	\vspace{-2.5em}
	\begin{columns}
	\begin{column}{0.50\textwidth}

	\begin{block}{ What is it?}
	Open OnDemand, is a National Science Foundation (NSF) funded science gateway project (OAC1534949, OAC183575) from the Ohio Supercomputer Center, with the goal of making access to HPC resources simpler through a web browser.\footnote[1,frame]{\tiny \href{https://openondemand.org/}{https://openondemand.org/}}
	\end{block}
	\begin{block}{ Who is using it?}
	  \begin{minipage}[t][0.20\textheight][t]{\linewidth} 		
	\begin{itemize}
	\item Universities
	\item The Federal Reserve Bank
	\item Super computing centers
	\end{itemize}
	\end{minipage}
	\end{block}
	\end{column}
	\begin{column}{0.50\textwidth} %%
	\vspace{-2.5em}
	\begin{center}
  \includegraphics[scale=0.15]{05.png}
	\end{center}
	\end{column}
	\end{columns}
\end{frame}

\section[Why]{What can it do for me?}
\begin{frame}
	\frametitle{What can it do for me?}
	\begin{itemize}
	\item Provide a uniform environment which can help simplify instruction in a classroom setting
	\item Built-in file browser with in browser download/upload capabilities
		\begin{itemize}
		\item Max upload file size is currently limited to 5GB
		\end{itemize}
	\item In browser SSH terminal
	\item Job management and monitoring
	\item The ability to proxy graphical output from applications like Juypter notebooks and Rstudio Server executed on a compute node
	\end{itemize}
\end{frame}

### 2020-09-10 OOD/sections/02_walking.tex
addition of open ondemand slides

\part{Connecting \& Tour}
\begin{frame}
			 \partpage
\end{frame}

\section[Connecting]{Prerequisite \& Connecting}

\begin{frame}
\frametitle{Prerequisite}
		\begin{block}{Prerequisite}
			\begin{itemize}
				\item An active Mana (UH-HPC) account 
				\item A web browser 
			\end{itemize}
			\end{block}
			\begin{exampleblock}{Supported Browsers}
			\begin{itemize}
				\item Chrome\footnote[1,frame]{\tiny Chrome is the recommended browser}\footnote[2,frame]{\tiny \url{https://osc.github.io/ood-documentation/latest/requirements.html\#browser-requirements}}, Firefox, Edge are all supported
				\item Any modern browser that supports ECMAScript 2016 should work
		\end{itemize}
		\end{exampleblock}
			\begin{alertblock}{Unsupported Browsers}
			\begin{itemize}
				\item Internet Explorer 11
		\end{itemize}
				\end{alertblock}
\end{frame}

\begin{frame}
\frametitle{Connecting}
			\begin{itemize}
				\item Open up a browser window in incognito mode
					\begin{itemize}
					\item Optional but recommended for security upon logout
					\end{itemize}						
				\item Direct your browser to \url{https://uhhpc.its.hawaii.edu}
				\item Should be prompted for your UH credentials by the UH gold screen
				\item If successful, should be directed to a page with the Mana logo
			\end{itemize}
	\begin{center}
  \includegraphics[scale=0.15]{05.png}
	\end{center}			
\end{frame}

\section[WT]{Tour}
\begin{frame}
\Huge{\centerline{Tour}}
\end{frame}

### 2020-09-10 OOD/sections/03_custom.tex
addition of open ondemand slides

\part{Customization}
\begin{frame}
			 \partpage
\end{frame}

\section[Rstudio]{Rstudio Server}
\subsection[rstudiodef]{Defaults and Quirks}
\begin{frame}
  \frametitle{Defaults and Quirks}
	\begin{block}{Defaults}
		\begin{itemize}
			\item Rstudio Server will utilize the default R module unless told otherwise
				\begin{itemize}
					\item Current default is R 4.0.0
				\end{itemize}
			\item Only the dependencies the R module provides are loaded
		\end{itemize}
	\end{block}
	\begin{block}{Quirks}
			\begin{itemize}
			\item The terminal in Rstudio does not start up knowing the modules command
			\begin{itemize}
			\item User can correct this by doing the following commands in the terminal
			\item Changes in this terminals environment (modules loaded/purged) do not change what the R environment sees
			\item . \ctilde{}/.bash\_profile
			\item . /opt/ohpc/admin/lmod/lmod/init/sh
			\end{itemize}
		\end{itemize}
	\end{block}
\end{frame}

\subsection[rstudiodeps]{User defined environment}
\begin{frame}
	\frametitle{Lmod Collections}
	\begin{block}{}
	\begin{itemize}
		\item Lmod has the ability to save/create a named environment (Collections) which can be restored when needed
		\begin{itemize}
			\item For example, a user may create a build environment that loads multiple dependencies and they don't want to load them every single time or remember what they loaded and save it with a name
			\item The user would save the collection by doing "module save build\_enviro" 
			\item The user can then restore the collection by doing "module restore build\_enviro"
		\end{itemize}
		\item  We utilize this mechanism to allow a user to customize their Rstudio environment when using Open OnDemand and allow for the user to:
		\begin{itemize}
			\item Pin the R language version
			\item Load additional dependencies that some of their R libraries require, e.g., netCDF, gdal, CUDA
		\end{itemize}
	\end{itemize}
	\end{block}
\end{frame}

\subsection[rstudioenv]{Requirements \& procedure}
\begin{frame}
	\frametitle{Requirements \& procedure}
	\begin{block}{Requirements}
		\begin{itemize}
			\item The Lmod collection must contain a working R environment
				\begin{itemize}
					\item At a minimum it requires an R module to be loaded - lang/R
				\end{itemize}
			\item The collection must be named "rstudio"
		\end{itemize}
	\end{block}
	\begin{block}{Procedure}
		\begin{enumerate}
			\item Bring up a terminal on the cluster
			\item purge your current module environment
				\begin{itemize}
				\item module purge
				\end{itemize}
			\item Load modules you want to use with R studio
				\begin{itemize}
				\item module load lang/R/3.5.1-intel-2018.5.274-Python-2.7.15 
				\item module load data/GDAL/2.2.3-intel-2018.5.274-Python-2.7.15
				\end{itemize}
			\item Save the module environment to the name "rstudio"
				\begin{itemize}
				\item module save rstudio
				\end{itemize}
			\item Start up a new Rstudio instance and Rstudio should now have the additional libraries and be using the version of R you specified.
		\end{enumerate}
	\end{block}
\end{frame}

\section[Jupyternote]{Jupyter Notebooks}
\subsection[jupenv]{Procedure}
\begin{frame}[fragile]
	\frametitle{Procedure}
	\begin{block}{}
	The following steps were adapted from \url{https://github.com/PSC-PublicHealth/pha-nbextensions}.~\\The following commands should be executed from a compute node in an interactive session prior to starting up a Juypter Notebook in Open OnDemand.
	\end{block}
	\begin{semiverbatim}\footnotesize
	[\$] module load lang/Anaconda3
	[\$] conda create \ddash{}name test\_env python=3
	[\$] source activate test\_env
	\#\#\# Install your required packages \#\#\#
	\# \ldots 
	\# \ldots 
	\#\#\# Finalize setup with these last two packages \#\#\#
	[\$(test\_env)] conda install ipykernel
	[\$(test\_env)] pip install git+https://github.com/PSC-PublicHealth/pha-nbextensions.git
	\#\#\# DONE and ready for use in Open OnDemand \#\#\#
 	\end{semiverbatim}
	\vspace{-1.5em}
	\begin{center}
  \includegraphics[scale=0.08]{juypter.png}
	\end{center}

\end{frame}

### 2020-09-10 hpc_101/sections/01_intro.tex
addition of open ondemand slides

@@
 		\item Nodes may be {\textbf{leased}} from the community pool
-		\item Additional permanent storage can be leased with up to a five year contract by faculty \& staff
+		\item Additional permanent storage can be leased by faculty \& staff
     \end{itemize}
@@
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 297                    & 6,308                      & 50 TB                   &  56                   & 80 TB                             &  150 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 298                    & 6,308                      & 50 TB                   &  64                   & 80 TB                             &  150 TB                        & 1 PB              \\ %\hline
 				\bottomrule
@@
   \begin{block}{Online documents \& FAQ}
-    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ~\\
-    \href{https://www.hawaii.edu/its/ci/xcat/}{xCAT cluster information, policies \& FAQ}
+    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ.~\\
+		\begin{itemize}
+		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
+		\item \href{http://go.hawaii.edu/jdG}{FAQ}
+		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}
+		\end{itemize}
   \end{block}

### 2020-10-15 hpc_101/sections/01_intro.tex
Changed mentioned of UH-HPC to Mana Changed SLURM to Slurm as it is no longer an acronym.

@@
     \begin{itemize}
-    \item The University~of~{\hawaii} High Performance Compute Cluster~({\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
+    \item The University~of~{\hawaii} High Performance Compute Cluster - \textbf{\mana} (formerly~known~as~the~{\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
     \item Community acquired nodes are equally accessible to all users
@@
     \end{itemize}
-		\begin{block}{UH-HPC Resource Summary}
+		\begin{block}{Mana Resource Summary}
   \begin{table}
@@
 		\begin{itemize}
-		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with the UH-HPC
+		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with {\mana}
 		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
@@
   \frametitle{Storage}
-  The UH-HPC has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
+  {\mana} has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
 	\begin{columns}
@@
     \begin{itemize}
-    \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on the UH-HPC
+    \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on {\mana}
     \end{itemize}
@@
     \begin{itemize}
-    \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on the UH-HPC
+    \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on {\mana}
     \end{itemize}
@@
 ~\\
-  Permanent storage options on the UH-HPC have the following attributes:
+  Permanent storage options on {\mana} have the following attributes:
   \begin{itemize}
@@
 ~\\
-  Scratch file systems on the UH-HPC have the following attributes:
+  Scratch file systems on {\mana} have the following attributes:
   \begin{itemize}

### 2020-10-15 hpc_101/sections/03_policies.tex
Changed mentioned of UH-HPC to Mana Changed SLURM to Slurm as it is no longer an acronym.

@@
 
-\section[UH-HPC Usage Policies]{UH-HPC Usage Policies}
+\section[Mana Usage Policies]{Mana Usage Policies}
 
@@
 \begin{block}{}
-The following actions are acceptable on the shared systems in the UH-HPC:
+The following actions are acceptable on the shared systems in {\mana}:
 \begin{enumerate}
@@
 \item Accounts that are idle for a total of 180 days will be purged from the cluster
-\item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use the UH-HPC
-\item Users that will be leaving UH or are no longer active at UH, will need to have an active UH faculty/staff member in good standing, request a UH-HPC account extension.
+\item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use {\mana}
+\item Users that will be leaving UH or are no longer active at UH, will need to have an active UH faculty/staff member in good standing, request a {\mana} account extension.
 \end{enumerate}
@@
 \begin{frame}
-\frametitle{Where to find all UH-HPC specific policies}
+\frametitle{Where to find all Mana specific policies}
 \begin{enumerate}

### 2020-11-25 OOD/sections/02_walking.tex
Update to slides to point to caltech j.notebook instructions.  Merged in changes made for a class specific version to include where to contact us for support.

@@
 \section[Connecting]{Prerequisite \& Connecting}
-
 \begin{frame}
@@
 				\item An active Mana (UH-HPC) account
+				\begin{itemize}
+				\item Accounts provisioned for a class are active until the end of the semester
+				\end{itemize}
+				\item Registered for \href{http://www.hawaii.edu/its/uhlogin/}{MFA/DUO}
 				\item A web browser 
@@
 				\item Chrome\footnote[1,frame]{\tiny Chrome is the recommended browser}\footnote[2,frame]{\tiny \url{https://osc.github.io/ood-documentation/latest/requirements.html\#browser-requirements}}, Firefox, Edge are all supported
-				\item Any modern browser that supports ECMAScript 2016 should work
 		\end{itemize}
@@
 \end{frame}
+
+\section{User Support}
+\begin{frame}
+  \frametitle{User Support}
+  \begin{block}{Contact Information}
+   For help, please contact us at: UH-HPC-Help@lists.hawaii.edu
+  \end{block}
+  \begin{block}{Online documents \& FAQ}
+		\begin{itemize}
+		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
+		\item \href{http://go.hawaii.edu/jdG}{FAQ}
+		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}
+		\end{itemize}
+  \end{block}
+\end{frame}

### 2020-11-25 OOD/sections/03_custom.tex
update slide titles
update last slide title
Typo correction
typo correction
Update to slides to point to caltech j.notebook instructions.  Merged in changes made for a class specific version to include where to contact us for support.

@@
 \begin{frame}
-  \frametitle{Defaults and Quirks}
+  \frametitle{Rstudio - Defaults and Quirks}
 	\begin{block}{Defaults}
@@
 \begin{frame}
-	\frametitle{Lmod Collections}
+	\frametitle{Rstudio - Lmod Collections}
 	\begin{block}{}
@@
 \begin{frame}
-	\frametitle{Requirements \& procedure}
+	\frametitle{Rstudio - Requirements \& procedure}
 	\begin{block}{Requirements}
@@
 \section[Jupyternote]{Jupyter Notebooks}
-\subsection[jupenv]{Procedure}
+\subsection[jupenv]{Jupyter Notebook - Custom Environment Procedure}
 \begin{frame}[fragile]
-	\frametitle{Procedure}
+	\frametitle{Jupyter Notebook - Custom Environment Procedure}
 	\begin{block}{}
-	The following steps were adapted from \url{https://github.com/PSC-PublicHealth/pha-nbextensions}.~\\The following commands should be executed from a compute node in an interactive session prior to starting up a Juypter Notebook in Open OnDemand.
+	The following steps were adapted from \url{https://hpc.sites.caltech.edu/documentation/software-and-modules/jupyter-notebook}.~\\The following commands should be executed from a compute node in an interactive session prior to starting up a Juypter Notebook in Open OnDemand.
 	\end{block}
@@
 	\# \ldots 
-	\# \ldots 
-	\#\#\# Finalize setup with these last two packages \#\#\#
+	\#\#\# Finalize setup with this last package \#\#\#
 	[\$(test\_env)] conda install ipykernel
-	[\$(test\_env)] pip install git+https://github.com/PSC-PublicHealth/pha-nbextensions.git
+	[\$(test\_env)] python -m ipykernel install \ddash{}user \ddash{}name test\_env \ddash{}display-name "test\_env"
 	\#\#\# DONE and ready for use in Open OnDemand \#\#\#

### 2021-03-07 OOD/sections/01_intro.tex
Update to slides to point to caltech j.notebook instructions.  Merged in changes made for a class specific version to include where to contact us for support.
bring the github version up to current

@@
 
-\section[Why]{What can it do for me?}
+\section[Why]{What can it do for me at UH?}
 \begin{frame}
-	\frametitle{What can it do for me?}
+	\frametitle{What can it do for me at UH?}
 	\begin{itemize}
@@
 	\item Job management and monitoring
-	\item The ability to proxy graphical output from applications like Juypter notebooks and Rstudio Server executed on a compute node
+	\item The ability run interactive applications like Juypter lab/notebook, Rstudio Server and remote desktop on a node in Mana
 	\end{itemize}

### 2021-03-07 OOD/sections/03_custom.tex
bring the github version up to current

@@
 				\end{itemize}
-			\item The collection must be named "rstudio"
+			\item The collection must be named "ood\_rstudio"
 		\end{itemize}
@@
 				\end{itemize}
-			\item Save the module environment to the name "rstudio"
+			\item Save the module environment to the name "ood\_rstudio"
 				\begin{itemize}
-				\item module save rstudio
+				\item module save ood\_rstudio
 				\end{itemize}
@@
 
-\section[Jupyternote]{Jupyter Notebooks}
-\subsection[jupenv]{Jupyter Notebook - Custom Environment Procedure}
+\section[Jupyternote]{Jupyter Environments}
 \begin{frame}[fragile]
-	\frametitle{Jupyter Notebook - Custom Environment Procedure}
+	\frametitle{Jupyter- Custom Kernel Environment}
 	\begin{block}{}
@@
 	\end{center}	
+\end{frame}
+
+\begin{frame}[fragile]
+	\frametitle{Jupyter - Custom version}
+	\begin{block}{}Users can also install a custom version of Jupyter Lab/notebook by creating a custom collection with the name "`ood\_jupyterlab"'\end{block}
+	\begin{semiverbatim}\footnotesize
+[\$] srun -p sandbox -c 2 --mem=6 -t 60 --pty /bin/bash
+[\$] module load lang/Anaconda3
+[\$] conda create -n ood_jupyterlab -c conda-forge jupyterlab=3 xeus-python
+[\$] exit 
+# go to OOD and create an interactive session using the existing "jupyter
+notebook app/server" 
+# check Jupyter notebook's about for version to verify.
+ 	\end{semiverbatim}
 
 \end{frame}
+
+\begin{frame}
+  \frametitle{Jupyter Quirks}
+	\begin{block}{Quirks}
+			\begin{itemize}
+			\item jupyter applies all parameters from a global config file and those specified at run time \url{https://jupyterlab.readthedocs.io/en/latest/user/directories.html\#labconfig-directories}.
+			\item Users will need to remove or rename their global config file so as to not conflict with the Open OnDemand launcher.
+For example:\\ .jupyter/jupyter\_server\_config.py~->~.jupyter/jupyter\_server\_config\_custom.py
+		\end{itemize}
+	\end{block}
+\end{frame}

### 2021-03-09 hpc_101/sections/01_intro.tex
update to slides to have a full version (online) and a presentation version.
Next steps are to make it simpler to generate both versions

@@
     \begin{itemize}
-    \item The University~of~{\hawaii} High Performance Compute Cluster - \textbf{\mana} (formerly~known~as~the~{\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
+    \item The University~of~{\hawaii} High Performance Compute Cluster - \textbf{\mana} is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
     \item Community acquired nodes are equally accessible to all users
@@
 
-              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & Storage for lease \\
+              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & \thead{\textbf{Storage for lease}} \\
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 298                    & 6,308                      & 50 TB                   &  64                   & 80 TB                             &  150 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 345                    & 8,452                      & 62.9 TB                   &  120                   & 80 TB                             &  211 TB                        & 1 PB              \\ %\hline
 				\bottomrule
@@
 
-\subsection{Condo Program \& Leasing}
-\begin{frame}
-  \frametitle{What is the Condo Program \& leasing?}
-	\begin{block}{Condo Program}\scriptsize
-		\begin{itemize}
-		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with {\mana}
-		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
-		\item Condo nodes are managed and maintained by ITS staff
-		\item No maintenance fee will be assessed until the node is off warranty (typically five years)
-		\end{itemize}
-	\end{block}
-	\begin{block}{Leasing}\scriptsize
-		\begin{itemize}
-		\item Nodes leased from the community can have a contract period from one month, up to one year
-		\item Node lessees are provided priority access to their leased hardware
-		\item Leases are not considered an equipment purchase
-		\end{itemize}
-	\end{block}
-\end{frame}
+\subsection{Storage}
 
-\subsection{Storage}
-\begin{frame}
-  \frametitle{Storage}
-  {\mana} has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
-	\begin{columns}
-		\begin{column}{0.46\textwidth}
-		\begin{block}{Free Storage}
-			\begin{enumerate}
-			\item Permanent Storage
-				\begin{itemize}
-				\item Home Storage
-				\item Group/Lab Storage
-				\end{itemize}
-			\item Scratch Storage
-				\begin{itemize}
-%				\item {\lustre}(TBD)
-				\item Network File System (NFS)
-				\end{itemize}
-		\end{enumerate}
-		\end{block}
-		\end{column}
-		\begin{column}{0.46\textwidth}
-		\begin{block}{For Fee Storage}
-			\begin{enumerate}
-				\item Long Term Storage (LTS)
-		\end{enumerate}
-		\end{block}
-		\end{column}
-	\end{columns}      
-\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{frame}
+  %\frametitle{Storage}
+  %{\mana} has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
+	%\begin{columns}
+		%\begin{column}{0.46\textwidth}
+		%\begin{block}{Free Storage}
+			%\begin{enumerate}
+			%\item Permanent Storage
+				%\begin{itemize}
+				%\item Home Storage
+				%\item Group/Lab Storage
+				%\end{itemize}
+			%\item Scratch Storage
+				%\begin{itemize}
+				%\item {\lustre}
+				%\item Network File System (NFS)
+				%\end{itemize}
+		%\end{enumerate}
+		%\end{block}
+		%\end{column}
+		%\begin{column}{0.46\textwidth}
+		%\begin{block}{For Fee Storage}
+			%\begin{enumerate}
+				%\item Long Term Storage (LTS)
+		%\end{enumerate}
+		%\end{block}
+		%\end{column}
+	%\end{columns}      
+%\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \begin{enumerate}
-%    \item {\lustre (TBD)}
-%      \begin{itemize}
-%      \item High performance parallel file system 
-%      \item 50 TB quota
-%      \end{itemize}
+    \item {\lustre}
+      \begin{itemize}
+			\item A parallel distributed file system 
+      \end{itemize}
     \item NFS
       \begin{itemize}
-%      \item Not a parallel file system
-      \item 5 TB quota
+			\item A distributed file system 
       \end{itemize}
@@
   \begin{itemize}
+	\item 5 TB quota
   \item \textbf{Purge policy} -- 10 days based on file modify time
@@
 
-\subsection{Networking}
-\begin{frame}
-  \frametitle{Networking}
-	\begin{columns}
-	\begin{column}{0.46\textwidth}
-		\begin{block}{Internal Networks}
-		\begin{itemize}
-			\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
-				\begin{itemize}
-					\item 40 Gbit
-					\item Older compute nodes
-					\item low latency ($\approx1.3\mu$s)
-					\item non-blocking
-				\end{itemize}
-				~\\
-			\item 25/100 Gbit Ethernet
-				\begin{itemize}
-					\item Newer compute nodes
-					\item Nodes connected @ 25~Gbit
-					\item non-blocking
-				\end{itemize}
-	        \end{itemize}
-                \end{block}
-	\end{column}
-	\begin{column}{0.46\textwidth}
-		\begin{block}{External Networks}
-		\begin{itemize}
-			\item 100 Gbit SciDMZ
-				\begin{itemize}
-					\item Login \& compute nodes\\connected via a firewall
-					\item DTNs are directly connected
-				\end{itemize}
-		\end{itemize}
-                \end{block}
-	\end{column}
-	\end{columns}
-\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\subsection{Networking}
+%\begin{frame}
+  %\frametitle{Networking}
+	%\begin{columns}
+	%\begin{column}{0.46\textwidth}
+		%\begin{block}{Internal Networks}
+		%\begin{itemize}
+			%\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
+				%\begin{itemize}
+					%\item 40 Gbit
+					%\item low latency ($\approx1.3\mu$s)
+					%\item non-blocking
+				%\end{itemize}
+			%\item High Data Rate (HDR) IB   
+				%\begin{itemize}
+					%\item 100 Gbit
+					%\item low latency ($\approx0.5\mu$s)
+					%%\item non-blocking
+				%\end{itemize}
+			%\item 1/25/100 Gbit Ethernet
+%
+	        %\end{itemize}
+                %\end{block}
+	%\end{column}
+	%\begin{column}{0.46\textwidth}
+		%\begin{block}{External Networks}
+		%\begin{itemize}
+			%\item 100 Gbit SciDMZ
+				%\begin{itemize}
+					%\item Login \& compute nodes\\connected via a firewall
+					%\item DTNs are directly connected
+				%\end{itemize}
+		%\end{itemize}
+                %\end{block}
+	%\end{column}
+	%\end{columns}
+%\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{frame}
+
+\subsection{Condo Program \& Leasing}
+\begin{frame}
+  \frametitle{What is the Condo Program \& leasing?}
+	\begin{block}{Condo Program}\scriptsize
+		\begin{itemize}
+		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with {\mana}
+		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
+		\item Condo nodes are managed and maintained by ITS staff
+		\item Nodes are purchased with a five year warranty
+		\end{itemize}
+	\end{block}
+	\begin{block}{Leasing}\scriptsize
+		\begin{itemize}
+		\item Nodes leased from the community can have a contract period from one month, up to one year
+		\item Node lessees are provided priority access to their leased hardware
+		\item Leases are not considered an equipment purchase
+		\end{itemize}
+	\end{block}
+\end{frame}

### 2022-02-04 hpc_101/sections/01_intro.tex
update to slides

@@
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 345                    & 8,452                      & 62.9 TB                   &  120                   & 80 TB                             &  211 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 346                    & 8,500                      & 63.2 TB                   &  120                   & 80 TB                             &  111 TB                        & 1 PB              \\ %\hline
 				\bottomrule
@@
 	\item 5 TB quota
-  \item \textbf{Purge policy} -- 10 days based on file modify time
+  \item \textbf{Purge policy} -- 20 days based on file modify time
   \item Available on all nodes
@@
   \begin{block}{Online documents \& FAQ}
-    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ.~\\
+    Users are encouraged to look through the online documentation \& User Guide prior to contacting ITS-CI directly.
+Many questions we receive are repeat questions and we try to capture them in our User Guide.~\\
 		\begin{itemize}
 		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
-		\item \href{http://go.hawaii.edu/jdG}{FAQ}
+		\item \href{http://go.hawaii.edu/JLA}{Resource Documentation \& User Guide}
 		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}

### 2022-09-12 hpc_101/sections/03_policies.tex
update to slides
update the slides to contain updated info.

@@
 Users are responsible for their own data and are highly encouraged to backup data that is important to them at other storage locations
-\item Files located on the scratch file systems are subject to a \textbf{10 day} purge policy that is based on the file/directory modified timestamp.
+\item Files located on the scratch file systems are subject to a \textbf{20 day} purge policy that is based on the file/directory modified timestamp.
 \textbf{Files that are purge cannot be recovered}.
@@
 
-\begin{frame}
-\frametitle{Where to find all Mana specific policies}
-\begin{enumerate}
-%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes}
-%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{University of Hawaii Executive Policy E2.210}
-\item \href{http://go.hawaii.edu/GSY}{Common systems \& Storage}
-\item \href{http://go.hawaii.edu/0SG}{User Account life cycle}
-\item \href{http://go.hawaii.edu/WSG}{Security}
-\item \href{http://go.hawaii.edu/YKG}{For lease storage}
-\item \href{http://go.hawaii.edu/GK0}{Condo Program \& Off Warranty condo nodes}
-%% \item \href{https://docs.google.com/document/d/1mGcCAsmzGT_hWXLcUNfT2LUwiggCnhqrln32k0IGZZI/}{Special allocations}
-\end{enumerate}
-\end{frame}
+%\begin{frame}
+%\frametitle{Where to find all Mana specific policies}
+%\begin{enumerate}
+%%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html\#appendixa}{Chapter 708, Hawaii Revised Statutes}
+%%\item \href{http://www.hawaii.edu/infotech/policies/itpolicy.html}{University of Hawaii Executive Policy E2.210}
+%\item \href{http://go.hawaii.edu/GSY}{Common systems \& Storage}
+%\item \href{http://go.hawaii.edu/0SG}{User Account life cycle}
+%\item \href{http://go.hawaii.edu/WSG}{Security}
+%\item \href{http://go.hawaii.edu/YKG}{For lease storage}
+%\item \href{http://go.hawaii.edu/GK0}{Condo Program \& Off Warranty condo nodes}
+%%% \item \href{https://docs.google.com/document/d/1mGcCAsmzGT_hWXLcUNfT2LUwiggCnhqrln32k0IGZZI/}{Special allocations}
+%\end{enumerate}
+%\end{frame}
 %

### 2023-04-30 hpc_101/sections/09_appendix.tex
Added appendix and moved terminology

--- 2020-10-15 hpc_101/sections/01_intro.tex
@@
-\part{Introduction}
-\begin{frame}
-			 \partpage
-\end{frame}
+\part{Appendix}
 
@@
 \end{frame}
-
-\section[University of Hawai'i High Performance Compute Cluster]{University of Hawai'i High Performance Compute Cluster}
-\begin{frame}
-    \frametitle{University of {\hawaii} High Performance Compute Cluster}
-    \begin{itemize}
-    \item The University~of~{\hawaii} High Performance Compute Cluster - \textbf{\mana} (formerly~known~as~the~{\uhhpc}) is \textbf{free} to use for all active faculty, staff, and students affiliated with the University of {\hawaii}
-    \item Community acquired nodes are equally accessible to all users
-    \item Nodes purchased through the {\textbf{condo program}} are shared with the community, but priority is given to the node owner and their agents
-		\item Nodes may be {\textbf{leased}} from the community pool
-		\item Additional permanent storage can be leased by faculty \& staff
-    \end{itemize}
-		\begin{block}{Mana Resource Summary}
-  \begin{table}
-    \centering
-    \resizebox{\textwidth}{!}{%
-      \begin{tabular}{l||l||l||l||l||l||l||l}
-      %  \multicolumn{8}{c}{ {\large {\textbf{UH-HPC Resource Summary} } } } \\ 
-			\toprule
-
-              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & Storage for lease \\
-							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 298                    & 6,308                      & 50 TB                   &  64                   & 80 TB                             &  150 TB                        & 1 PB              \\ %\hline
-				\bottomrule
-      \end{tabular}%
-    }
-  \end{table}
-	\end{block}
-\end{frame}
-
-\subsection{Condo Program \& Leasing}
-\begin{frame}
-  \frametitle{What is the Condo Program \& leasing?}
-	\begin{block}{Condo Program}\scriptsize
-		\begin{itemize}
-		\item The condo program allows faculty \& staff to purchase nodes and have them integrated with {\mana}
-		\item Condo nodes can take advantage of networking and storage infrastructure that one may not typically have access to
-		\item Condo nodes are managed and maintained by ITS staff
-		\item No maintenance fee will be assessed until the node is off warranty (typically five years)
-		\end{itemize}
-	\end{block}
-	\begin{block}{Leasing}\scriptsize
-		\begin{itemize}
-		\item Nodes leased from the community can have a contract period from one month, up to one year
-		\item Node lessees are provided priority access to their leased hardware
-		\item Leases are not considered an equipment purchase
-		\end{itemize}
-	\end{block}
-\end{frame}
-
-\subsection{Storage}
-\begin{frame}
-  \frametitle{Storage}
-  {\mana} has two classes and three types of storage that users can potentially access.\\Each type of storage has their own attributes and restrictions
-	\begin{columns}
-		\begin{column}{0.46\textwidth}
-		\begin{block}{Free Storage}
-			\begin{enumerate}
-			\item Permanent Storage
-				\begin{itemize}
-				\item Home Storage
-				\item Group/Lab Storage
-				\end{itemize}
-			\item Scratch Storage
-				\begin{itemize}
-%				\item {\lustre}(TBD)
-				\item Network File System (NFS)
-				\end{itemize}
-		\end{enumerate}
-		\end{block}
-		\end{column}
-		\begin{column}{0.46\textwidth}
-		\begin{block}{For Fee Storage}
-			\begin{enumerate}
-				\item Long Term Storage (LTS)
-		\end{enumerate}
-		\end{block}
-		\end{column}
-	\end{columns}      
-\end{frame}
-
-\begin{frame}
-  \frametitle{Permanent Storage}
-  \begin{enumerate}
-  \item Home Storage
-    \begin{itemize}
-    \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on {\mana}
-    \end{itemize}
-  \item Group/Lab Storage
-    \begin{itemize}
-    \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on {\mana}
-    \end{itemize}
-  \end{enumerate}
-~\\
-  Permanent storage options on {\mana} have the following attributes:
-  \begin{itemize}
-  \item 50 GB default quota with a max quota of 300 GB
-  \item Quota increases are re-evaluated annually
-  \item Available on all nodes
-  \item Freely available to users
-  \end{itemize}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Scratch Storage}
-  \begin{enumerate}
-%    \item {\lustre (TBD)}
-%      \begin{itemize}
-%      \item High performance parallel file system 
-%      \item 50 TB quota
-%      \end{itemize}
-    \item NFS
-      \begin{itemize}
-%      \item Not a parallel file system
-      \item 5 TB quota
-      \end{itemize}
-  \end{enumerate}
-~\\
-  Scratch file systems on {\mana} have the following attributes:
-  \begin{itemize}
-  \item \textbf{Purge policy} -- 10 days based on file modify time
-  \item Available on all nodes
-  \item Freely available to users
-  \end{itemize}
-
-\end{frame}
-
-%\begin{frame}
-  %\frametitle{For Fee Storage}
-  %\begin{enumerate}
-    %\item ValueStorage
-      %\begin{itemize}
-      %\item Enterprise class equipment 
-      %\item 10 Gbit connected to the UH-HPC
-      %\item Currently only accessible on the DTNs and login nodes
-      %\end{itemize}
-    %\item LTS  
-      %\begin{itemize}
-      %\item Cheap-n-Deep
-      %\item 100 Gbit connected to the UH-HPC
-      %\item Accessible on all nodes
-      %\end{itemize}
-  %\end{enumerate}
-%\end{frame}
-
-\subsection{Networking}
-\begin{frame}
-  \frametitle{Networking}
-	\begin{columns}
-	\begin{column}{0.46\textwidth}
-		\begin{block}{Internal Networks}
-		\begin{itemize}
-			\item Quad Data Rate (QDR)\\InfiniBand{\regtrademark} (IB)   
-				\begin{itemize}
-					\item 40 Gbit
-					\item Older compute nodes
-					\item low latency ($\approx1.3\mu$s)
-					\item non-blocking
-				\end{itemize}
-				~\\
-			\item 25/100 Gbit Ethernet
-				\begin{itemize}
-					\item Newer compute nodes
-					\item Nodes connected @ 25~Gbit
-					\item non-blocking
-				\end{itemize}
-	        \end{itemize}
-                \end{block}
-	\end{column}
-	\begin{column}{0.46\textwidth}
-		\begin{block}{External Networks}
-		\begin{itemize}
-			\item 100 Gbit SciDMZ
-				\begin{itemize}
-					\item Login \& compute nodes\\connected via a firewall
-					\item DTNs are directly connected
-				\end{itemize}
-		\end{itemize}
-                \end{block}
-	\end{column}
-	\end{columns}
-\end{frame}
-
-\subsection{User Support}
-\begin{frame}
-  \frametitle{User Support}
-
-  \begin{block}{Online documents \& FAQ}
-    Users are encouraged to look through the online documentation \& FAQ prior to contacting ITS-CI directly.  Many questions we receive are repeat questions and we try to capture them in our FAQ.~\\
-		\begin{itemize}
-		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
-		\item \href{http://go.hawaii.edu/jdG}{FAQ}
-		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}
-		\end{itemize}
-  \end{block}
-  \begin{block}{Contact Information}
-    If your question is not answered in our online documentation, please contact us at: UH-HPC-Help@lists.hawaii.edu
-
-    \begin{itemize}
-    \item For batch jobs \ldots
-      \begin{itemize}
-      \item[--] Job ID, path to submission script, submission command, error file location, output files
-      \end{itemize} 
-    \item For other problems \ldots
-      \begin{itemize}
-      \item[--] State the problem, command issued, host, directory, remote host, error messages
-      \end{itemize}
-    \end{itemize}
-  \end{block}
-\end{frame}

### 2023-04-30 hpc_101/sections/01_intro.tex
update the slides to contain updated info.
Group/lab storage creation point

@@
 			 \partpage
-\end{frame}
-
-\section[Terminology]{Terminology}
-
-\begin{frame}
-	\frametitle{Terminology}
-	\begin{itemize}
-	\item \textbf{Node} -- Another name for a server or computer
-        \item \textbf{Login node} -- A specialized node that users connect to in order to submit work to a computer cluster
-        \item \textbf{Computer cluster} -- A set of loosely or tightly connected nodes that work together so that, in many respects, they can be viewed as a single system\footnote{\label{wiki_ccluster}\tiny\url{https://en.wikipedia.org/wiki/Computer_cluster}}
-        \item \textbf{Data transfer node (DTN)} -- Specialized nodes that minimize the impedance on the network to access the full capability of the network
-	\item \textbf{Science DMZ (SciDMZ)} --  A portion of the network configured with equipment and security policies in order to optimize for high-performance scientific applications rather than for general-purpose business systems or “enterprise” computing
-	\item \textbf{Multi-factor Authentication (MFA)} -- An authentication method in which a computer user is granted access only after successfully presenting two or more pieces of evidence or factors to an authentication mechanism, e.g., DUO
-
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Terminology}
-	\begin{itemize}
-	\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
-	\item \textbf{Command-line interface/interpreter (CLI)} -- A text-based user interface used to view and manage computer files 
-	\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
-	\item \textbf{High Performance Compute (HPC)} -- A computing paradigm in which applications are typically a tightly coupled parallel job that benefit from a low-latency interconnect
-	\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
-        \item \textbf{Modified time} -- The last time the file was modified (content has been modified)\footnote{\label{Types of timestamps}\tiny\url{https://unix.stackexchange.com/a/2465}}
-        \item \textbf{Shell script (script)} -- A computer program designed to be run by a CLI\footnote{\label{Shell script}\tiny\url{https://en.wikipedia.org/wiki/Shell_script}}
-%        \item Pleasantly Parallel -- Processes are independent and no communication is necessary 
-	\end{itemize}
-
 \end{frame}
@@
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 346                    & 8,500                      & 63.2 TB                   &  120                   & 80 TB                             &  111 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 366                    & 9,396                      & 67.3 TB                   &  120                   & 80 TB                             &  111 TB                        & 1 PB              \\ %\hline
 				\bottomrule
@@
     \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on {\mana}
+    \item PI must contact \textbf{uh-hpc-help@lists.hawaii.edu} to set up group/lab storage
     \end{itemize}
@@
 		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
-		\item \href{http://go.hawaii.edu/JLA}{Resource Documentation \& User Guide}
+		\item \href{http://go.hawaii.edu/ex2}{Resource Documentation \& User Guide}
 		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}

### 2023-05-03 hpc_101/sections/09_appendix.tex
monitoring examples added

@@
 \end{frame}
+\section{Monitoring Jobs and Partitions}{Monitoring Jobs and Partitions}
+\begin{frame}[fragile]
+\frametitle{Monitoring Running Jobs}
+\begin{block}{\textbf{squeue}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ squeue --me
+             JOBID PARTITION     NAME     USER ST       TIME  NODES NODELIST(REASON)
+          43486015   sandbox     bash victorgc  R    2:05:37      1 node-0093
+\end{semiverbatim}
+\end{block}
+The --me flag can be removed and \textbf{squeue} will report all running jobs in the system.
+\begin{block}{\textbf{squeue}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ squeue | head
+             JOBID PARTITION     NAME     USER ST       TIME  NODES NODELIST(REASON)
+          43078226   bioinfo    sgSQL     user  R 23-02:12:35      1 node-0327
+          43078227   bioinfo cromwell     user  R 23-02:12:09      1 node-0327
+          43485890   bioinfo       QC     user  R    4:45:59      1 node-0327
+          43399271   bioinfo      TQC     user  R 3-20:49:10      1 node-0328
+          43485868 exclusive    MoSe2     user  R    5:20:14      1 node-0296
+          43485867 exclusive    MoSe2     user  R    5:30:46      1 node-0295
+          43485866 exclusive    MoSe2     user  R    5:35:43      1 node-0294
+          43485815 exclusive    MoSe2     user  R    5:57:03      1 node-0281
+\end{semiverbatim}
+\end{block}
+\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/squeue.html}{https://slurm.schedmd.com/squeue.html}}
+
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Completed Jobs Reports}
+\begin{block}{\textbf{sacct}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ sacct -u victorgc -X | head
+JobID           JobName  Partition    Account  AllocCPUS      State ExitCode
+------------ ---------- ---------- ---------- ---------- ---------- --------
+43486049_1     spectrum     shared         uh          1  COMPLETED      0:0
+43486049_12    spectrum     shared         uh          1  COMPLETED      0:0
+\end{semiverbatim}
+\end{block}
+\begin{block}{\textbf{seff}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ seff 43486049_12
+Job ID: 43486061
+Array Job ID: 43486049_12
+Cluster: uh_hpc
+User/Group: victorgc/victorgc
+State: COMPLETED (exit code 0)
+Cores: 1
+CPU Utilized: 00:00:33
+CPU Efficiency: 97.06\% of 00:00:34 core-walltime
+Job Wall-clock time: 00:00:34
+Memory Utilized: 75.91 MB
+Memory Efficiency: 1.27\% of 5.86 GB
+\end{semiverbatim}
+\end{block}
+The output of \textbf{sacct} is highly customizable.
+\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/sacct.html}{https://slurm.schedmd.com/sacct.html}}
+
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Partition Information}
+\begin{block}{\textbf{sinfo}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ sinfo | head
+PARTITION       AVAIL  TIMELIMIT  NODES  STATE NODELIST
+gpu                up 3-00:00:00      9    mix gpu-[0013,0017-0024]
+gpu                up 3-00:00:00      1  alloc gpu-0016
+sandbox            up    4:00:00      1    mix node-0093
+sandbox            up    4:00:00      3  alloc node-[0094-0096]
+shared*            up 3-00:00:00      4  down* node-[0102,0120,0122,0178]
+shared*            up 3-00:00:00      8  drain node-[0111-0112,0126,0130,0150,0175,0344,0350]
+\end{semiverbatim}
+\end{block}
+Partition information can be printed in a per node basis.
+\begin{block}{\textbf{sinfo}}
+\begin{semiverbatim}\tiny \texttt
+[victorgc@node-0093 HDM]\$ sinfo -Nel | head
+Wed May 03 01:20:33 2023
+NODELIST   NODES       PARTITION       STATE CPUS    S:C:T MEMORY TMP_DISK WEIGHT AVAIL_FE REASON
+gpu-0001       1     gpu-sandbox       mixed 20     2:10:1 128500        0  10000 x86,inte none
+gpu-0001       1  kill-exclusive       mixed 20     2:10:1 128500        0  10000 x86,inte none
+gpu-0001       1     kill-shared       mixed 20     2:10:1 128500        0  10000 x86,inte none
+gpu-0002       1  kill-exclusive        idle 20     2:10:1 128500        0  10000 x86,inte none
+gpu-0002       1     kill-shared        idle 20     2:10:1 128500        0  10000 x86,inte none
+\end{semiverbatim}
+\end{block}
+\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/sinfo.html}{https://slurm.schedmd.com/sinfo.htmls}}
+\end{frame}

### 2023-05-06 hpc_101/sections/01_intro.tex
LTS frame added

@@
 
+\begin{frame}
+  \frametitle{Permanent Storage }
+  \begin{enumerate}
+  \item Long Term Storage (LTS)
+    \begin{itemize}
+    \item {\textbf{Purpose}}: \footnotesize Additional Personal and/or group storage for applications and active data when all other storage options have been exhausted{\mana}\footnote{\label{LTS}\tiny\ \url{https://hawaii.kualibuild.com/app/builder/app/6062a8170d0d97001cb7d3af/run}}
+
+    \end{itemize}
+  \end{enumerate}
+~\\
+  LTS storage on {\mana} has the following attributes:
+  \begin{itemize}
+  \item \$20/TB with a minimum purchase of 2TB 
+  \item LTS purchases are done in a per year basis
+  \item Available on all nodes
+  \end{itemize}
+
+\end{frame}
+
 %\begin{frame}

### 2023-05-06 hpc_101/sections/09_appendix.tex
small corrections

@@
 \end{frame}
-\section{Monitoring Jobs and Partitions}{Monitoring Jobs and Partitions}
+\section[Monitoring Jobs and Partitions]{Monitoring Jobs and Partitions}
 \begin{frame}[fragile]
@@
 \begin{semiverbatim}\tiny \texttt
-[victorgc@node-0093 HDM]\$ squeue --me
+[victorgc@node-0093 HDM]\$ squeue -u victorgc
              JOBID PARTITION     NAME     USER ST       TIME  NODES NODELIST(REASON)
@@
 \end{block}
-The --me flag can be removed and \textbf{squeue} will report all running jobs in the system.
+By default, \textbf{squeue} will report all running jobs in the system.
 \begin{block}{\textbf{squeue}}

### 2023-06-22 hpc_101/sections/01_intro.tex
Added information about Mana migration to Koa

--- 2023-04-30 hpc_101/sections/01_intro.tex
@@
     \item Nodes purchased through the {\textbf{condo program}} are shared with the community, but priority is given to the node owner and their agents
+    \itmem Mana is currently undergoing a migration, resource summary subject to change. 
 		\item Nodes may be {\textbf{leased}} from the community pool
@@
     \end{itemize}
-		\begin{block}{Mana Resource Summary}
+		\begin{block}{Mana Resource Summary\footnotemark}
   \begin{table}
@@
 	\end{block}
+ \footnotetext{Mana is currently undergoing a migration, resource summary subject to change over the next few months.}
 \end{frame}
@@
   \end{itemize}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Permanent Storage }
+  \begin{enumerate}
+  \item Long Term Storage (LTS)
+    \begin{itemize}
+    \item {\textbf{Purpose}}: \footnotesize Additional Personal and/or group storage for applications and active data when all other storage options have been exhausted in {\mana}\footnote{\label{LTS}\tiny\ \url{https://hawaii.kualibuild.com/app/builder/app/6062a8170d0d97001cb7d3af/run}}
+
+    \end{itemize}
+  \end{enumerate}
+~\\
+  LTS storage on {\mana} has the following attributes:
+  \begin{itemize}
+  \item \$20/TB with a minimum purchase of 2TB 
+  \item LTS purchases are done in a per year basis
+  \item Available on all nodes
+  \end{itemize}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Mana Migration}
+~\\
+  The new Mana/Koa cluster will have the following attributes:
+  \begin{itemize}
+	  \item {\textbf{Storage}}: New separate storage system. Files will need to be migrated, and files remaining in Mana will be deleted once the old cluster is turned off. 
+	  \item{\textbf{Home}}:Permanent storage with 50GB quota for personal applications and data
+	  \item{\textbf{Scratch}}: 800 TB and 400,000,000 file quota shared across all users with a 90 day purge policy. 
+  \item{\textbf{Workspace}}: Factulty and staff members will allocated 1TB of workspace upon request, to be shared by lab and group members. 
+  \item {\textbf{Timeline}}: Progressive migration in 6 differnt stages \footnotemark
+  \end{itemize}
+  \footnotetext{\tiny\url{https://datascience.hawaii.edu/mana-upgrade/}}
 

### 2024-01-18 hpc_101/sections/01_intro.tex
workspace info to TBD
update to slides for onboarding

--- 2023-04-30 hpc_101/sections/01_intro.tex
@@
 \end{frame}
+
+%\section[Terminology]{Terminology}
+
+%\begin{frame}
+	%\frametitle{Terminology}
+	%\begin{itemize}
+	%\item \textbf{Node} -- Another name for a server or computer
+        %\item \textbf{Login node} -- A specialized node that users connect to in order to submit work to a computer cluster
+        %\item \textbf{Computer cluster} -- A set of loosely or tightly connected nodes that work together so that, in many respects, they can be viewed as a single system\footnote{\label{wiki_ccluster}\tiny\url{https://en.wikipedia.org/wiki/Computer_cluster}}
+        %\item \textbf{Data transfer node (DTN)} -- Specialized nodes that minimize the impedance on the network to access the full capability of the network
+	%\item \textbf{Science DMZ (SciDMZ)} --  A portion of the network configured with equipment and security policies in order to optimize for high-performance scientific applications rather than for general-purpose business systems or “enterprise” computing
+	%\item \textbf{Multi-factor Authentication (MFA)} -- An authentication method in which a computer user is granted access only after successfully presenting two or more pieces of evidence or factors to an authentication mechanism, e.g., DUO
+%
+	%\end{itemize}
+%\end{frame}
+
+%\begin{frame}
+	%\frametitle{Terminology}
+	%\begin{itemize}
+	%\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
+	%\item \textbf{Command-line interface/interpreter (CLI)} -- A text-based user interface used to view and manage computer files 
+	%\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
+	%\item \textbf{High Performance Compute (HPC)} -- A computing paradigm in which applications are typically a tightly coupled parallel job that benefit from a low-latency interconnect
+	%\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
+        %\item \textbf{Modified time} -- The last time the file was modified (content has been modified)\footnote{\label{Types of timestamps}\tiny\url{https://unix.stackexchange.com/a/2465}}
+        %\item \textbf{Shell script (script)} -- A computer program designed to be run by a CLI\footnote{\label{Shell script}\tiny\url{https://en.wikipedia.org/wiki/Shell_script}}
+        %\item Pleasantly Parallel -- Processes are independent and no communication is necessary 
+	%\end{itemize}
+%
+%
+%\end{frame}
+%
 
@@
     \end{itemize}
-		\begin{block}{Mana Resource Summary}
+		\begin{block}{Koa Resource Summary}
   \begin{table}
@@
 
-              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home/Group Space}} & \thead{\textbf{Scratch Space}} & \thead{\textbf{Storage for lease}} \\
+              {}               & \thead{\textbf{Nodes}} & \thead{\textbf{CPU Cores}} & \thead{\textbf{Memory}} & \thead{\textbf{GPUs}} & \thead{\textbf{Home Space}} & \thead{\textbf{Scratch Space}} & \thead{\textbf{Storage for lease}} \\
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 366                    & 9,396                      & 67.3 TB                   &  120                   & 80 TB                             &  111 TB                        & 1 PB              \\ %\hline
+        \thead{\textbf{Total}} & 269                    & 7,501                      & 53 TB                   &  146                   & 50 GB per user                             &  800 TB                        & 2 PB              \\ %\hline
 				\bottomrule
@@
     \item {\textbf{Purpose}}: Personal storage for applications and active data that needs to persist on {\mana}
+		\item {\textbf{Quota}}: 50 GB per user
     \end{itemize}
-  \item Group/Lab Storage
+  \item Lab Storage
     \begin{itemize}
-    \item {\textbf{Purpose}}: Group/Lab storage allows users to share data \& applications with a need for persistence on {\mana}
-    \item PI must contact \textbf{uh-hpc-help@lists.hawaii.edu} to set up group/lab storage
+    \item {\textbf{Purpose}}: Lab storage allows users to share data \& applications with a need for persistence on {\mana}
+		\item {\textbf{Quota}}: 500 GB and 512,000 files/folders (inodes)
     \end{itemize}
@@
 ~\\
-  Permanent storage options on {\mana} have the following attributes:
+  Permanent storage options on {\mana} share the following attributes:
   \begin{itemize}
-  \item 50 GB default quota with a max quota of 300 GB
-  \item Quota increases are re-evaluated annually
   \item Available on all nodes
@@
       \end{itemize}
-    \item NFS
-      \begin{itemize}
-			\item A distributed file system 
-      \end{itemize}
+    %\item NFS
+      %\begin{itemize}
+			%\item A distributed file system 
+      %\end{itemize}
   \end{enumerate}
@@
   \begin{itemize}
-	\item 5 TB quota
-  \item \textbf{Purge policy} -- 20 days based on file modify time
+	\item 800 TB and 400K file/folder (inode) quota shared by all users
+  \item \textbf{Purge policy} -- 90 days based on file modify time
   \item Available on all nodes
@@
 
-%\begin{frame}
-  %\frametitle{For Fee Storage}
-  %\begin{enumerate}
-    %\item ValueStorage
-      %\begin{itemize}
-      %\item Enterprise class equipment 
-      %\item 10 Gbit connected to the UH-HPC
-      %\item Currently only accessible on the DTNs and login nodes
-      %\end{itemize}
-    %\item LTS  
-      %\begin{itemize}
-      %\item Cheap-n-Deep
-      %\item 100 Gbit connected to the UH-HPC
-      %\item Accessible on all nodes
-      %\end{itemize}
-  %\end{enumerate}
-%\end{frame}
+\begin{frame}
+  \frametitle{Storage as a Service}
+  \begin{enumerate}
+    \item KoaStore
+      \begin{itemize}
+      \item \textbf{{Purpose:}} Additional permanent storage for applications and active data on {\mana}.
+			\item \$50/TB/year
+			\item Purchased in year increments
+			\item Available on all nodes
+      \end{itemize}
+    \item KoaCloud
+      \begin{itemize}
+      \item \textbf{{Purpose:}} Dropbox/Google Drive like storage that can be used outside of {\mana} for research data.
+			\item \$55/TB/year
+			\item Purchased in year increments
+      \end{itemize}
+  \end{enumerate}
+\end{frame}
 

### 2024-01-18 hpc_101/sections/03_policies.tex
Removed account purge policy
update to slides for onboarding

@@
 
-\section[Mana Usage Policies]{Mana Usage Policies}
+\section[Koa Usage Policies]{Koa Usage Policies}
 
-\subsection{Login nodes \& DTN nodes}
+\subsection{Login nodes}
 \begin{frame}
-\frametitle{Login nodes \& DTN nodes}
+\frametitle{Login nodes}
 \begin{block}{}
-The following actions are acceptable on the shared systems in {\mana}:
+The following actions are acceptable on the login nodes of {\mana}:
 \begin{enumerate}
-\item File/Directory management [ Login \& DTN nodes ]
-\item Text editing with a text editor: vi/vim, emacs, nano [ Login \& DTN nodes ]
-\item Transferring files to and from the cluster (scp, rsync, SFTP, globus, aspera, lftp, etc.) [ Login \& DTN nodes ]
-\item Shall be used to submit batch and interactive jobs [Login nodes]
-\item SSH shell access [ Login \& DTN nodes ]
+\item File/Directory management
+\item Text editing with a text editor: vi/vim, emacs, nano
+\item Transferring files to and from the cluster (scp, rsync, SFTP, etc.)
+\item Submit batch and interactive jobs
+\item SSH shell access
 \end{enumerate}
@@
 Users are responsible for their own data and are highly encouraged to backup data that is important to them at other storage locations
-\item Files located on the scratch file systems are subject to a \textbf{20 day} purge policy that is based on the file/directory modified timestamp.
+\item Files located on the scratch file systems are subject to a \textbf{90 day} purge policy that is based on the file/directory modified timestamp.
 \textbf{Files that are purge cannot be recovered}.
@@
 \item Accounts that have been locked can be unlocked upon request with no further action then emailing us a request to unlock their account
-\item Accounts that are idle for a total of 180 days will be purged from the cluster
-\item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use {\mana}
+%\item Accounts that are idle for a total of 180 days will be purged from the cluster
+%\item Users that have been purged may request a reactivation of their account, but may be required to take a quiz to confirm retention of how to use {\mana}
 \item Users that will be leaving UH or are no longer active at UH, will need to have an active UH faculty/staff member in good standing, request a {\mana} account extension.

### 2025-03-14 hpc_101/sections/00_title.tex
corrections to early slides of hpc101, addition of new members, correction to node prices.
added another team member
switch over to new presentation for new cluster
update to slides for stale information
update the slides to contain updated info.
update to slides for onboarding
update cluster stats and group info

@@
 
-\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
-\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
+\author{Acting Director: Sean Cleveland\\Adriana Comerford, Ron Merrill, David Schanzenbach}
+%\author{Director: Gwen Jacobs\\Associate Director: Sean Cleveland\\Victor Goicoechea Casanueva,\\Ron Merrill, David Schanzenbach}
+%\author{Director: Gwen Jacobs\\Sean Cleveland, Adriana Comerford, Jennifer Geis,\\Ron Merrill, David Schanzenbach}
+\institute[University of {\hawaii} -- ITS--RCI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
 {
@@
 \medskip
-\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
+\textbf{\url{https://datascience.hawaii.edu/}}\\
+%\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
 \textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address

### 2025-03-14 hpc_101/sections/01_intro.tex
update cluster stats and group info

@@
 							\midrule  \midrule %\hline
-        \thead{\textbf{Total}} & 269                    & 7,501                      & 53 TB                   &  146                   & 50 GB per user                             &  800 TB                        & 2 PB              \\ %\hline
+        \thead{\textbf{Total}} & 287                    & 8,559                      & 61 TB                   &  140                   & 50 GB per user                             &  800 TB                        & 2 PB              \\ %\hline
 				\bottomrule
@@
   \begin{itemize}
-	\item 800 TB and 400K file/folder (inode) quota shared by all users
+	\item 800 TB and 400M file/folder (inode) quota shared by all users
   \item \textbf{Purge policy} -- 90 days based on file modify time

### 2025-03-14 hpc_101/sections/01_intro.tex
rename to ITS-RCI

@@
   \begin{block}{Online documents \& FAQ}
-    Users are encouraged to look through the online documentation \& User Guide prior to contacting ITS-CI directly.
+    Users are encouraged to look through the online documentation \& User Guide prior to contacting ITS-RCI directly.
 Many questions we receive are repeat questions and we try to capture them in our User Guide.~\\
